### 2017-05-16 doc/response.tex
paper files added

\documentclass{article}
\usepackage{a4wide}
\usepackage{xr-hyper} 
\usepackage{hyperref}
\externaldocument{paper}
\title{Response to the Reviewers}
%\author{Ákos Kádár \and Grzegorz Chrupala \and Afra Alishahi}
\author{Submission number: 882 \and
Action editor:  Christopher Potts }
\date{}
\begin{document}

\maketitle
The Editor's and reviewers' comments were of great help and enabled us
to make important improvements on our previous submission. Some minor or major
changes were made to almost all of the sections of the paper.
One of the major changes in the revised version is the addition of
discussions on the generality and applicability of the proposed
methods in sections \ref{sec:intro}, \ref{sec:computeomission}, 
\ref{sec:beyondlexical} and \ref{sec:reprdim}. The revised version
still focuses on the {\sc Imaginet} architecture, but in addition 
allocates more space and more emphasis on the discussion
of the introduced analysis techniques.
The other major revision concerns the re-structuring 
of Section \ref{sec:macro}. In the revised version we 
separated the description of the omission score computation 
(Section \ref{sec:computeomission}) and
the discussion regarding the distribution of the omission scores over 
categories of words for the different models 
(Section \ref{sec:omitimaginet}). 
Furthermore, in Section \ref{sec:beyondlexical}
we introduced a common methodology
to assess the impact of position and grammatical
function of tokens (while controlling for word-types) based on the 
omission scores. We also provide finer-grained analysis
of these factors separately in sections \ref{sec:gramfunc} and
\ref{subsec:information-struct}. Thanks to the reviewers' comments
we realized that the original submission gave the impression that our main
result was the amount of attention
different models pay to words of different categories.  We hope that the
revised version helps the reader to also focus on
sections \ref{sec:beyondlexical}, \ref{sec:gramfunc} and
\ref{subsec:information-struct} which explore linguistic features
that the networks encode on top of simple lexical features. These and
the additional less substantial revisions are discussed in more
detail below.

\begin{verbatim}

>> Editor's Comments <<

The paper is pitched in very broad terms (methods for analyzing RNNs), 
but the methods are actually quite specific to classifier-type RNNs, 
and many aspects of the work seem applicable only to the Imaginet 
setting. I would suggest clarifying the scope of the claims. Similarly, 
as Reviewer A points out, some of the rhetoric is misaligned with the 
methods and results.  
\end{verbatim}

This point was raised by both \emph{Reviewer A} and \emph{Reviewer B} and 
had a major impact on our paper.
Sections \ref{sec:intro}, \ref{sec:computeomission}, 
\ref{sec:beyondlexical} and \ref{sec:reprdim} include new paragraphs
aimed to make it clear in which settings the described methods could be meaningfully applied.

\begin{itemize}
  \item In Section \ref{sec:intro} ``Introduction'' on page
    \pageref{generalintro}, we point out that we applied the introduced methods
   only to {\sc Imaginet}, 
  but that we also discuss their generalizability and offer pointers
  on how to apply them in other settings.
  \item On page \pageref{edit:introomissiongeneral} after briefly introducing the 
   \emph{omission score} technique we point out that it can be applied to
   other architectures too. The next paragraph on page
   \pageref{edit:introtopkgeneral}, after introducing the
   \emph{top-K-contexts} method itself, discusses its
  applicability to other settings. 
  \item We added a paragraph to the end of Section \ref{sec:computeomission} on
  page \pageref{edit:omitgeneral} that points out in a bit more detail 
  how the \emph{omission} technique can be generalized to other
  RNN variants, recursive neural networks and CNNs. 
  We hope that it will be clear to the reader that the analysis 
  provided in Section \ref{sec:omitimaginet} can be applied to the 
  same models as the \emph{omission} method.
  \item We also added text in the first paragraph of Section \ref{sec:beyondlexical}
  on page \pageref{edit:beyonlexicalgeneral} to clarify in which settings 
  the described methods could be meaningfully applied.
  \item In Section \ref{sec:reprdim} we explain that the provided 
  analysis can be straight-forwardly applied to other RNN variants and 
  suggest how to apply it to multi-layer architectures. We also point
  out that it is likely to be limited to uni-directional models.
\end{itemize}

\begin{verbatim}
The core results for content words and functions 
words seem like the expected ones given their 
linguistic content. In particular, the omission
score method seems to show that the networks learn 
to identify discriminating content words, and that
changing or removing those words has a dramatic effect 
on final predictions. This seems like a prerequisite 
(given the linguistic facts) for learning an effective classifier. 
Regular train/dev/test experimental analyses already
show that these models are effective. Could you further 
articulate what we learn via the omission score method 
that we don't get from standard evaluations? 
\end{verbatim}

As mentioned above, the fact that the {\sc Visual} network learns to
identify discriminating content words is only one of several results
in our paper. Even this basic finding is not as obvious as it
may seem: in the revised version, the last paragraph in Section \ref{sec:omitimaginet} on page 
\pageref{edit:generality} discusses that the relatively low omission scores
for verbs for the {\sc Visual} pathway are somewhat unexpected as one 
would expect high omission
scores for content words in general. We go on to discuss that the caption to
image prediction setting 
seems to lead to more task specific representations than expected.

Furthermore, much of the paper is devoted to investigating patterns
over and above attention paid to particular lexical items: for
example in our experiment in Section~\ref{sec:macro} we focus on the
effect of grammatical role and position in the sentence, {\it while
  controling for word-type.} (See also the comment below).

We believe these insights go well beyond what standard dev/test evaluations
can tell us.

\begin{verbatim}

The connections with POS tags and dependency relations are intriguing.
Could you, though, clarify the significance of these findings? My concern is
that they just reflect the fact that the content-word/function-word
distinction is highly correlated with POS and dependency relation.
\end{verbatim}

This comment really helped us restructure Section \ref{sec:macro}, with the aim
to describe and highlight our main findings in a more clear manner. 
Firstly in the new version of Section \ref{sec:omitimaginet} we do not emphasize so much
the content- vs. function-word distinction and point out the relatively
low omission of verbs in case of {\sc Visual}. Furthermore, in the new version
we dedicate Section \ref{sec:beyondlexical} to raise the question about 
the correlation between word-types, dependency relations and position in the 
sentence and describe how we disentangle these factors using linear
regression modeling. Concretely, we show
that the omission scores both in case of {\sc Visual} and {\sc Textual} do not
only depend on the word types, but on the grammatical function 
and position of the tokens, while controling for word types.
To further illustrate this point, we introduce a non-sequential baseline model {\sc Sum}, 
in which we replace the RNN with a
vector-summation operation over learned word-embeddings.

\begin{verbatim}
I had trouble understanding the information structure section. 
I agree with the core insight about old/topic before new/comment 
in languages like English. I had trouble relating this to what 
I would expect from the RNN. For instance, the structure of the 
RNN seems to ensure that information nearer the end of the sequence
will have a larger impact. Is the Textual omission score result due 
to this factor or is it truly due to something about language? 
Conversely, for Visual, I guess I would have anticipated exactly 
the opposite of what is reported here: the comment part should
be the most important, corresponding to the new and visually 
salient elements, rather than those that can fade into the background.
\end{verbatim}

Motivated by this and the previous comment in the revised version
sections \ref{sec:gramfunc} and \ref{subsec:information-struct} use the same methodology
for the analysis of the significance of grammatical function and position in the sentence.
We hope this modification makes our methodology and the interpretation of our 
results more straight-forward. Also for clarity, we streamlined the visualization
of comparing how the position of tokens affect the omission scores of the different models
- Figure \ref{fig:posrqs} on page \pageref{fig:posrqs}. 
The current plot shows the same effect for {\sc Textual}
and similar for {\sc Visual} as before, but hopefully in a more interpretable
way. Furthermore, we have added a short discussion 
on topic vs. comment in the image-description setting on page 
\pageref{edit:topiccomment}, which we agree was definitely needed.

\begin{verbatim}

>> Reviewer A <<

One question about the omission score: the lower the score is, the low the
similarity is. Thus, when dropping the important words, the similarity
should go down -- so a lower omission score. It is not clear to me why
dropping important words will lead to a higher omission score.

\end{verbatim}

Thank you for pointing this mistake out, we actually
compute the cosine distance $1-cos(x,y)$ not cosine.
We modified Equation \ref{eg:omit} on page \pageref{eg:omit} accordingly.

\begin{verbatim}
First, why is the architecture of
Chrupala et al (2015) selected only? Is it representative enough for general
RNNs? What about LSTM and its variants for language modelling task? How
would they be different? It would be good to see such discussions. 
\end{verbatim}

Section \ref{sec:intro} (``Introduction'')
discusses the reason why we choose this particular architecture
on page \pageref{explainimaginet}. The model contains 
two separate RNNs: a language model and a
model that predicts image-vectors given sentences. 
These ``pathways'' are trained on the same training data and 
share word-embeddings. We choose such a setup as we
are interested in 1) analyzing the hidden activation patterns 
(rather than word embeddings), and 2) developing methods that allow for comparing models.
As described in the first answer to the 
Editor's first question, not discussing the generality and
applicability of our methods was definitely a shortcoming of the paper and
in the new version sections \ref{sec:computeomission},
\ref{sec:beyondlexical}, \ref{sec:reprdim} discuss
how general our techniques are and how to apply them in new settings.

\begin{verbatim}
Second, most of the paper presents qualitative results, without rigorous
quantitative justifications. It would be nice, however, if the authors could
collect some manual annotations and conduct automatic evaluations based on
such annotations. For example, such annotations could include human's
judgements on how important each word (or word class, syntactic class) in
the sentence are etc.
\end{verbatim}
Regarding the collection of manual annotations, this is a great
suggestion and we also have been planning 
to assess the relationship between our results and human 
performance. Conducting such experiments with 
reliable results is labor- and resource-intensive and for the time 
being we consider such an experiment to be out of the scope of our current
paper, but as a great direction for our future work. 
We discuss this idea in the new version of Section 
\ref{conclusion} ``Conclusion''.

Regarding qualitative results, we believe the paper offers a good
balance of quantitative and qualitative analyses, even though our
quantitative results do not take the form of accuracy scores. Our
quantitative results include:
\begin{itemize}
 \item We introduce omission scores in Section \ref{sec:computeomission}
 with the aim to quantify the importance of words to 
 the final prediction.
 \item Section \ref{sec:omitimaginet} compares {\sc Visual} and {\sc Textual} 
 with respect to the difference between their omission score distributions
 over POS tags and deprel categories.
 \item The new Section \ref{sec:beyondlexical} fits linear models with 
 various sets of predictor variables to predict omission score as outcome and
 assesses the contribution of these variables. 
 The analysis in sections \ref{sec:gramfunc}, 
 \ref{subsec:information-struct} also utilize these models.
 \item In Section \ref{seq:describe} we compare 
 {\sc Visual} and {\sc Textual}
 by measuring the strength of association between the 
 binned activation values
 of the hidden units and the token/deprel n-grams in the corpus with mutual
 information. 
\end{itemize}

\begin{verbatim}
Finally, although the visualisation is helpful 
and it might help the readers understand the 
underlying activities of the RNN, little is known about why. 
I understand it is just an empirical paper, however, it would 
be helpful to at least share with the readers some high level 
insights so that the readers could learn more from this paper. 
One question that I have in mind after reading this paper is: 
these results related to RNN presented here are expected. 
It's good that RNN is able to capture such things. 
However, how do I make use of the results here in my research? 
How does this have some impact on my work?
\end{verbatim}

RNN variants proved to be powerful tools for modeling various aspects of human
language and for NLP applications, and yet there has been little work on showing 
what these models learn about language. In our paper we explore the 
learned representations from a linguistic point of view, in the spirit
of basic research.  
Engineering applications are not the focus of this paper, and so in
the body of the paper we try not to speculate on the relevance of our
findings to practical NLP.

That said, using our omission score results
we do observe that the task of predicting images 
from descriptions on MS-COCO does
not promote the learning of representations for verbs. 
Furthermore, one can observe a recency
effect in the meaning representation of {\sc Textual}, 
which probably also leads to too task specific representations. 
Thus, we do hope that our methods can provide help for such
engineering research in the future.

\begin{verbatim}
The authors claimed at the end of the paper that 
"The goal of our paper is to propose novel methods
for the encoding of linguistic knowledge in RNNs" -- I am not sure if I am
convinced that this is goal of this work.  It is largely an empirical paper
that tries to visualise/understand the RNN which was previously regarded as
a blackbox, but does not really "propose" methods for encoding linguistic
knowledge in RNN.
\end{verbatim}

This was an unfortunate mistake to leave this typo in the final version;
we meant to write ``The goal of our paper is to propose novel methods
for the ANALYSIS of the encoding of linguistic knowledge in RNNs''. We fixed it and 
thank you for noticing it!

\begin{verbatim}
 >> Reviewer B <<

The main criticism I would have of this paper is that it 
sets out as a providing generic new method for analysing RNNs 
with text input, amongst other modalities. 
I think this is overstating the contents of the paper: 
the methods of analysis are fairly tied to the nature of the model
(unidirectional RNN without attention) and broad task 
(producing a final representation that can be used for classification or generation). 
It would have been interesting to the same analyses performed on other, 
related models (bi-directional RNN, word-based log-bilinear models, etc), 
and on RNN tasks with other objectives (machine translation, language modelling, 
sequence labelling). I realise it's a tall order to add this to the current work,
but perhaps re-stating the scope of the work 
(providing an example of how to perform further qualitative 
and quantitative analysis of a an existing model trained
 on a specific task) 
 and briefly discussing how to adapt the experiments to other domains would be good.
\end{verbatim}

We really appreciated this comment! As described in our first answer, 
this observation had a large impact on our paper. We did add 
discussions in sections \ref{sec:intro}, \ref{sec:computeomission} 
,\ref{sec:beyondlexical}, \ref{sec:reprdim} on how to apply our 
introduced methods to different types of models.
We hope the Reviewer agrees that the changes we made based on 
this comment made our paper much stronger.

\begin{verbatim}
Likewise, a discussion comparing and contrasting these methods
of analysis with other (complementary?) visualisation methods 
for RNNs with attention such at Badhanau et al. 2015 or 
Rocktaschel et al. 2015 [ICLR 2016] (which could be cited here) would be welcome.
\end{verbatim}

This was also a great suggestion and we did include a 
discussion of the attention-weight 
visualization on page \pageref{edit:attention} in 
Section \ref{sec:related} based on the suggested works
``Related Work''. This has made the section much more complete.

\begin{verbatim}
It would be nice to get a clear statement of the objective being optimised
against in eq 9.
\end{verbatim}

The two loss functions are now described by equations \ref{eq:lossce}, and 
\ref{eq:losscos} and the multi-task loss by Equation \ref{eq:losscombo}.

\begin{verbatim}
Figure 2 a little hard to read.
\end{verbatim}

We were considering ways to make the metioned figure less cluttered, but
we are already reducing clutter by
only including labels with frequency at least 500. We feel increasing
this threshold further or merging categories would risk 
removing too much information and end up with a misleading picture. 
To help the understanding of the figure the main trends are explained in the text.

\begin{verbatim}
Figure 6 very hard to read.
\end{verbatim}

We agree that both the figure and the methodology in the previous
version concerning the models' sensitivity to position in
sentences was  not easy to interpret. Section 
\ref{subsec:information-struct} reports new results 
regarding the significance of the position of tokens along
with a less cluttered Figure \ref{fig:posrqs}.

\begin{verbatim}
Section 5.1 "textual is more sensitive to ..." -> it looks like textual is
more sensitive to most words being removed to some extent. I don't know if
it's fair to single out these categories. Instead, just contrast more
clearly with the heavy focus on content words (mainly NNs) for Visual.
\end{verbatim}

For clarity in the revised version of the paper we separated the
introduction of the omission scores and the results based on them. 
Section \ref{sec:computeomission} motivates omission scores and describes
how to compute them and Section \ref{sec:omitimaginet} describes the
distribution of the omission scores over POS-tags and deprels for 
{\sc Textual} and {\sc Visual}. On page \ref{edit:textualomission} 
in Section \ref{sec:omitimaginet} we point out that {\sc Visual} is 
only sensitive to a couple of word-types, 
while {\sc Textual} has a flatter omission score distribution in general.
Only after this observation we go on to single out a couple of
categories for a more detailed comparison.

\begin{verbatim}
Section 5.2, model 1 vs model 2 <- is the a significant difference in
number of features here? Model 2 seems like it would have more features, and
intuitively, more info == better model (modulo feature quality), so quick
comment on this here would be good.
\end{verbatim}

We re-structured the sections analyzing the impact of position 
and grammatical function. In the new version sections
\ref{sec:gramfunc} and \ref{subsec:information-struct} 
apply the same methodology to explore both factors. 
Instead of commenting  on the difference in the number of features, in the new
version for the comparison of {\sc Visual} and {\sc Textual} 
we introduce a non-sequential baseline model - {\sc Sum}, and apply
the same regression models with the same features to this baseline.
Adding position and deprel variables in the linear model
to predict omission scores does slightly increase the $R^2$ 
even in case of {\sc Sum} - probably
due to unseen words -, but much more so for {\sc Visual} and {\sc
  Textual}.

\begin{verbatim}
Section 5.3 is very tied to structure of the model use in this paper
(would expect different results for bidirection RNN, recursive neural net).
Comment on this here would be good.
\end{verbatim}

The first paragraph in section Section \ref{sec:beyondlexical}
on page \pageref{edit:beyonlexicalgeneral} argues that the analysis
presented regarding the significance of word position are tied to RNNs. 
We also added a footnote on page \pageref{edit:foot} 
pointing out that recursive neural networks
and CNNs are not affected by left-to-right and right-to-left word order
in the way RNNs and bi-RNNs are. We agree that the results 
would be probably different for bidirectional RNNs, but these 
architectures are still affected by word-order.

\begin{verbatim}
Not a fan of qualitative analysis as done in Table 1/2. Too easy
over-interpret, but I realise this is personal taste. Would not be offended
if this is left in, but would suggest scrapping this to create space for
discussions suggested above.
\end{verbatim}

Both in the Section \ref{sec:intro} and Section \ref{sec:reprdim}
we point out that the provided results are based on qualitative methods.
We agree that Tables \ref{tab:contexts}, \ref{tab:syntax} 
and also Figure \ref{fig:dimheat} are easy 
to over-interpret. This is arguably a danger with 
qualitative analysis in general, but we do believe that it is interesting
to complement our quantitative analysis with a detailed exploration
of the function of individual hidden units from a closer look.

\begin{verbatim}
 >> Reviewer C <<

The paper builds heavily on previous work "IMAGINET", is densely described
but lacking details in certain sections. For example, I find it rather
challenging to understand the following:

- the process of image vector prediction is vague (eqn 7). Does it mean
prediction at the pixel level of an entire image, or a blob or principal
image segment of interest? Also, how does the image retrieval shown in
Figure 1 operate?

\end{verbatim}

To make these points more clear:

\begin{itemize}
	\item On page \pageref{edit:dumdumeddy} we 
added a footnote ``Representing the full image, extracted from 
a pre-trained Convolutional Neural Network''.
	\item On page \pageref{edit:retrievalexplain} in Section 
	\ref{sec:computeomission} we added an explanation about the retrieval
	process.
\end{itemize}

\begin{verbatim}
- Figure 6 would benefit from showing the aggregated POS groupings such as
"VBN,VBG,VB" all grouped into "VERB" etc. for benefit of clarity in a small
graph (too cluttered)
\end{verbatim}

Thank you for pointing it out, we hope that the reviewer agrees that 
the new Figure \ref{fig:posrqs} is less cluttered and more intuitive.

\end{document}

### 2017-05-16 doc/categories.tex
paper files added

\section{Analysis of hidden activation vectors}
\label{sec:macro}

\begin{figure*}[t]
\hspace*{-0.5in}
\setlength{\tabcolsep}{0.01pt}
    \begin{tabular}{cccc}
    \includegraphics[scale=0.14]{new_omission_examples/textual1} &
    \includegraphics[scale=0.14]{new_omission_examples/visual1} &
    \includegraphics[scale=0.33]{new_omission_examples/img1} & 
    \includegraphics[scale=0.28]{new_omission_examples/img1_omit} \\
    \includegraphics[scale=0.16]{new_omission_examples/textual2} &
    \includegraphics[scale=0.16]{new_omission_examples/visual2} &
    \includegraphics[scale=0.28]{new_omission_examples/img2} & 
    \includegraphics[scale=0.30]{new_omission_examples/img2_omit}
    \end{tabular}
    \caption{The omission scores of {\sc Visual} and {\sc Textual} for two example sentences, 
and the best retrieved images for the original sentence (left) and the sentence with the most 
important word removed (right). 
%Top row example caption is {\it a bottle next to a windowsill with
%  light coming through},  bottom row: {\it a train on a track next to
%  many bushes}.
}
    \label{fig:omissionex}
\end{figure*}

%Understanding the behavior of the learned models in terms of which features do they assign 
%the most attention to is straight forward with linear 
%models such as linear/logistic regression, 
%naive-bayes classifiers or with some of the non-parametric models as well such as decision 
%trees. This information allows researchers to not only train learning algorithms to perform 
%particular tasks, but also to uncover some underlying patterns in the training data.

%\subsection{Salience of categories of words}
%\label{sec:salience}
In this section we propose novel techniques for interpreting the
activation patterns of neural networks trained on language tasks
from a linguistic point of view and focus
on the high level understanding of what types of words the networks
pay most attention to. Furthermore, we investigate if the networks
learn to assign appropriate amounts of importance to tokens depending
on their position and grammatical function in the sentences.\label{edit:whyposdep}
In Section \ref{sec:computeomission} we introduce \emph{omission scores};
a metric to measure the contribution of each token to the prediction of the networks.
Section \ref{sec:omitimaginet} aggregates the omission scores in terms
of dependency relations and part-of-speech categories and compares {\sc Visual}
and {\sc Textual}. Lastly Section \ref{sec:beyondlexical} investigates the extent to which
the importance of words for the different pathways depend on the words themselves,
their position and their grammatical function in the sentences.
% MAYBE THIS WOULD BE TOO MUCH
%We apply the proposed methods to explore
%the representations of {\sc Imaginet}, but also discuss how to generalize
%them to other architectures.

\subsection{Computing Omission Scores}
\label{sec:computeomission}

In both pathways of {\sc Imaginet} the full sentences are represented by the
activation vector at the end-of-sentence symbol
($\mathbf{h}_\text{end}$). We measure the salience of each word $S_i$
in an input sentence $S_{1:n}$ based on how much the representation of the
partial sentence $S_{\setminus i} = S_{1:i-1}S_{i+1:n}$, with
the omitted word $S_i$, deviates from that of the original sentence
representation. For example, the distance
between $\mathbf{h}_\text{end}(${\it the black dog is running}$)$
and $\mathbf{h}_\text{end}(${\it the dog is running}$)$ determines
the importance of {\it black} in the first sentence. We introduce the
measure $\mathrm{omission}(i,S)$ for estimating the salience of a word $S_i$:

\begin{equation}
\label{eg:omit}
\mathrm{omission}(i,S) = 1-\mathrm{cosine}(\mathbf{h}_\text{end}(S),
\mathbf{h}_\text{end}(S_{\setminus i}))
\end{equation}

\noindent Figure~\ref{fig:omissionex} demonstrates the omission
measure for the {\sc Visual} and {\sc Textual} pathways for two
example captions. For both captions the $\mathrm{omission}$ scores are
plotted along with the first image retrieved by {\sc Visual} for the
full sentence and for the sentence with the word with the highest
$\mathrm{omission}(i,S)$ removed. The images are retrieved from the
validation set of MS-COCO by: 1) computing the image representation of the 
given sentence with {\sc Visual}; 2) extracting the CNN features for the 
images from the set; and 3) finding the image that minimizes the cosine distance
to the query.\label{edit:retrievalexplain} 
In the first example, the omission
scores for {\sc Visual} suggest that the model interpreted {\it  pizza} as the most important
word in the sequence and returned an image that depicts a pizza on a plate. 
Removing the word {\it pizza} promotes {\it salad} and {\it bowl} as
the main theme of the sentence and the model retrieves an image with a dining table 
with salad-like dishes. 
For the second example, the omission scores for {\sc
  Visual} show that the model paid attention mostly to {\it baby} and
{\it bed} and slightly to {\it laptop} and retrieved an image depicting a baby 
sitting on a bed with a laptop.
Removing the word {\it baby} leads
to an image that depicts an adult male laying on a bed. Figure~\ref{fig:omissionex}
also shows that in contrast to {\sc Visual}, in both examples {\sc
  Textual} distributes its attention more evenly across time steps
instead of focusing on the types of words related to the corresponding
visual scene.

For other RNN architectures such as GRUs, LSTMs \label{edit:omitgeneral}
and their bi-directional variants to measure the contribution
of tokens to their predictions the $\mathrm{omission}$ 
can be straight-forwardly computed using their hidden state 
at the last time step used for prediction. Furthermore, the technique 
can be applied in general to other architectures, which
embed variable length linguistic expressions to the same fixed dimensional
space and perform predictions based on these embeddings. 
This includes tree-structured Recursive Neural Network models such as the Tree-LSTM
introduced in \namecite{kai2015treelstm} or the CNN architecture of \namecite{yoonneural2014} 
for sentence classification. In both cases the pre-softmax activations can be extracted 
from the models as the representations of the full and partial sentences.

% Given a corpus segmented into sentences, each word in an input sentence is tagged with its 
%part-of-speech category (POS) and dependency-relation (DepRel). For dependency relations 
%only the label of the ark (and not the ark itself) is taken into account. For each input sentence 
%of 
%length $n$ the RNNs produce $n$ hidden activation vectors $h_{1}, \ldots , h_{n}$. Each word 
%in the input sentence is tagged with POS and DepRel categories and the contribution of the $<
%$word, POS, DepRel$>$ tuples to the meaning of the sentence is measured by computing the 
%following scores:
%

%\paragraph{$omission$} Given a sentence "the black dog", we generate 3 sentence by 
%omitting one of the words each time: "the black", "the dog", "black dog". For each of these 
%sentence we compute a sentence representation $h_{full}$ and calculate the cosine 
%similarities 
%between the original and the omitted sentences. This measured is used to indicate the overall 
%importance of word categories. \todo{Example with a short sentence: ab, bc, cd, cosines}
%\\

%-----------------
\subsection{Omission score distributions}
\label{sec:omitimaginet}

%\begin{figure*}
 %   \includegraphics[scale=0.6]{omission-stat/sent-omission-pos-boxplot.png} 
 %   \includegraphics[scale=0.6]{omission-stat/sent-omission-deprel-boxplot.png} 
 %   \caption{Distributions of omission scores per POS category (left)
 %     and dependency relation (right) for {\sc Sentiment}. Only labels
 %   which occurred at least 300 times are included.}
%\label{fig:omission-sent}
%\end{figure*}

%\subsubsection{Omission results for {\sc Sentiment}}
%\label{sec:omitsentiment}
%Figure~\ref{fig:omission-sent} shows the distribution of the omission scores per category for the {\sc Sentiment} model. Superlative adjectives (JJS) such as {\it best, least, scaries} and {\it funniest} are by far the most influential POS category, followed by Adjectives (JJ), Verbs (VBZ, VBG), Nouns (NN, NNS) and Adverbs (RB). Similarly, tokens with function {\sc root} (various verbs, nouns or adjectives central to the meaning of the sentence) are recognized by the model as the most important category, followed by adjectival modifier {\sc amod}. These results support the validity of $\mathrm{omission}$ as a measure of saliency and provide evidence that {\sc Sentiment} learns representations that enable the model to successfully filter out irrelevant parts of the input.

%-----------------
\label{subsec:omission-text-vis}
\begin{figure*}[t]
%\hspace*{-0.3in}
\setlength{\tabcolsep}{0pt}
  \begin{tabular}{cc}
  \includegraphics[scale=0.55]{omission-stat/imaginet-omission-pos-boxplot.png} &
  \includegraphics[scale=0.55]{omission-stat/imaginet-omission-deprel-boxplot.png} \\
  \end{tabular}

\caption{Distribution of omission scores for POS (left) and deprel labels
  (right), for the {\sc Textual} and {\sc Visual} pathways. Only labels which occur at least 500 times are included.}
\label{fig:omission-imaginet}
\end{figure*}

\begin{figure*}[t]
  \centering
  \hspace*{-0.2in}
  \setlength{\tabcolsep}{0pt}
  \begin{tabular}{cc}
  \includegraphics[scale=0.55]{omission-stat/omission-pos-boxplot.png} &
  \includegraphics[scale=0.55]{omission-stat/omission-deprel-boxplot.png} \\  
  \end{tabular}
  \caption{Distributions of log ratios of omission scores of {\sc Textual} to {\sc Visual} per
    POS (left) and deprel labels (right). Only labels which occur at least 500 times are included.}
\label{fig:omission-imaginet-ratio}
\end{figure*}

The $\mathrm{omission}$ scores can be used not only to
estimate the importance of individual words, but also of syntactic
categories. We estimate the salience of each syntactic category by
accumulating the omission scores for all words in that category. We
tag every word in a sentence with the part-of-speech (POS) category
and the dependency relation (deprel) label of its incoming arc. For
example, for the sentence \emph{the black dog}, we get ({\it
  the},~DT,~det), 
({\it black},~JJ,~amod), ({\it dog},~NN,~root). 
Both POS tagging and dependency parsing are performed jointly
using the TurboParser dependency parser \cite{martins2013turning}.\footnote{Available at
  \url{https://github.com/andre-martins/TurboParser}{github.com/andre-martins/TurboParser}.} The POS tags used are the Penn Treebank tags and the dependencies are 
the Stanford basic dependencies.

Figure~\ref{fig:omission-imaginet} shows the distribution of omission
scores per POS and deprel label for the two pathways of {\sc Imaginet}. 
The general trend is that for the {\sc Visual} pathway, the omission scores are
high for a small subset of labels - corresponding mostly to nouns, less so for
adjectives and even less for verbs - and low for the rest (mostly function words and
various types of verbs). For {\sc Textual} the differences are \label{edit:textualomission}
smaller, and the pathway seems to be sensitive to the omission of most
types of words.

Figure~\ref{fig:omission-imaginet-ratio} compares the
two pathways directly using the log of the ratio of the {\sc Visual}
to {\sc Textual} omission scores, and plots the distribution of this
ratio for different POS and deprel labels.  Log ratios above zero
indicate stronger association with the {\sc Visual} pathway and below
zero with the {\sc Textual} pathway. We see that in relative terms,
{\sc Visual} is more sensitive to nouns (NNP, NNS, NN), numerals (CD)
and adjectives (JJ), and {\sc Textual} to the prepositions (TO, IN),
some types of verbs (VBZ, VBP), determiners (DET, WDT) and
particles (RP). This picture is complemented by the analysis of the
relative importance of dependency relations: {\sc Visual} pays most
attention to the relations {\sc nsubjpass, nsubj, pobj, root, dobj,
  nn, conj, dep, amod} and {\sc num}, whereas {\sc Textual} is more
sensitive to {\sc det, mark, aux, auxpass, prt, prep, poss, cc} and
{\sc cop}.

As expected, {\sc Visual} is more focused on grammatical
functions typically filled by semantically contentful words, while
{\sc Textual} distributes its attention more uniformly and 
attends relatively more to purely grammatical functions. 
It is worth noting, however, the relatively low omission scores for verbs in case \label{edit:generality}
of {\sc Visual}. One might expect that the task of image prediction from
descriptions requires general language understanding and so high omission
scores for all content words in general, however, the results
suggest that this setting is not optimal for learning useful representations of verbs,
which possibly leads to representations that are too task specific 
and not transferable across tasks.

%The results on syntactic categories for {\sc Visual} are somewhat similar to 
%{\sc Sentiment} in that both models seem to learn to focus on one
%particular category (Nouns for {\sc Visual}, Adjectives for {\sc Sentiment}). 
%{\sc Visual} still incorporates adjectives, personal
%pronouns and verbs to the representations, but tends to ignore
%words of other categories.

\subsection{Beyond Lexical Cues}
\label{sec:beyondlexical}

Models that utilize the sequential structure of natural-language 
have the capacity to interpret the same word-type differently depending on
the context. The omission score distributions in Section \ref{sec:omitimaginet} 
show that in the case of {\sc Imaginet} the 
pathways are differentially sensitive to content vs.\ function
words. This may be either just due to purely lexical features or the model 
may actually learn to pay more attention to the same word type in appropriate
contexts. This section investigates to what extent {\sc Visual} and {\sc Textual}
discriminate between occurrences of a given word in different positions and 
grammatical functions. The analysis we described here takes \label{edit:beyonlexicalgeneral}
the omission scores as input data, therefore it can be potentially applied 
to any architecture for which the omission scores can be computed. However,
the presented analysis and results regarding word positions can only be meaningful
for Recurrent Neural Networks as they compute their representations sequentially and are not
limited by fixed windowsizes.\footnote{CNNs with multi-word filters
and tree-structured recursive neural networks do not incrementaly build representations
of sentences in a left-to-right or right-to-left fashion. 
Bi-directional RNNs, however, are affected by word-order and can potentially
learn to handle the same word in different positions differently. \label{edit:foot}}

We fit four linear models which predict the omission 
scores per token with the following predictor variables:

\begin{enumerate}
	\item {\sc LM word}: word type
	\item {\sc LM deprel}: word type, dependency label and their interaction 
	\item {\sc LM position}: word type, position (binned as first, second, third, middle,
	antepenult, penult, last) and their interaction
	\item {\sc LM full}: word type, dependency label, position, word-deprel interaction, 
	word-position interaction
\end{enumerate}

\noindent We split the whole MSCOCO validation set into two parts, fit the 
models (regularized via L2 penalty) on the first part, and compute the 
proportion of variance explained on the second part. 
For comparison we also train a version of {\sc Imaginet},
where the GRU in the {\sc Visual} pathway is replaced by a simple vector-summation operation
over word-embeddings and thus does not have access to sequential cues. This
model we refer to as {\sc Sum}.

Figure~\ref{fig:rsquared} shows the increase in $R^2$  for the models 
per pathway relative to {\sc LM word}. The plot reveals that 
adding additional information on top of lexical features in case of {\sc Sum} 
does increase the explained variance slightly, which is probably due to 
the unseen words in the held out set. More interestingly, the
relative increase in $R^2$ between {\sc LM word} and {\sc LM full} in
case of {\sc Visual} is larger, suggesting that 
the omission scores in case of {\sc Visual} do not only depend on the words themselves,
but their grammatical function and position in sentences.  The raw  $R^2$
scores show that for the {\sc Textual} model the word-type predicts the omission-score
to much smaller degree compared to {\sc Visual} (0.38 against 0.48), and the 
increase in $R^2$ shows that the position of the word adds considerable 
amount of information.

\begin{figure}
\centering
  \includegraphics[scale=0.35]{omission-stat/position-new.png}
\caption{Proportion of variance in omission scores explained by the models
 for {\sc Sum}, {\sc Visual} and {\sc Textual} pathways.}
\label{fig:rsquared}
\end{figure}

\subsubsection{Sensitivity to grammatical function}
\label{sec:gramfunc}

%\begin{figure}[t]
%  \centering
%  \includegraphics[scale=0.4]{omission-stat/scores-vs-alpha.pdf}
%  \caption{Proportion of variance in omission scores explained by {\sc Model 1} vs {\sc Model 2} as a function of regularization parameter $\alpha$.}
%  \label{fig:scores-vs-alpha}
%\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.35]{omission-stat/top_words.pdf}
  \caption{Distribution of omission scores per deprel label for the selected word types.}
  \label{fig:top_words}
\end{figure}
%\todo{At some point redo this with words ranked by ratio between the errors not the difference}

In order to find out some of the specific configurations leading to
increase in $R^2$ in case of {\sc Visual}, we next considered all word types with
occurrence counts of at least 100 and ranked them according to how much
better, on average, {\sc LM deprel} predicted their omission scores 
compared to {\sc LM word}. Figure~\ref{fig:top_words} shows the per-dependency 
omission-score distributions for the five top ranked words plus the word {\it  water}. 
There are clear and large differences in how these words
impact the network's representation depending on what grammatical
function they fulfill. They all have large omission scores when they
occur as {\sc nsubj} (nominal subject) or {\sc root}, likely due to the fact that these
grammatical functions typically have a large contribution to the
complete meaning of a sentence.  Conversely, all have small omission 
scores when appearing as {\sc conj} (conjunct): this is probably because in this position
they share their contribution with the first, often more important,
member of the conjunction (e.g. {\it A cow and its baby eating
  grass}).

The pattern for {\sc nn} (nominal modifier) is a bit more complicated: for
four of the words shown (as well as for most other words not shown in
the figure), the score is very low in this grammatical function--presumably 
because most words contribute less to the sentence meaning when used  
as modifiers than as heads (e.g.\ {\it a clock tower}). However,
for the words {\it zebra} and {\it water}, omission scores are high
when they act as a nominal modifier {\sc nn}. This appears due to two reasons:
\begin{enumerate}
\item For {\it zebra}, there are frequent erroneous parses such as {\it
    zebra/{\sc nn} browsing/{\sc root}} instead of {\it zebra/{\sc nsubj} browsing/{\sc
    root}}. The network does not make this mistake, and treats these
    occurrences of {\it zebra} according to its importance as {\sc nsubj}.
\item {\it Water} as a modifier often changes the meaning of its head in
 a visually salient way: e.g.\ {\it water fall, water balloon, water scene, water skiing}, and
  thus the network learns that this particular word is important in the modifier position.
\end{enumerate}

\subsubsection{Sensitivity to information structure}
\label{subsec:information-struct}

\begin{figure}
\centering
 \includegraphics[scale=0.4]{omission-stat/position-coef.png}
 \caption{Coefficients on the y-axis of {\sc LM full} corresponding to the
position variables on the x-axis.}
 \label{fig:posrqs}
\end{figure}

As observed in Section \ref{sec:beyondlexical}, 
adding extra information about the position of words 
explains more of the variance in case of {\sc Visual} and especially {\sc Textual}.
Figure \ref{fig:posrqs} shows the coefficients corresponding to the
position variables in {\sc LM full}. Since the omission scores 
are measured at the end-of-sentence token, the expectation is that 
for {\sc Textual}, as a language model,  
the words appearing closer to the end of the sentence would have a
stronger effect on the omission scores. This seems to be confirmed by
the plot as the coefficients up until the \emph{penultimate} are all negative.

For the {\sc Visual} model it is less clear what to
expect: on the one hand due to their chain structure,  
RNNs are better at keeping track of
short-distance rather than long-distance dependencies and thus we can expect 
tokens in positions closer to the end of the sentence to be more important. 
On the other hand in English the information structure (or pragmatic structure) 
of a sentence is expressed via linear ordering: the {\sc topic} of a 
sentence appears sentence-initially, and the {\sc comment} follows. 
In the context of other text types such as dialog or multi-sentence \label{edit:topiccomment}
narrative structure, we would expect {\sc comment} to often be more
important than {\sc topic} as {\sc comment} will often 
contain new information in these cases. In our setting of image captions 
however, it is the {\sc topic} that typically contains the most important 
objects depicted in the image, e.g. {\it {\underline{two zebras}} are grazing in tall grass on a savannah.}
Thus, for the task of predicting features of the visual scene, it would 
be advantageous to detect the topic of the sentence and up-weight its
importance in the final meaning representation. Figure \ref{fig:posrqs}
appears to support this hypothesis and the network does 
learn to pay more attention to  words appearing
sentence-initially. This effect seems to be to some extent mixed with the recency
bias of RNNs as perhaps indicated by the relatively high coefficient of the {\it last}
position for {\sc Visual}.

% The basic idea is to focus on
% words which occur in several grammatical functions, and see whether
% the network treats the different occurrences differently based on the
% dependency label.

% The omission score results for {\sc Visual}
% suggest that the model pays relatively more attention to 
% categories {\sc nsubj, pobj, root} than to {\sc nn}. 
% Since the same word types can potentially
% assume these categories in different contexts which leads to the
% hypothesis that {\sc Visual} learned to recognize small syntactic 
% constructions. In the following experiment we consider word types 
% that appear with all four of the following dependency labels: 
% {\sc nsubj, pobj, root, nn}. Table \ref{tab:nnexample} demonstrates
% examples of the systematic pattern that in when tokens with deprel {\sc nn}
% have the lowest

% A systematic pattern that provides evidence to support
% the hypothesis is that when nouns assume the role of noun complement
% {\sc nn}, their omission score is the lowest. This result is illustrated
% in

% \iffalse
% \subsection{Distribution of omission scores}

% Determining feature importances from RNN hidden states for NLP tasks
% allows to uncover what kinds linguistic features models focus on.  We
% showed that omission scores allow for interesting qualitative
% comparison between RNN based models, but also provides insight into
% the nature of the combinations of data sets and tasks. The omission
% score results shows that to estimate the polarity of the sentence a
% model should primarily focus on adjectives. More interestingly the
% relative higher omission scores for Nouns for {\sc Visual} suggests
% that combination of the MS-COCO data set and the pre-trained VGG-CNN
% image-representation prediction as objective does not seem to promote
% the model to learn to pay attention to verbs. This analysis suggests
% that this particular setting does not allow the model to learn general
% representations for phrases/sentences as it promotes filtering out the
% information content of verbs.  Relative to {\sc Visual} the
% omission-score distribution of the {\sc Skip-gram} is much flatter
% providing an indication that the model does not throw away as much
% important linguistic information.  \todo{show some plot of the
%   distributions VISUAL vs. skip-gram} \fi

### 2017-05-16 doc/conclusions.tex
paper files added

\section{Conclusion}
\label{conclusion}

The goal of our paper is to propose novel methods for the 
analysis of the encoding of linguistic knowledge in RNNs trained on language tasks.
We focused on developing quantitative methods to measure the importance
of different kinds of words for the performance of such models. Furthermore, we
proposed techniques to explore what kinds of linguistic features the models
learn to exploit on top of lexical cues.

Using the {\sc Imaginet} model as our case study,
our analyses of the hidden activation patterns show that the {\sc Visual} model
learns an abstract representation of the information structure of the
language, and pays selective attention to lexical categories and
grammatical functions that carry semantic information. In contrast,
the language model {\sc Textual} is sensitive to features of a more
syntactic nature. We have also shown that each network contains
specialized units which are tuned to both lexical and structural
patterns that are useful for the task at hand, some of which can carry
activations to later time steps to encode long-term dependencies.

In future we would like to apply the techniques introduced in this paper 
to analyze the encoding of linguistic form and function of 
recurrent neural models trained on different objectives, 
such as neural machine translation systems
\cite{sutskever2014sequence} or the purely distributional
sentence embedding system of \cite{kiros2015skip}. A number of
recurrent neural models rely on a so called attention mechanism, first
introduced by \namecite{bahdanau2014neural} under the name of soft
alignment. In these networks attention is explicitly represented and it
would be interesting to see how our method of discovering implicit
attention, the omission score, compares. For future work we also propose to
collect data where humans assess the importance of each word in sentences
and explore the relationsship between \emph{omission scores} for various models
and human annotations.\label{edit:humanjudgement} Finally, one of the benefits
of understanding how linguistic form and function is represented in RNNs is that
it can provide insight into how to improve systems. We plan to draw on
lessons learned from our analyses in order to develop models with better 
general-purpose sentence representations.

% \iffalse
% We developed \emph{macro} and \emph{micro} level methods to analyze the 
% activation patterns of Recurrent Neural Networks
% from a linguistic point of view. On the \emph{macro} level
% we introduced the $\mathrm{omission}$ score
% to measure the salience of tokens in sentences and
% showed how aggregating these scores in terms of part-of-speech categories and
% grammatical functions allows for a more in-depth understanding of the kinds of
% linguistic structure RNNs learn from linguistic data. In particular we have shown
% that {\sc Visual} learns to pay attention to tokens depending of their grammatical
% function. In these experiments we focused on how the model interprets
% the same \emph{nouns} fulfilling different grammatical functions, but in future work 
% this method can be straight-forwardly applied to other classes of content-words 
% such as \emph{verbs} or \emph{adjectives}. In addition we provided evidence 
% that {\sc Visual} pays more attention to nouns that are closer to 
% the beginning of the sentence, which is motivated by the information structure of English.
% In our \emph{micro} level analyses we developed a simple and general purpose method
% dubbed \emph{top K contexts}, which describes the function of a particular hidden unit 
% by ranking the contexts that produce the highest activation values for that unit.
% We performed exploratory analysis on these contexts, especially focusing on 
% hidden units whose activations are predictive of the grammatical function of tokens.
% We observed that in numerous cases units are highly activated for contexts that represent
% combined syntactic semantic templates. Furthermore, we explored dimensions that
% carry over their activations to represent longer term dependencies and demonstrated 
% a visualization technique to explore these patterns. Lastly, we performed a comparison
% between {\sc Textual} and {\sc Visual} using the mutual information between the activation
% values of their hidden units and the contexts. Our analysis showed a high level difference
% between the kinds of features the two pathways extract, in that the
% linguistic regularities encoded by hidden units of {\sc Textual} 
% seem to be more characterized by syntactic templates than in case of {\sc Visual}.
% In addition to the insights provided by the methods implemented in this paper
% we believe them to be general enough to allow  cognitive linguistics research 
% to further explore the linguistic knowledge in the activation patterns
% of RNNs trained on large scale data sets and on challenging tasks.

% \fi

### 2017-05-16 doc/experiments.tex
paper files added

\section{Experiments}
\label{sec:experiments}

In the following sections, we report a series of experiments in which
we explore the kinds of linguistic regularities the
{\sc Textual} and {\sc Visual} pathways of {\sc Imaginet} learn from
word-level input. Section~\ref{sec:macro} presents the macro-level
analyses where we propose methods to analyze the final hidden activation 
vectors of the recurrent pathways from linguistic point of view. 
Section~\ref{sec:micro} reports exploratory experiments on
the linguistic features encoded by individual hidden units. 
In both sections we report our findings based on the {\sc Imaginet} model
and discuss the generalizabilty of our methods to other architectures.
\label{edit:experimentsgeneral}
For all the experiments, we trained {\sc Imaginet} on the training portion of the
MSCOCO image-caption dataset \cite{lin2014microsoft}, and analyzed the
representations of the sentences in the validation set. The target
image representations were extracted from the pre-softmax layer of the
16-layer CNN of \namecite{simonyan2014very}.

% \iffalse

% {\bf Section~\ref{sec:macro}} includes experiments that provide linguistic analysis of the activation patterns on the activation vector level with results more quantitative in nature. Section~\ref{sec:salience} describes a method to estimate the salience of tokens as a function of their part-of-speech category and grammatical function (dependency relation) in sentences.

%  Sections~\ref{sec:gramfunc} and \ref{subsec:information-struct} apply this method and provide evidence that the {\sc Visual} pathway of {\sc Imaginet} learns to interpret the same word fullfilling different grammatical functions differently, and is sensitive to the information structure of sentences.

% {\bf Section~\ref{sec:micro}} compliments the macro level analysis and reports a series of more qualitative experiments exploring the kinds of linguistic abstraction represented by individual hidden units. Section~\ref{sec:reprdim} proposes a simple method for representing
% the function of hidden units by identifying the top K contexts that yield the highest activation values, and Section~\ref{sec:topk} provides a general qualitative analysis of these contexts.

% For the experiments using the {\sc Imaginet} model we use the validation set of MS-COCO.
% Finally Section~\ref{sec:syntacticdim} goes more into detail and shows examples 
% of hidden units that are active for particular multi-word constructions with joint
% semantical and syntactic regularities.  
% \fi

%\begin{enumerate}
%\item How accurately does each model represent syntactic knowledge? WE DONT DO THIS
%\item What types of grammatical functions each model pays most attention to? WE DO THIS
%\item What is the level of separation between syntactic and semantic 
%representations learned by the models? NO CHANCE
%\item What functions are the individual units in each network specialized for? WE DO THIS 
%\item On each of the above dimensions, what are the systematic differences 
%between models that are optimized for different tasks?  WE ALSO DO THIS
%\end{enumerate}

%Section~\ref{sec:categories} looks at the salience of two sets of syntactic categories in 
%each model, by estimating how much input tokens from each category influence 
%the final performance in each task. In Section~\ref{sec:syntax}, we examine the 
%degree to which syntactic information is encoded in the internal representations 
%acquired by each model. Section~\ref{sec:dimension} focuses on analyzing the 
%reaction patterns of individual units in each model, and on finding out specialized 
%functions that these units perform.

%Finally in Section~\ref{sec:visualize}, we look at 
%how the meaning representations of word sequences evolve over time for each 
%model. \todo{not sure how to relate this last one to our objectives.}
%In all our experiments, we compare the nature and degree of the acquired 
%syntactic knowledge among models that are trained for performing different tasks.

%\noindent{\bf Syntactic categories.} In all the following analyses, we use two different sets of syntactic categories: part-of-speech categories (POS) and dependency relations (DepRel). The tagging is performed jointly using the TurboParser dependency parser\footnote{This implementation is available at https://github.com/andre-martins/TurboParser.} \cite{martins2013turning}. For POS tags we use the Penn Treebank tagset and for the dependency relations the Stanford basic dependencies.

%\input{regression}
\input{categories}
\input{dimensions}
%\input{visualization}

### 2017-05-16 doc/intro.tex
paper files added

\begin{abstract}

  We present novel methods for analyzing the activation patterns of
  RNNs from a linguistic point of view and explore the types of 
  linguistic structure they learn. As a case study, we use a multi-task gated
  recurrent network architecture consisting of two parallel pathways with
  shared word embeddings trained on predicting the
  representations of the visual scene corresponding to an input
  sentence, and predicting the next word in the same sentence. 
  Based on our proposed method to estimate the amount of contribution
  of individual tokens in the input to the final prediction of the networks   
  we show that the image prediction pathway: a) is sensitive to
  the information structure of the sentence, b) pays
  selective attention to lexical categories and grammatical functions
  that carry semantic information, and c) learns to treat the same
  input token differently depending on its grammatical functions in
  the sentence. In contrast the language model is comparatively more
  sensitive to words with a syntactic function. 
  Furthermore, we propose methods to explore the function
  of individual hidden units in RNNs and show that the two pathways
  of the architecture in our case study contain
  specialized units tuned to patterns informative
  for the task, some of which can carry activations to later time
  steps to encode long-term dependencies.

%Recurrent Neural Network (RNN) models with word embeddings learn
%representations of variable-length linguistic units from text. 
%In the fact that there are no obvious cues to the
%kinds of linguistic information these representations encode lies a vast 
%new area of research questions. The current paper develops general
%methods for the analysis of the learned representations with a focus on
%linguistic \emph{structure} and using \emph{comparative} experiments on two RNNs
%trained on the same data set with different objectives.
%The experiments in the paper consider \emph{macro} and
%\emph{micro} level analysis on the hidden activations of RNNs. On the 
%\emph{macro} level we develop a novel method to measure the saliency of 
%tokens in sentences and show how aggregating 
%these scores in terms of part-of-speech categories
%and dependency relations provides deeper insights into what the hidden
%activations of RNNs learn about linguistic structure. On the \emph{micro}
%level we develop a simple general method to describe
%the functionality of individual hidden units. This technique allows us for the
%exploration of the lingusitic regularities hidden units learn to encode. In
%addition we develop a visualization technique to trace the activations of individual
%units as they carry over their information through multiple time steps. Both on 
%\emph{macro}
%and \emph{micro} level we explore the differences between the examined models
%
%We provide evidence that {\sc Visual}
%learns to treat the same tokens differently depending on their grammatical function.
%In addition we show that it learns to focus on the first Noun in the sentences exploting
%the linear information structure of English.

\end{abstract}

\section{Introduction}
\label{sec:intro}
Recurrent neural networks (RNNs) were introduced by
%\cite{elman1990finding} 
as a connectionist architecture with the
ability to model the temporal dimension. They have proved popular for
modeling language data as they learn representations of words and
larger linguistic units directly from the input data, without feature
engineering. Variations of the RNN architectures have been applied in
several NLP domains such as parsing \cite{vinyals2015grammar} and
machine translation \cite{bahdanau2014neural}, as well as in computer
vision applications such as image generation \cite{gregor2015draw} and
object segmentation \cite{visin2015reseg}. RNNs are also important
components of systems integrating Vision and Language, e.g.\ image
\cite{karpathy2015deep} and video captioning \cite{yu2015video}.

These networks can represent variable-length linguistic expressions by
encoding them into a fixed-size low-dimensional vector. The nature and
the role of the components of these representations are not directly
interpretable as they are a complex, non-linear function of the
input. There have recently been numerous efforts to visualize deep
models such as convolutional neural networks in the domain of computer
vision, but much less so for variants of RNNs and for language
processing.

The present paper develops novel methods for uncovering abstract
linguistic knowledge encoded by the distributed representations of RNNs,
with a specific focus on analyzing the hidden activation patterns rather 
than word embeddings and on the syntactic generalizations 
that models learn to capture. In the current work we apply our methods
to a specific architecture trained on specific tasks, but also provide
pointers about how to generalize the proposed analysis to other settings.\label{generalintro}

As our case study we picked the {\sc Imaginet} model introduced by \label{explainimaginet}
\cite{chrupala2015learning}. It is a multi-task, multi-modal
architecture consisting of two Gated-Recurrent Unit (GRU)
\cite{cho2014properties,chung2014empirical} pathways and
a shared word embedding matrix. One of the GRUs ({\sc Visual}) is
trained to predict image vectors given image descriptions, while the other
pathway ({\sc Textual}) is a language model. This particular
architecture allows a comparative analysis of the hidden activations
patterns between networks trained on two very different tasks, while
keeping the training data and the word embeddings fixed. Recurrent neural
language models akin to {\sc Textual} which are trained to predict the
next symbol in a sequence are relatively well understood, and there
have been some attempts to analyze their internal states e.g.
\cite{elman1991distributed,karpathy2015visualizing}. In
constrast, {\sc Visual} maps a complete sequence of words to
a representation of a corresponding visual scene and is a less
commonly encountered, but a more interesting model from the point of
view of representing meaning conveyed via linguistic structure.

We report a thorough \emph{macro level} quantitative analysis to
provide a linguistic interpretation of the networks' activation
patterns. These experiments rely on a novel method we
call \emph{omission score} to measure the importance of input tokens 
to the final prediction of models that compute distributed representations 
of sentences. This includes various RNN architectures, Recursive Neural Networks 
and Convolutional Neural Networks.\label{edit:introomissiongeneral}
Based on these omission scores our experiments show that the image prediction 
pathway in general pays special attention to words of syntactic categories that 
carry semantic content. Furthermore, we observe that on top of
lexical features it learns important aspects of the information structure of sentences,
and treats the same input tokens differently depending on their
grammatical function in the sentence. In contrast, the language model
is more sensitive to the local syntactic characteristics of the input
sentences.

%On the macro level, we 
%develop a novel method to estimate the salience of tokens as a function of their syntactic 
%category and grammatical function. We show that due to the nature of its objective, the 
%{\sc Visual} pathway of {\sc Imaginet} learns important information structural properties of 
%language, and treats the same tokens differently depending on their grammatical functions 
%in the sentence. \todo{What can we say about Textual?}

We also perform a \emph{micro level} analysis to explore \label{edit:introtopkgeneral}
the function of individual hidden units by introducing a method we dub \emph{top-k-contexts}.
It involves identifying the sentential contexts which yield the 
highest activation values and the analysis we present is applicable 
to uni-directional RNN architectures.
Through a qualitative examination of these contexts we
observe that in case of {\sc Imaginet} some of the contexts represent a particular 
syntactic category or dependency function, while others correspond to a semantic theme. 
We identify cases where the hidden units encode characteristics of 
the context that go beyond their lexical properties, and represent abstract 
patterns that are useful for the network's task. 
Furthermore, we explore and visualize units that carry their
activation over to later time steps to 
extract longer dependencies and more complex linguistic features. 
We also quantitatively show that features encoded by the language model are more associated 
with syntactic constructions than in case of the image prediction pathway.

%\todo{No mutual information results. should we say something about that?}
%
%Lastly, we perform a comparative analysis between {\sc Visual} and {\sc Textual} and 
%present quantitative evidence that the features encoded by the hidden units of {\sc 
%Textual} are more focused on syntactic patterns than in case of {\sc Visual}.
%\todo{Not sure about this last line...}

%In all our experiments we use fixed RNN architecture with Gated Recurrent Units (GRU) 
% and train them on there tasks: 1) sentiment analysis ;
% 2) language modelling ; and 3) image prediction (predicting the visual 
% representation of an input word sequences). These 
%three tasks require different features to be extracted from
%linguistic data, which forces us to develop methods that generalize
%to a variety of models.

%Recurrent Neural Networks (RNNs) were introduced by 
%\cite{elman1990finding} as a connectionist architecture with the ability to model the temporal
%dimension. This feature makes RNNs particularly suitable for modelling 
%tasks that process input sequences, such as a variety of language 
%processing tasks. \cite{elman1991distributed} himself investigated the nature of the
%distributed representations built by a simple recurrent network when
%reading sentences and learning to predict each coming word, and showed 
%that the network grasped some notion of lexical categories among the 
%(very) small set of words in the input which show a syntactically similar 
%behaviour (e.g., edible or inanimate objects).

%RNNs can usually detect and generalise structural information that is implicitly present 
%in the sequence data and is useful for the task they are optimised for; however, the exact 
%nature of the structural  knowledge learned by RNNs is hard to pinpoint.

%With the recent surge of interest in artificial neural networks comes the 
%need for techniques for analysing the internal dynamics as well as the 
%observable behaviour of such networks in order to understand the type of 
%structural, higher-level knowledge they extract and use from the input data. 
%Some efforts have been made in the domain of computer vision: .... 
%\todo[inline]{guys, please summarise the current approaches in a couple of 
%sentences.}

%Similar studies are needed in the NLP domain, where the emergence of 
%syntactic knowledge from sequence of words is of special interest. The 
%extent and complexity of the acquired syntactic knowledge depends on the 
%nature of the task a model is trained for. In this paper we explore what 
%linguistic features RNNs learn from unannotated word sequences in order 
%to solve tasks that involve different levels of natural language 
%understanding. More specifically, we are interested in 
%\begin{itemize}
%We focus on different levels of describing the models as well as different 
%types of the tasks these models solve. The experiments explore:
%\item the nature of the syntactic structure the RNNs learn;
%\item the syntactic categories of words different models pay most attention to;
%\item the specific function of (some of) the individual neurons in these 
%models;
%\item the level of separation between syntactic and semantic 
%representation.
%\end{itemize}

%In this study we follow in this tradition but take onboard all the recent
%developments in recurrent architectures and their application in a
%variety of domains. Specifically, w

%We work with gated units able to
%learn long dependencies, and in addition to language modeling, we
%consider other tasks such as predicting sentiment and predicting
%features of the visual scene corresponding to a sentence.  Based on
%our experiments we show that depending on the nature of the prediction
%task on which a network is trained, it learns to encode either (a)
%aspects of the syntactic structure of the sentence, or (b) aspects of
%its meaning: concretely, a language modeling network will be
%comparatively more sensitive to function words encoding syntactic
%information, while a network predicting visual features pays most
%attention to semantically rich content words. Additionally we also
%show how the representations built by recurrent neural network evolve
%over time while incorporating each coming word: syntactically parallel
%and semantically related sentences trace similar trajectories in the
%vector space of the hidden layer states.

### 2017-05-16 doc/models.tex
paper files added

\section{Models}

%\subsection{Gated Recurrent Neural Networks}
%\label{sec:gru}

One of the main difficulties for training traditional Elman networks
arises from the fact that they overwrite their hidden states at every
time step with a new value computed from the current input $x_{t}$ and
the previous hidden state $\mathbf{h_{t-1}}$. Similarly to LSTMs,
Gated Recurrent Unit networks introduce a mechanism which facilitates the retention of 
information over multiple time steps.
Specifically, the GRU computes the hidden state at current time step, $\mathbf{h}_{t}$, as the
linear combination of previous activation $\mathbf{h_{t-1}}$, and a new
{\it candidate} activation $\mathbf{\tilde{h}}_t$:
%
\vspace{-.2cm}
\begin{equation}
  \mathrm{GRU}(\mathbf{h}_{t-1}, \mathbf{x}_t) = (1 - \mathbf{z}_t)\odot \mathbf{h}_{t-1} + \mathbf{z}_t \odot \mathbf{\tilde{h}}_t
\vspace{-.1cm}
\end{equation}
%
where $\odot$ is elementwise multiplication, and the update gate
activation $\mathbf{z_{t}}$ determines the amount of new information
mixed in the current state:
%
\vspace{-.1cm}
\begin{equation}
\label{eq:gru-update}
   \mathbf{z}_t = \sigma_s(\mathbf{W}_z \mathbf{x}_t + \mathbf{U}_z \mathbf{h}_{t-1})
\vspace{-.1cm}
\end{equation}
%
The candidate activation is computed as:
%
\vspace{-.2cm}
\begin{equation}
\label{eq:gru-cand}
   \mathbf{\tilde{h}}_t = \sigma(\mathbf{W} \mathbf{x}_t + \mathbf{U}(\mathbf{r}_t \odot \mathbf{h}_{t-1}))
\vspace{-.1cm}
\end{equation}
%
The reset gate $\mathbf{r_{t}}$ determines how much of the current
input $\mathbf{x_{t}}$ is mixed in the previous state
$\mathbf{h}_{t-1}$ to form the candidate activation:
%
\vspace{-.2cm}
\begin{equation}
\label{eq:gru-reset}
   \mathbf{r}_t = \sigma_s(\mathbf{W}_r \mathbf{x}_t + \mathbf{U}_r \mathbf{h}_{t-1})
\vspace{-.1cm}
\end{equation}

% \iffalse
% \subsection{Sentiment Classifier}
% \label{sec:sentiment}
% \todo{Describe sentiment task and classifier in proper detail}

% For some of the experiments we trained a sentiment classifier on the
% Sentiment Tree-bank \cite{socher2013recursive} a widely used data set
% to test neural network models for natural-language processing. It is a
% phrase level sentiment classification data set containing sentiment
% labels for parse tree constituents, from sentences to word level, for
% 215,154 phrases in 11,855 sentences. We trained our sentiment
% classifier to perform the 5 level fine-grained (very positive,
% positive, neutral, negative and very negative) sentiment
% classification task. The input parse trees were flattened, by
% transforming each parse tree node into a sequence of tokens. Each
% token sequence of length $n$ is first mapped to hidden representations
% $h_{t_{0}}, \ldots, h_{t_{n}}$ and the last representation is fed into
% a softmax classifier layer to predict the final sentiment score. The
% architecture we used is a simple single layer Gated Recurrent Unit
% Network:

% \begin{enumerate}
% \item Embedding layer
% \item GRU layer with clipped rectified activation
% \item Softmax Output layer
% \end{enumerate}

% The model was trained with the Adam optimizer using mini-batches, $L_{2}$ regularization and no dropout. The number of hidden neurons and the dimensionality of the word embeddings was set to 50. 
% \fi

\subsection{Imaginet}
\label{sec:imaginet}

{\sc Imaginet} introduced in \cite{chrupala2015learning} is a
multi-modal GRU network architecture that learns visually grounded
meaning representations from textual and visual input. It consists of
two GRU pathways, {\sc Textual} and {\sc Visual}, with a shared
word-embedding matrix. The inputs to the model are pairs of image descriptions 
and their corresponding images. Each sentence is mapped to two
sequences of hidden states, one by {\sc Visual} and the other by {\sc
  Textual}:
\vspace{-.2cm}
\begin{align}
  \mathbf{h}^{V}_t = \mathrm{GRU}^V(\mathbf{h}^{V}_{t-1},\mathbf{x}_t)\\
  \mathbf{h}^{T}_t = \mathrm{GRU}^T(\mathbf{h}^{T}_{t-1},\mathbf{x}_t)
\vspace{-.1cm}
\end{align}
%
At each time step {\sc Textual} predicts the next word in the sentence
$S$ from its current hidden state $\mathbf{h}^{T}_t$, while {\sc
  Visual} predicts the image-vector\footnote{Representing the full image, 
  extracted from the pre-trained Convolutional Neural Network of 
  \namecite{simonyan2014very}. \label{edit:dumdumeddy}}
$\hat{\mathbf{i}}$ from its last
hidden representation $\mathbf{h}^{V}_t$.
%
\vspace{-.2cm}
\begin{align}
   \hat{\mathbf{i}} &= \mathbf{V} \mathbf{h}^{V}_\tau \\
    p(S_{t+1}|S_{1:t}) &= \mathrm{softmax}(\mathbf{L}\mathbf{h}^T_t)
\vspace{-.1cm}
\end{align}
%
The loss function is a multi-task objective which penalizes error on
the visual and the textual targets simultaneously. The objective
combines cross-entropy loss $L^{T}$ for the word predictions
and cosine distance $L^V$ for the image
predictions.
%
\vspace{-.2cm}

\begin{equation}
\label{eq:lossce}
 L^{T}(\theta) = {-} \frac{1}{\tau}\sum_{t=1}^\tau \log p(S_t|S_{1:t})
\end{equation}

\begin{equation}
\label{eq:losscos}
 L^V(\theta) =  1 - \frac{\hat{\mathbf{i}} \cdot \mathbf{i}}{\|
     \hat{\mathbf{i}} \| \| \mathbf{i} \|}
\end{equation}

\begin{equation}
\label{eq:losscombo}
     L = \alpha L^T + (1-\alpha)L^{V}
\end{equation}
%L^V(\theta) &= \frac{1}{K}\sum_{k=1}^K(\hat{i}_k-i_k)^2
%

For more details about the {\sc Imaginet} model and 
its performance see \namecite{chrupala2015learning}.

In our version of the model we add two minor modifications 
compared to the original {\sc Imaginet}. As the performance of
{\sc Visual} is measured on image-retrieval, which is based on cosine 
distances we use cosine distance as the visual loss.
Moreover, we observe that using standardized image vectors, where each dimension 
is transformed by subtracting the mean and dividing by standard deviation further
improves performance.

%\todo{should we include somewhere where the image representations are coming from?}
% The model achieves near state-of-the-art results on MS-COCO image
% retrieval and good performance on other natural language
% understanding tasks. For the detailed results and more information
% please consult

% he model performance was evaluated on the test on image-retrieval
% and paraphrase retrieval - we consider captions belonging to the
% same image as paraphrases. Image retrieval is performed by the {\sc
% Visual} pathway by projecting the validation captions to the image
% space and ranking the images according to there cosine similarity to
% the projection. The model achieves a performance of Recall@1=0.06,
% Recall@5=0.20 and Recall@10=0.29. The paraphrase retrieval task is
% performed using the last hidden states of {\sc Textual} or {\sc
% Visual} to represent the captions in the validation set. As
% demonstrated by Table \ref{tab:retrieval} the representations
% computed by {\sc Textual} captures the paraphrase relationship
% better in the particular setting.

### 2017-05-16 doc/omission-stat/analysis.Rmd
paper files added

---
output: html_document
---
## Analysis of omission scores for textual and visual pathway of Imaginet

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
read.data <- function(f) { 
  data <- read.csv(f, col.names= c('omit','id','word','pos','deprel', 'score')) %>% 
  select(-omit) %>% na.omit()
  return(data[! data[,'pos'] %in%c('<END>','END'),])
}
data_v <- read.data('../data/visual_full_omission.csv')
data_v$pathway <- "visual"
data_v$deprel <- reorder(data_v$deprel, data_v$score, median)
data_v$pos <- reorder(data_v$pos, data_v$score, median)
data_t <- read.data('../data/textual_full_omission.csv')
data_t$pathway <- "textual"
data_t$deprel <- reorder(data_t$deprel, data_v$score, median)
data_t$pos <- reorder(data_t$pos, data_v$score, median)

data <- rbind(data_t , data_v) %>% mutate(pathway=as.factor(pathway))
```
```{r}
filter_by_count <- function(dat, col, count) {
  dat[dat[,col] %in% names(table(dat[,col]))[table(dat[,col]) > count],]
}

MINCOUNT<-500
subset <- filter_by_count(data %>% filter(pathway=='visual'),"pos", MINCOUNT)
ggplot(subset, aes(x=pos, y=score)) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="visual-omission-pos-boxplot.png",width=5, height=5)

subset <- filter_by_count(data %>% filter(pathway=='visual'),"deprel",MINCOUNT)
ggplot(subset, aes(x=deprel, y=score)) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="visual-omission-deprel-boxplot.png",width=5, height=5)

subset <- filter_by_count(data %>% filter(pathway=='textual'),"pos",MINCOUNT)
ggplot(subset, aes(x=pos, y=score)) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="textual-omission-pos-boxplot.png",
       width=5, height=5)

subset <- filter_by_count(data %>% filter(pathway=='textual'),"deprel",MINCOUNT)
ggplot(subset, aes(x=deprel, y=score)) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="textual-omission-deprel-boxplot.png",
       width=5, height=5)

subset <- filter_by_count(data,"pos",MINCOUNT*2)
ggplot(subset, aes(x=pos, y=score, color=pathway)) + 
  geom_boxplot(outlier.size = 1, outlier.colour='grey', varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="imaginet-omission-pos-boxplot.png",
       width=5, height=5)

subset <- filter_by_count(data,"deprel",MINCOUNT*2)
ggplot(subset, aes(x=deprel, y=score, color=pathway)) + 
  geom_boxplot(outlier.size = 1, outlier.colour='grey', varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="imaginet-omission-deprel-boxplot.png",
       width=5, height=5)

```
## Skipthought
```{r}
#data_skip <- rbind(data_v, read.data("../data//skipthought_omission.csv") %>% mutate(pathway="skipthought")) %>% rename(model=pathway)

#ggplot(data_skip %>% filter_by_count("deprel",MINCOUNT*2), aes(x=deprel, y=score, color=model)) +
#  geom_boxplot(outlier.size=1, varwidth=FALSE) + 
#   coord_flip() +
#   theme(text=element_text(size=15))
# ggsave(file="skipthought-omission-deprel-boxplot.png",
#        width=5, height=5)
```

### Log scores ratios per dependency
```{r}
ratio <- filter(data, pathway=='visual')$score / filter(data, pathway=='textual')$score  
data_r <- filter(data, pathway=='visual')
data_r$ratio <- ratio
data_r$deprel <- reorder(data_r$deprel, data_r$ratio, median)
data_r$pos <- reorder(data_r$pos, data_r$ratio, median)
```
```{r}
ggplot(filter_by_count(data_r,"pos",MINCOUNT) %>% filter, 
       aes(x=pos, y=log(ratio))) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) + 
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="omission-pos-boxplot.png", width=5, height=5)
ggplot(filter_by_count(data_r,"deprel",MINCOUNT), aes(x=deprel, y=log(ratio))) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) + 
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="omission-deprel-boxplot.png", width=5, height=5)
m1 <- lm(log(ratio) ~ pos, data=data_r)
m2 <- lm(log(ratio) ~ deprel, data=data_r)
m3 <- lm(log(ratio) ~ deprel+pos, data=data_r)

summary(m1)
summary(m2)
summary(m3)
```
### Sentiment
```{r}
# data_sent <- read.data("../data/sentiment_omission.csv") 
# data_sent$deprel <- reorder(data_sent$deprel, data_sent$score, median)
# data_sent$pos <- reorder(data_sent$pos, data_sent$score, median)
# 
#             
# subset <- filter_by_count(data_sent,"pos",300)
# p.sent.pos <- ggplot(subset, aes(x=pos, y=score)) + 
#   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) 
# ylim1 <- boxplot.stats(filter(subset, pos=="JJS")$score)$stats[c(1, 5)]
# p.sent.pos + 
#   coord_flip(ylim=ylim1*1.05) +
#   theme(text=element_text(size=15))
# ggsave(file="sent-omission-pos-boxplot.png",width=5, height=5)
# 
# subset <- filter_by_count(data_sent,"deprel",300)
# p.sent.deprel <- ggplot(subset, aes(x=deprel, y=score)) + 
#   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) 
# ylim1 <- boxplot.stats(filter(subset, deprel=="root")$score)$stats[c(1, 5)]
# p.sent.deprel + 
#   coord_flip(ylim=ylim1*1.05) +
#   theme(text=element_text(size=15))
# ggsave(file="sent-omission-deprel-boxplot.png",
#        width=5, height=5)
# 
# ggplot(data_r %>% filter(pos=="VBZ") %>% filter_by_count(col = "deprel", 100), aes(x=deprel, y=log(ratio))) + geom_boxplot(varwidth = TRUE) + coord_flip()
# ggsave(file="omission-vbz-boxplot.pdf")
```

```{r}
data <- read.csv("visual_omission_with_position.csv")
data <- data %>% mutate(pos=reorder(pos, score),
                        deprel=reorder(deprel, score)) #%>% filter(pos %in% c("NN","NNS"))
#m0 <- lm(log(score) ~ pos, data=data)
#m1 <- lm(log(score) ~ pos + as.factor(position), data=data)
#summary(m0)
#summary(m1)
cor(data %>% select(position, score), method = "pearson")
cor(data  %>% filter(pos %in% c("NN","NNP","NNS")) %>% select(position, score), method = "pearson")

ggplot(data %>% filter_by_count("pos",MINCOUNT*2), 
            aes(x=position, color=pos, y=score)) +
  geom_point(position = "jitter", alpha=0.03) +
  geom_smooth() +
  xlim(0,20) +
  ylim(0.0, 0.5) +
  theme(aspect.ratio=2/3, text=element_text(size=25)) + 
  ggtitle("Visual")
ggsave(file="visual-omission-by-position.png")

```

```{r}
data <- read.csv("textual_omission_with_position.csv")
data <- data %>% mutate(pos=reorder(pos, score),
                        deprel=reorder(deprel, score))
cor(data  %>% select(position, score), method = "pearson")
#m0 <- lm(log(score) ~ pos, data=data)
#m1 <- lm(log(score) ~ pos + as.factor(position), data=data)
#summary(m0)
#summary(m1)

ggplot(data %>% filter_by_count("pos",MINCOUNT*2), 
            aes(x=position, color=pos, y=score)) +
  geom_point(position = "jitter", alpha=0.03) +
  geom_smooth() +
  xlim(0,20) +
  ylim(0.0, 0.5) + 
  theme(aspect.ratio=2/3, text=element_text(size=25)) +
  ggtitle("Textual")
ggsave(file="textual-omission-by-position.png")

```
```{r}
library(dplyr)
library(ggplot2)

top_words <- read.csv("top_words.csv") %>% 
  mutate(deprel = reorder(deprel, score))
ggplot(top_words %>% filter_by_count("deprel",70), 
       aes(x=word, y=score, color=deprel)) +
  geom_boxplot(position = "dodge") +
   theme(text=element_text(size=25))
ggsave(file="top_words.pdf")
```

### 2017-05-16 doc/original_reviews.tex
paper files added

\documentclass[10pt,a4paper]{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}

\begin{document}

I'm writing to you in my capacity as the TACL action editor for your
submission 882, "Representation of linguistic form and function in recurrent
neural networks". While we are not accepting your paper in its current form,
we do encourage you to revise and resubmit it within 3-6 months.

We received three reports. These are included below. Please do consider all
of the points they make as you revise the work. Some high-level themes:

* The paper is pitched in very broad terms (methods for analyzing RNNs), but
the methods are actually quite specific to classifier-type RNNs, and many
aspects of the work seem applicable only to the Imaginet setting. I would
suggest clarifying the scope of the claims.

* Similarly, as Reviewer A points out, some of the rhetoric is misaligned
with the methods and results.

* The core results for content words and functions words seem like the
expected ones given their linguistic content. In particular, the omission
score method seems to show that the networks learn to identify
discriminating content words, and that
changing or removing those words has a dramatic effect on final predictions.
This seems like a prerequisite (given the linguistic facts) for learning an
effective classifier. Regular train/dev/test experimental analyses already
show that these models are effective. Could you further articulate what we
learn via the omission score method that we don't get from standard
evaluations?

* The connections with POS tags and dependency relations are intriguing.
Could you, though, clarify the significance of these findings? My concern is
that they just reflect the fact that the content-word/function-word
distinction is highly correlated with POS and dependency relation.

* I had trouble understanding the information structure section. I agree
with the core insight about old/topic before new/comment in languages like
English. I had trouble relating this to what I would expect from the RNN.
For instance, the structure of the RNN seems to ensure that information
nearer the end of the sequence will have a larger impact. Is the Textual
omission score result due to this factor or is it truly due to something
about language? Conversely, for Visual, I guess I would have anticipated
exactly the opposite of what is reported here: the comment part should be
the most important, corresponding to the new and visually salient elements,
rather than those that can fade into the background.

For additional questions and concerns, and some suggestions for how to
improve accessibility and readability, please see the full reports.

If you do choose to revise and resubmit, please make use a *new* submission
number, and follow the instructions in section "Revision and Resubmission
Policy for TACL Submissions" at
\url{https://tacl2013.cs.columbia.edu/ojs/index.php/tacl/about/submissions#authorGuidelines.}
We  allow you 1-2 additional pages in the revised version for addressing the
referees' concerns.

Please understand that while we have endeavored to provide some guidance on
how to revise the manuscript, the paper will be reviewed afresh should you
choose to resubmit (possibly involving a change of action editor and
reviewers), with no guarantee of acceptance.

Thank you for considering TACL for your work, and again, I encourage you to
revise and resubmit within the specified timeframe.

---Chris

Christopher Potts
Stanford University
cgpotts@stanford.edu

-----------------------
THE REVIEWS
----------------------- \\
Reviewer A:

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        3. Mostly understandable to me with some effort.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        2. Pedestrian: Obvious, or a minor improvement on familiar techniques.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        3. Fairly reasonable work. The approach is not bad, and at least the main
claims are probably correct, but I am not entirely ready to accept them
(based on the material in the paper).

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        3. Bibliography and comparison are somewhat helpful, but it could be hard
for a reader to determine exactly how this work relates to previous work or
what its benefits and limitations are.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        2. Work in progress. There are enough good ideas, but perhaps not enough
results yet.

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        3. Interesting but not too influential. The work will be cited, but mainly
for comparison or as a source of minor contributions.

REPLICABILITY: Will members of the ACL community be able to reproduce or
verify the results in this paper?:
        3. could reproduce the results with some difficulty. The settings of
parameters are underspecified or subjectively determined; the
training/evaluation data are not widely available.

IMPACT OF PROMISED SOFTWARE:  If the authors state (in anonymous fashion)
that their software will be available, what is the expected impact of the
software package?:
        1. No usable software released.

IMPACT OF PROMISED DATASET(S): If the authors state (in anonymous fashion)
that datasets will be released, how valuable will they be to others?:
        1. No usable datasets submitted.

TACL-WORTHY AS IS? In answering, think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
recommend it. If a paper is solid but you could live without it, let us know
that you're ambivalent.

Note: after you submit this review form, you'll need to answer a related but
different question via a pull-down menu: how long would it take for the
authors to revise the submission to be TACL-worthy?
:
        3. Ambivalent: OK but does not seem up to the standards of TACL.

Detailed Comments for the Authors:
        The paper presents some empirical studies on the task of visualising the
internal representations of the RNN on some language tasks. The authors
specifically focused on the RNN architectures proposed by Chrupala et al
(2015) and focused on two tasks - language modelling and predicting visual
representations from word sequences.

The authors focused on the discussions on the understanding of the hidden
activation vectors learned at the end of RNN as well as the understanding of
the dimensions in the learned vector representations. For the first task,
the authors used the "omission" trick -- excluding a certain single word in
the sentence, and generate a different vector representation at the end, and
make comparisons with the original vector representation. The authors then
tried to draw some conclusions on the importance of each word based on the
omission scores.

One question about the omission score: the lower the score is, the low the
similarity is. Thus, when dropping the important words, the similarity
should go down -- so a lower omission score. It is not clear to me why
dropping important words will lead to a higher omission score.

Although I agree it's interesting to do such analysis: especially here we
have two different tasks -- one involves visual aspects which is very nice,
there are two important questions. First, why is the architecture of
Chrupala et al (2015) selected only? Is it representative enough for general
RNNs? What about LSTM and its variants for language modelling task? How
would they be different? It would be good to see such discussions. Second,
most of the paper presents qualitative results, without rigorous
quantitative justifications. It would be nice, however, if the authors could
collect some manual annotations and conduct automatic evaluations based on
such annotations. For example, such annotations could include human's
judgements on how important each word (or word class, syntactic class) in
the sentence are etc. Finally, although the visualisation is helpful and it
might help the readers understand the underlying activities of the RNN,
little is known about why. I understand it is just an empirical paper,
however, it would be helpful to at least share with the readers some high
level insights so that the readers could learn more from this paper. One
question that I have in mind after reading this paper is: these results
related to RNN presented here are expected. It's good that RNN is able to
capture such things. However, how do I make use of the results here in my
research? How does this have some impact on my work? The authors claimed at
the end of the paper that "The goal of our paper is to propose novel methods
for the encoding of linguistic knowledge in RNNs" -- I am not sure if I am
convinced that this is goal of this work. It is largely an empirical paper
that tries to visualise/understand the RNN which was previously regarded as
a blackbox, but does not really "propose" methods for encoding linguistic
knowledge in RNN.

The explanations of several places, such as Sec 6.1 and Sec 6.2 could be
improved. It was a little difficult for me to follow.

Minor: diffeent -> different.

REVIEWER CONFIDENCE:
        3. Pretty sure, but there's a chance I missed something. Although I have a
good feel for this area in general, I did not carefully check the paper's
details, e.g., the math, experimental design, or novelty.

-----------------------
Reviewer B:
-----------------------

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        5. Very clear.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        3. Respectable: A nice research contribution that represents a notable
extension of prior approaches or methodologies.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        4. Generally solid work, although there are some aspects of the approach or
evaluation I am not sure about.

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        4. Mostly solid bibliography and comparison, but there are a few additional
references that should be included. Discussion of benefits and limitations
is acceptable but not enlightening.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        4. Represents an appropriate amount of work for a publication in this
journal. (most submissions)

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        4. Some of the ideas or results will substantially help other people's
ongoing research.

REPLICABILITY: Will members of the ACL community be able to reproduce or
verify the results in this paper?:
        4. could mostly reproduce the results, but there may be some
variation because of sample variance or minor variations in their
interpretation of the protocol or method.

IMPACT OF PROMISED SOFTWARE:  If the authors state (in anonymous fashion)
that their software will be available, what is the expected impact of the
software package?:
        1. No usable software released.

IMPACT OF PROMISED DATASET(S): If the authors state (in anonymous fashion)
that datasets will be released, how valuable will they be to others?:
        1. No usable datasets submitted.

TACL-WORTHY AS IS? In answering, think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
recommend it. If a paper is solid but you could live without it, let us know
that you're ambivalent.

Note: after you submit this review form, you'll need to answer a related but
different question via a pull-down menu: how long would it take for the
authors to revise the submission to be TACL-worthy?
:
        4. Worthy: A good paper that is worthy of being published in TACL.

Detailed Comments for the Authors:
        This paper provides an in-depth qualitative and quantitative of a joint
caption-generation and image selection model (IMAGINET). It casts the said
analysis as a general procedure for introspecting a trained model through
several methods. It examines the effect of removing words on the final
representation of a text-predicting RNN and an image predicting RNN as a
measure of word, POS, dependency relation, and positional importance. It
also examines how various words and word classes contribute to increases in
hidden layer activation, or preservation thereof. The results are entirely
unsurprising, in that content words matter for image matching, and function
words affect textual meaning while preserving information in the hidden
layer. The fact that these results match obvious intuitions should not be
taken as criticism of the paper in any way, and I am satisfied that the
method by which they were obtained is sound.

Overall, the paper is clear, well written, well presented. Such attempts to
analyse trained recurrent neural networks from a linguistic perspective is
furthermore a timely and welcome contribution, given their near-ubiquity in
*ACL publications (at least in semantics and ML tracks). The results of the
analysis are convincing, and I would support publication of this paper in
TACL. Below are a few criticisms and suggested improvements that would help
improve the paper.

The main criticism I would have of this paper is that it sets out as a
providing generic new method for analysing RNNs with text input, amongst
other modalities. I think this is overstating the contents of the paper: the
methods of analysis are fairly tied to the nature of the model
(unidirectional RNN without attention) and broad task (producing a final
representation that can be used for classification or generation). It would
have been interesting to the same analyses performed on other, related
models (bi-directional RNN, word-based log-bilinear models, etc), and on RNN
tasks with other objectives (machine translation, language modelling,
sequence labelling). I realise it's a tall order to add this to the current
work, but perhaps re-stating the scope of the work (providing an example of
how to perform further qualitative and quantitative analysis of a an
existing model trained on a specific task) and briefly discussing how to
adapt the experiments to other domains would be good. Likewise, a discussion
comparing and contrasting these methods of analysis with other
(complementary?) visualisation methods for RNNs with attention such at
Badhanau et al. 2015 or Rocktaschel et al. 2015 [ICLR 2016] (which could be
cited here) would be welcome. To summarise this point, I do not believe the
authors have willingly misrepresented the aim and scope of their work, but
think it would benefit from being restated a bit more clearly, bearing in
mind the comments made above.

Other more minor comments:

* It would be nice to get a clear statement of the objective being optimised
against in eq 9.
* Figure 2 a little hard to read.
* Figure 6 very hard to read.
* Section 5.1 "textual is more sensitive to ..." -> it looks like textual is
more sensitive to most words being removed to some extent. I don't know if
it's fair to single out these categories. Instead, just contrast more
clearly with the heavy focus on content words (mainly NNs) for Visual.
* Section 5.2, model 1 vs model 2 <- is the a significant difference in
number of features here? Model 2 seems like it would have more features, and
intuitively, more info == better model (modulo feature quality), so quick
comment on this here would be good.
* Section 5.3 is very tied to structure of the model use in this paper
(would expect different results for bidirection RNN, recursive neural net).
Comment on this here would be good.
* Not a fan of qualitative analysis as done in Table 1/2. Too easy
over-interpret, but I realise this is personal taste. Would not be offended
if this is left in, but would suggest scrapping this to create space for
discussions suggested above.

REVIEWER CONFIDENCE:
        5. Positive that my evaluation is correct. I read the paper very carefully
and am familiar with related work.

-----------------------
Reviewer C:
-----------------------

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        3. Mostly understandable to me with some effort.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        4. Creative: An intriguing problem, technique, or approach that is
substantially different from previous research.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        4. Generally solid work, although there are some aspects of the approach or
evaluation I am not sure about.

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        4. Mostly solid bibliography and comparison, but there are a few additional
references that should be included. Discussion of benefits and limitations
is acceptable but not enlightening.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        4. Represents an appropriate amount of work for a publication in this
journal. (most submissions)

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        5. Will affect the field by altering other people's choice of research
topics or basic approach.

REPLICABILITY: Will members of the ACL community be able to reproduce or
verify the results in this paper?:
        3. could reproduce the results with some difficulty. The settings of
parameters are underspecified or subjectively determined; the
training/evaluation data are not widely available.

IMPACT OF PROMISED SOFTWARE:  If the authors state (in anonymous fashion)
that their software will be available, what is the expected impact of the
software package?:
        1. No usable software released.

IMPACT OF PROMISED DATASET(S): If the authors state (in anonymous fashion)
that datasets will be released, how valuable will they be to others?:
        4. Useful: I would recommend the new datasets to other researchers or
developers for their ongoing work.

TACL-WORTHY AS IS? In answering, think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
recommend it. If a paper is solid but you could live without it, let us know
that you're ambivalent.

Note: after you submit this review form, you'll need to answer a related but
different question via a pull-down menu: how long would it take for the
authors to revise the submission to be TACL-worthy?
:
        5. Strong: I'd like to see it accepted; it will be one of the better papers
in TACL.

Detailed Comments for the Authors:
        Summary :

This paper addresses questions on visualization and understanding of
activation patterns of recurrent neural networks given a paired
linguistic-visual representational input. Though sharing the same word
embeddings, the visual and linguistic neural pathways triggered different
activation patterns, manifested in the form of different omission scores
distributions at the end of each pathway, when variants of the same input
sentence are given as inputs to the RNN. The paper also gave an in-depth
analysis of the hidden units by comparative evaluations of "contexts", which
are encodings at the level of the hidden units, between the two neural
pathways.

Merits:

- novel visualization into inner workings of NN, and from an
easier-to-grasp, higher level linguistic perspective. To my knowledge, NN
are known to outperform many other learners, yet efforts on
visualization/understanding of their inner layers is rare due to their
inherent complexity.

- this work has the ability to generalize from dual linguistics
representations to n-modal representation, where there are increasingly more
work on multimodal-feature approach to solving old tasks. The principled
approach used in this paper can be used to explain away the orthogonal
contribution provided by each neural pathway trigged by each of the
modalities to gain a better understanding of the network.

Issues/concerns:

The paper builds heavily on previous work "IMAGINET", is densely described
but lacking details in certain sections. For example, I find it rather
challenging to understand the following:

- the process of image vector prediction is vague (eqn 7). Does it mean
prediction at the pixel level of an entire image, or a blob or principal
image segment of interest? Also, how does the image retrieval shown in
Figure 1 operate?

- Figure 6 would benefit from showing the aggregated POS groupings such as
"VBN,VBG,VB" all grouped into "VERB" etc. for benefit of clarity in a small
graph (too cluttered)

REVIEWER CONFIDENCE:
        3. Pretty sure, but there's a chance I missed something. Although I have a
good feel for this area in general, I did not carefully check the paper's
details, e.g., the math, experimental design, or novelty.

\end{document}

### 2017-05-16 doc/regression.tex
paper files added

\section{Representing syntactic information}
\label{sec:regression}

It is assumed that to perform a task that involves some level of language 
understanding, RNNs need to learn some level syntactic knowledge from the 
linguistic input. To uncover if the networks indeed represent syntactic 
information, we trained a number of Logistic Regression models on various 
combinations of n-gram tokens in the input as well as the hidden states of the 
networks to predict the syntactic categories of the tokens at each time-step $t$. 
From the sentences we extracted unigram and n-gram features up to a window 
size of 4; for example, to predict the label for {\it dog} in the sentence {\it the nice 
dog} we extract {\it the}$_2$, {\it nice}$_1$, {\it dog}$_0$, {\it the$_2$ nice$_1$}, 
{\it nice$_1$ dog$_0$}, and {\it the$_2$ nice$_1$ dog$_0$}, etc.

We trained several models on a combination of the following features:
\begin{itemize}
  \item current word $w_t$
  \item current word-embedding of {\sc Imaginet} $e_t^I$ (same for {\sc Visual} 
and {\sc Textual})
  \item current word-embedding of {\sc Sentiment} $e_t^S$
  \item 4-gram features
  \item hidden state of {\sc Visual} $h_t^V$
  \item hidden state of {\sc Textual} $h_t^T$
  \item hidden state of {\sc Sentiment} $h_t^S$
%  \item hidden state of {\sc Visual} $h_t^V$ + 4-gram features
%  \item hidden state of {\sc Textual} $h_t^T$ + 4-gram features
\end{itemize}

The motivation behind this experiment is that if the networks generalize over 
syntactic categories, then using their activation vectors as features will lead to a 
more accurate model f$h_t$or syntactic category prediction than only using the tokens 
themselves. If the embeddings $e_t$ and hidden activations $h_t$ outperform 
the word form $w_t$ and 4-gram models respectively, then the linguistic 
representations of the model do generalize over syntactic categories.

\begin{table}[]
\centering
\caption{Results of predicting syntactic categories for each token in the input to 
SENTIMENT and IMAGINET models, by Logistic Regression classifiers trained 
on features in the first column. The second and third columns report the micro F-
scores on DepRel and POS categories respectively.}

\label{tab:logistic}
\begin{tabular}{lll}
\hline
 \multicolumn{3}{c}{\sc Sentiment}           \\
 \hline 
{\bf Features}      &  {\bf deprel} &  {\bf POS}     \\
\hline 
word $w_t$                   		  & .50  & .65   \\
4-gram                       		  & .62  & .71   \\
$e^S_{t}$          & .33  & .38   \\ 
$h_t^S$                      		  & .30  & .32    \\
\hline 
\hline
\multicolumn{3}{c}{\sc Imaginet}                          \\
 \hline 
{\bf Features}      &  {\bf deprel} &  {\bf POS}     \\
\hline 
$w_t$                   		  	      & .710   & .93  \\
4-gram                         		  & .807  & .93  \\
$e^I_{t}$          			  	      & .73   & .95  \\ 
$h_t^V$                      		  & .751  & .88  \\
$h_t^T$                      		  & .811  & .94  \\
$h_t^T$ + $h_t^V$            		  & .81   & .94  \\
$h_t^V$ + 4-gram             		  & .809   & .93  \\
$h_t^T$ + 4-gram             		  & .823   & .95
\end{tabular}
\end{table}

Table~\ref{tab:logistic} shows the results of predicting the syntactic categories of 
tokens in the input sentences using different (combinations of) features. 
\todo{Are we going to change this to test set?}

The results for {\sc Sentiment} show that the representations (both 
$e^S_{t}$ and $h_t^S$) are less predictive of the grammatical categories of the 
target words than using the unigram and 4-gram features.

%In fact there are only 8 categories out of 44 with non-zero f-scores for deprels: 
%amod (0.28), cc (.64), conj (.13), det (.56), nsubj (.29), pobj (.30), prep (.43) and 
%root (.26).
However, in case of {\sc Imaginet}, using the embeddings $e_t$ leads to higher F-score for both dependency category and part-of-speech tag prediction of the tokens $w_t$ than using $w_{t}$'s 
themselves. This result provides evidence that syntactic information is indeed encoded in the learned word embeddings. 
Similarly, the classifier trained on the hidden activations of {\sc Textual } $h_t^T$ 
performs better on both DepRel and POS prediction than using 4-grams features, but only by a 
narrow margin.

The more interesting results, however, are the low F-scores when 
using $h_{t}^V$ as features. These results suggest that {\sc Visual} represents less 
syntactic information then {\sc Textual} about sequences of words. We also observe 
that the maximum F-score is achieved using the combinations of $h_{t}^T$ and the 4-
gram features, suggesting that perhaps some information about the word identities is 
lost in $h_{t}^T$.

### 2017-05-16 doc/related.tex
paper files added

\section{Related work}
\label{sec:related}

The direct predecessors of modern architectures were first proposed in the 
seminal paper of \cite{elman1990finding}. He modifies the recurrent 
neural network architecture of \cite{jordan1986attractor} by changing the output-to-memory 
feedback connections to hidden-to-memory recurrence, enabling Elman networks to 
represent arbitrary dynamic systems. \namecite{elman1991distributed} he trains an RNN on a small synthetic
sentence  dataset and analyzes the activation patterns of the hidden
layer. His analysis shows that these distributed representations  
encode lexical categories, grammatical relationships and hierarchical constituent 
structures. \cite{giles1992extracting} trains RNNs similar to Elman
networks on strings generated by small 
deterministic regular grammars with the objective to recognize positive and reject negative 
strings, and develops the \emph{dynamic state partitioning} technique to extract the learned 
grammar from the networks in the form of deterministic finite state
automatons.
% Dynamic state partitioning follows the general pattern employed in the
% large body of literature of RNN rule extraction: (i) using quantization
% to create a mapping from the continuous RNN state- space to a finite
% set of discrete spaces; (ii) passing the RNNs input and observing the
% state transitions; (iii) rule construction; and (iv) rule
% minimization. We refer the reader to \cite{jacobsson2005rule} for a
% comprehensive survey on RNN rule extraction.

More closely related is the recent work of \cite{li2015visualizing}, who develop techniques 
for a deeper understanding of the activation patterns of RNNs, but focus on models with 
modern architectures trained on large scale data sets. More specifically, they train Long Short-Term Memory networks (LSTM) \cite{hochreiter1997long} 
for phrase level sentiment analysis 
and present novel methods 
to explore the inner workings of RNNs. They measure the salience of tokens
in sentences by taking the first-order derivatives of the loss with respect to the
word embeddings and provide evidence 
that LSTMs can learn to attend to important tokens in sentences. Furthermore,
they plot the activation values of hidden units through time using heat maps 
and visualize local semantic compositionality in RNNs. 
In comparison, the present work focuses more on exploring structure 
learning in RNNs and on developing methods for a comparative analysis between RNNs.

Adding an explicit attention mechanism that allows the RNNs to focus on \label{edit:attention}
different parts of the input was recently introduced by \namecite{bahdanau2014neural}
in the context of extending the sequence-to-sequence 
RNN architecture for neural machine translation. 
At the decoding side this neural module assigns weights 
$\alpha_{1},\ldots,\alpha_{i}$ to the hidden states of the decoder 
$h_{1},\ldots,h_{i}$, which allows the decoder to selectively pay varying degrees of
attention to different phrases in the source sentence at different decoding time-steps.
They also provide qualitative analysis by visualizing the attention
weigths and exploring the importance of the source encodings at various decoding steps.
Similarly \namecite{rocktaschel2016reasoning} use an attentive neural 
network architecture to perform natural language inference and visualize which parts
of the hypotheses and premises the model pays attention to when deciding on the 
entailment relationship. Conversely, the present work focuses on RNNs without an 
explicit attention mechanism.

\namecite{karpathy2015visualizing} also take up the challenge of
rendering RNN activation patterns understandable, but use character
level language models and rather than taking a linguistic point of
view, focus on error analysis and training dynamics of LSTMs and
GRUs.  They show that certain dimensions in the RNN hidden activation
vectors have specific and interpretable functions. One of the goals of
the present paper is also to explore the specific functions encoded
by individual hidden units, but in a linguistically-informed way.

\namecite{li2015convergent} train a Convolutional Neural Networks (CNN)
with different random initializations
%based on the architecture introduced in \cite{krizhevsky2012imagenet} 
on the ImageNet dataset \cite{krizhevsky2012imagenet}. For each layer
in all networks they store the activation values produced on the
validation set of ILSVRC and align similar neurons of different
networks. They conclude that while some features are learned across
networks, some seem to depend on the initialization. Other works on
visualizing the role of individual hidden units in deep models for vision
synthesize images by optimizing random images through backpropagation
to maximize the activity of units
\cite{erhan2009visualizing,simonyan2013deep,yosinski2015understanding,nguyen2016multifaceted}
or to approximate the activation vectors of particular layers
\cite{mahendran2015visualizing,dosovitskiy2015inverting}.

In general, there has been a growing interest within computer vision
in understanding deep models, with a number of papers dedicated to
visualizing learned CNN filters and pixel saliencies
\cite{simonyan2013deep,yosinski2015understanding,mahendran2015understanding}. These
techniques have also led to improvements in model performance
\cite{eigen2013understanding} and transferability of features
\cite{zhou2014object}. To date there has been much less work on such
issues within computational linguistics. We aim to fill this gap by
adapting existing methods as well as developing novel techniques to
explore the linguistic structure learned by recurrent networks.

### 2017-05-16 doc/tacl_bundler.tex
paper files added

\documentclass{article}[11pt,oneside]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INPUTS FOR THE NEW PAPER
\newcommand{\subtitle}{Representation of linguistic form and function in recurrent neural networks}
\title{Resubmission of TACL, \\ \subtitle.   \\}

\newcommand{\resubmission}{paper.pdf}  %% revised version
\newcommand{\origdecision}{original_reviews.pdf}  %% name of anonymized original-decision-letter file
\newcommand{\coverfile}{response.pdf}  %% name of anonymized cover-letter file

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{hyperref}
\hypersetup{
    colorlinks=true, %set true if you want colored links
    linktoc=all,     %set to all if you want both sections and subsections linked
    linkcolor=blue,  %choose some color if you want links to stand out
}
\usepackage{ifthen}
\usepackage{xspace}

\usepackage{times}
\usepackage{fullpage}
\usepackage{pdfpages}

\special{papersize=8.5in,11in}
\pdfpageheight\paperheight
\pdfpagewidth\paperwidth
\setlength\topmargin{-7mm} \setlength\oddsidemargin{-0cm}
\setlength\textheight{22cm} \setlength\textwidth{15.8cm}
\setlength\columnsep{0.25in}  \newlength\titlebox \setlength\titlebox{2.00in}
\setlength\headheight{5pt}   \setlength\headsep{0pt}
\setlength\footskip{1.9cm}
\setlength\leftmargin{0.0in}
\pagestyle{plain}

\setlength{\footskip}{55pt}

\begin{document}
\maketitle

\tableofcontents

\section{Author(s) cover letter responding to the original reviews} Starts on next page.
\includepdf[pages=-,offset=0mm 6mm]{\coverfile}
\section{Revised submission} Starts on next page.
\includepdf[pages=-,offset=0mm 6mm]{\resubmission}
\section{Original decision letter and reviews} Starts on next page.
\includepdf[pages=-,offset=0mm 6mm]{\origdecision}

\end{document}

### 2017-05-16 doc/mi-stat/stat.R

@@ -1,0 +1,58 @@
+setwd("/home/gchrupala/repos/naacl2016/mi-stat")
+library(dplyr)
+library(ggplot2)
+
+
+data <- data.frame()
+for (n in 1:3) {
+  for (condition in c("deprel","token")) {
+    file <- paste("MI/new_results/",
+                  condition,
+                  "s",
+                  n,
+                  ".csv", sep="")
+    scores <- read.csv(file)
+    cond <- if (condition=="token") {"word"} else {"deprel"}
+    data <- rbind(data, data.frame(pathway="Visual",
+                           n=n,
+                           condition=cond,
+                           mi=as.numeric(scores[1,2:1025])))          
+    data <- rbind(data, data.frame(pathway="Textual",
+                                 n=n,
+                                 condition=cond,
+                                 mi=as.numeric(scores[2,2:1025])))          
+    }
+  }
+
+ggplot(data %>% filter(n==1), aes(x=condition, y=mi, color=pathway)) + 
+  geom_boxplot()
+
+library(boot)
+
+log_ratio <- function(data, ix) {
+  w <- log(median(filter(data[ix,], pathway=="Textual", condition=="word")$mi) / 
+        median(filter(data[ix,], pathway=="Visual", condition=="word")$mi),base=exp(1))
+  d <- log(median(filter(data[ix,], pathway=="Textual", condition=="deprel")$mi) / 
+             median(filter(data[ix,], pathway=="Visual", condition=="deprel")$mi),base=exp(1))
+  return(c(w=w,d=d))
+}
+
+
+library(reshape2)
+
+data_boot <- data.frame()
+for (i in 1:3) {  
+  res <- boot(data %>% filter(n==i), log_ratio, R=5000)
+  data_boot <- rbind(data_boot, data.frame(n=i, condition="word", log.ratio=res$t[,1]))
+  data_boot <- rbind(data_boot, data.frame(n=i, condition="deprel", log.ratio=res$t[,2]))
+}
+ggplot(data=data_boot,
+       aes(x=as.factor(n), color=condition, y=log.ratio)) +
+  geom_boxplot() +
+  coord_flip() +
+  xlab("n-gram range") +
+  ylab(expression(log(MI[C]^T / MI[C]^V ))) +
+  theme(text=element_text(size=20)) +
+  ggsave(file="bootstrappedMI.png",
+         width=6, height=4)
+

### 2017-05-16 doc/omission-stat/analysis.R

@@ -1,0 +1,180 @@
+# INSTALLATION
+
+#install.packages("Rcpp", dependencies=TRUE, repos="http://cran.r-project.org" )
+#install.packages("dplyr", dependencies=TRUE, repos="http://cran.r-project.org" )
+#install.packages("ggplot2", dependencies=TRUE, repos="http://cran.r-project.org")
+
+library("dplyr")
+library("ggplot2")
+# READING DATA
+
+read.data <- function(f) {
+  data <- read.csv(f, col.names= c('omit','id','word','pos','deprel', 'score')) %>%
+  select(-omit) %>% na.omit()
+  return(data[! data[,'pos'] %in%c('<END>','END'),])
+}
+data_v <- read.data('../data/visual_full_omission.csv')
+data_v$pathway <- "visual"
+data_v$deprel <- reorder(data_v$deprel, data_v$score, median)
+data_v$pos <- reorder(data_v$pos, data_v$score, median)
+data_t <- read.data('../data/textual_full_omission.csv')
+data_t$pathway <- "textual"
+data_t$deprel <- reorder(data_t$deprel, data_v$score, median)
+data_t$pos <- reorder(data_t$pos, data_v$score, median)
+
+data <- rbind(data_t , data_v) %>% mutate(pathway=as.factor(pathway))
+
+filter_by_count <- function(dat, col, count) {
+  dat[dat[,col] %in% names(table(dat[,col]))[table(dat[,col]) > count],]
+}
+
+MINCOUNT<-500
+subset <- filter_by_count(data %>% filter(pathway=='visual'),"pos", MINCOUNT)
+
+
+
+# OMISSION BOXPLOTS
+
+ggplot(subset, aes(x=pos, y=score)) +
+  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
+  coord_flip() +
+  theme(text=element_text(size=15))
+ggsave(file="visual-omission-pos-boxplot.png",width=5, height=5)
+
+subset <- filter_by_count(data %>% filter(pathway=='visual'),"deprel",MINCOUNT)
+ggplot(subset, aes(x=deprel, y=score)) +
+  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
+  coord_flip() +
+  theme(text=element_text(size=15))
+ggsave(file="visual-omission-deprel-boxplot.png",width=5, height=5)
+
[…]

### 2017-05-16 doc/omission-stat/me.R

@@ -1,0 +1,25 @@
+library(dplyr)
+library(ggplot2)
+library(lme4)
+
+# Visualization of predicting omission scores using various predictors
+data <- read.table(file="predictor_rsq.txt", header=TRUE) 
+for (model in levels(data$model)) {
+  data[data$model==model,"score"] = data[data$model==model,"rsq"]/data[data$model==model & data$predictors=="word","rsq"]
+}
+data <- data %>% mutate(predictors = reorder(predictors, rsq))
+ggplot(data, aes(x=model, color=predictors, y=score)) + geom_boxplot() +  theme(aspect.ratio=2/3, text=element_text(size=25)) +
+   ylab("R Squared relative to word")
+ggsave("position-new.png")
+
+data_coef<- read.table(file="position_coef.txt", header=TRUE)
+values <- c("first","second","third","middle","antepenult","penult","last")
+data_coef$predictor <- factor(data_coef$predictor, levels=values, ordered=TRUE)
+ggplot(data_coef, aes(x=as.numeric(predictor), y=coef, color=model)) + geom_point() + geom_line() +
+  xlab("Position") +
+  ylab("Coefficient") +
+  scale_x_continuous(breaks=1:length(values), 
+                     labels=values) +
+  theme(aspect.ratio=2/3, text=element_text(size=25))
+ggsave("position-coef.png")
+

### 2017-05-17 doc/models.tex
minor edits

@@
 
-%\subsection{Gated Recurrent Neural Networks}
-%\label{sec:gru}
+\subsection{Gated Recurrent Neural Networks}
+\label{sec:gru}
 
@@
 
-{\sc Imaginet} introduced in \cite{chrupala2015learning} is a
+{\sc Imaginet} introduced in \namecite{chrupala2015learning} is a
 multi-modal GRU network architecture that learns visually grounded

### 2017-05-17 doc/related.tex
minor edits

@@
 The direct predecessors of modern architectures were first proposed in the 
-seminal paper of \cite{elman1990finding}. He modifies the recurrent 
-neural network architecture of \cite{jordan1986attractor} by changing the output-to-memory 
+seminal paper of \namecite{elman1990finding}. He modifies the recurrent 
+neural network architecture of \namecite{jordan1986attractor} by changing the output-to-memory 
 feedback connections to hidden-to-memory recurrence, enabling Elman networks to 
-represent arbitrary dynamic systems. \namecite{elman1991distributed} he trains an RNN on a small synthetic
+represent arbitrary dynamic systems. \namecite{elman1991distributed} trains an RNN on a small synthetic
 sentence  dataset and analyzes the activation patterns of the hidden
@@
 encode lexical categories, grammatical relationships and hierarchical constituent 
-structures. \cite{giles1992extracting} trains RNNs similar to Elman
+structures. \namecite{giles1992extracting} trains RNNs similar to Elman
 networks on strings generated by small 
@@
 
-More closely related is the recent work of \cite{li2015visualizing}, who develop techniques 
+More closely related is the recent work of \namecite{li2015visualizing}, who develop techniques 
 for a deeper understanding of the activation patterns of RNNs, but focus on models with 

### 2017-05-17 doc/response.tex
added reviews at the end to add responses to later

@@
-\documentclass{article}
-\usepackage{a4wide}
-\usepackage{xr-hyper} 
-\usepackage{hyperref}
-\externaldocument{paper}
-\title{Response to the Reviewers}
-%\author{Ákos Kádár \and Grzegorz Chrupala \and Afra Alishahi}
-\author{Submission number: 882 \and
-Action editor:  Christopher Potts }
-\date{}
-\begin{document}
-
-\maketitle
-The Editor's and reviewers' comments were of great help and enabled us
-to make important improvements on our previous submission. Some minor or major
-changes were made to almost all of the sections of the paper.
-One of the major changes in the revised version is the addition of
-discussions on the generality and applicability of the proposed
-methods in sections \ref{sec:intro}, \ref{sec:computeomission}, 
-\ref{sec:beyondlexical} and \ref{sec:reprdim}. The revised version
-still focuses on the {\sc Imaginet} architecture, but in addition 
-allocates more space and more emphasis on the discussion
-of the introduced analysis techniques.
-The other major revision concerns the re-structuring 
-of Section \ref{sec:macro}. In the revised version we 
-separated the description of the omission score computation 
-(Section \ref{sec:computeomission}) and
-the discussion regarding the distribution of the omission scores over 
-categories of words for the different models 
-(Section \ref{sec:omitimaginet}). 
-Furthermore, in Section \ref{sec:beyondlexical}
-we introduced a common methodology
-to assess the impact of position and grammatical
-function of tokens (while controlling for word-types) based on the 
-omission scores. We also provide finer-grained analysis
-of these factors separately in sections \ref{sec:gramfunc} and
-\ref{subsec:information-struct}. Thanks to the reviewers' comments
-we realized that the original submission gave the impression that our main
-result was the amount of attention
-different models pay to words of different categories.  We hope that the
-revised version helps the reader to also focus on
-sections \ref{sec:beyondlexical}, \ref{sec:gramfunc} and
-\ref{subsec:information-struct} which explore linguistic features
-that the networks encode on top of simple lexical features. These and
-the additional less substantial revisions are discussed in more
-detail below.
+\pagenumbering{gobble}
+\newpage
 
 \begin{verbatim}
+We have reached a decision regarding your submission to Computational
+Linguistics, "Representation of linguistic form and function in
+recurrent neural networks". I include here the reviews for your
+perusal.
 
->> Editor's Comments <<
+The reviewers and I like the area of work in which this paper is
+situated: model analysis. While the reviews correctly point out that
+several shortcomings remain, a discussion has identified the
+modifications that we think are really required, to make the paper
+stronger. Other ideas are suggestions that we leave to you for future
+work.
 
-The paper is pitched in very broad terms (methods for analyzing RNNs), 
-but the methods are actually quite specific to classifier-type RNNs, 
-and many aspects of the work seem applicable only to the Imaginet 
-setting. I would suggest clarifying the scope of the claims. Similarly, 
-as Reviewer A points out, some of the rhetoric is misaligned with the 
-methods and results.  
+The requested changes concern an added comparisons to unimodal models
+(or at least to unimodal versions of Imaginet). On the other hand, all
+the reviewers thought that Section 6 is less convincing than the rest
+of the paper, at least in the current writeup. So we think that this
+Section could be reduced in favor of adding the unimodal experiments .
+
+I would be grateful if you could provide an appropriately revised
+version of your paper by 10th February 2017. Please email the
+editorial assistant (editorialoffice@cljournal.org) within one week,
+to confirm whether you will be able to meet this target date, or to
+negotiate a new target date.
+
+You will find details of the resubmission process at
+http://cljournal.org/ojs-help.html.  When you submit your revised
+paper, please enclose a letter that responds specifically to each of
+the points made by the reviewers, outlining the changes you have made,
+and those you have not addressed, together with an electronic copy of
+the revised manuscript as a PDF file.
+
+Thank you for submitting to Computational Linguistics.
+
+Paola Merlo
+University of Geneva
+Paola.Merlo@unige.ch
+------------------------------------------------------
+Reviewer A:
+
+2 What is this paper about? [Help] :
+The paper presents methodologies for analyzing what is learned by
+recurrent neural networks. It does this by suggesting two
+methodologies, and then applying them to two similar tasks -- one that
+that is based on language modeling, and another which is based on
+visual recognition. The analysis highlights some of the things that
+are encoded and not encoded in each of the representations.
+
+3 Strengths and Weaknesses [Help]
+
+Strengths:
+
+The topic of understanding what is learned by neural network models,
+in particular recurrent ones, is of great importance and interest, and
+papers of this kind are very much needed. While most work tries to use
+"deep learning" and RNNs to "improve" on various language tasks,
+understanding their capabilities and inner working, and making them
+more transparent, is much more challenging, and there are not enough
+works in this area. The current work attempts to do precisely that,
+and it also delivers on its promise with an interesting methodology
+and insights. In particular, the first part (section 5) is
+particularly well done. Section 6 is somewhat less convincing, but
+still is on-par or better than any other work that attempts to deliver
+analysis and insights.
+
+Weaknesses:
+
+Some parts of the papers were not clear, and could be improved. I also
+have someissues with some of the experiments in section 6, which can
+either be presented better or even removed altogether, in my view.
+
+- In terms of clarity, I would appreciate a better description of the
+  IMAGINET model, which is being analyzed. The model is described in
+  terms of its equations, but I would like to see a better
+  presentation of what is the intent behind it -
+
+- what is being modeled and what is being learned? this is available
+  in the original IMAGINET paper, but I think the description should
+  be imported to this paper also, in order to make it self contained
+  and stronger.
+
+- Also in terms of clarity, the graphs are not described in enough
+  detail in my view. Whiel the box plots (figure 2 and others) are
+  somewhat standard, they are not mainstream in the NLP commuity, and
+  I would appreciate seeing a more detailed description in the main
+  text or the caption of what each component means, both generally and
+  in the context of the specific graph.
+
+- Figure 7 is confusing -- what are the different colors between,
+  e.g., the words "in" and "tree" in the first line, and between "a"
+  and "wii" in the second? I assume this is some "gradient" between
+  two colors, but it looks like there are several data points and is
+  very confusing. Please fix this.
+
+In terms of content:
+
+- In section 6.1, why is the sorting by magnitude and not by absolute
+  value? are negative activations not important?
+
+- When choosing the "top k context of each unit" (section 6.1), it is
+  not clear what a "unit" is and how it is chosen.
+
+- I did not understand the last sentence in section 6.1, and would
+  appreciate an elaboration on this point (or a removal).
+
+- In section 6.2, it is not clear how trigrams with high activations
+  are chosen. How is a trigram with high activation defined? this is
+  not clear from the description of the method.
+
+- I am not convinced by the experiments in section 6.3 and their
+  importance -- either they are not described properly and I missed
+  something, or they are somewhat meaningless: when training a
+  logistic regression model to predict grammatical function, it is not
+  surprising that ones finds units that are predictive of grammatical
+  function.
+
+~ Where is section 6.4?
+
+~ a related (unpublished, but available on arxiv) work is
+https://arxiv.org/abs/1608.04207
+
+4 Substantive Revisions Required [Help]
+
+Complete this section if either 'Revise and Resubmit' or 'Reject' has
+been recommended.
+
+Revisions to be Required:
+:
+
+Revisions to be Encouraged:
+:
+
+5 Minor Revisions Required [Help]
+:
+
+See the points marked with (-) in the weaknesses above. Points marked
+with (~) are less important. For section 6.3, my recommendation is
+that it is either substantially revised, or removed altogether -- I
+believe the paper has enough merit without it.
+
+6 Typographic Errors [Help]
+:
+
+- "by as" in the first sentence of the intro.
+
+- There are some cases of using parenthesis in citation where they
+  shouldn't be, for example the first two citations in section 2 ("of
+  (Elman 1990)" should be "of Elman (1990)").
+
+------------------------------------------------------
+
+------------------------------------------------------
+Reviewer E:
+
+2 What is this paper about? [Help]
+:
+
+The paper presents a set of methods for analyzing the activations of
+recurrent neural networks from a linguistic perspective using:
+
+a) omission scores (to examine the contribution of a single token to
+the prediction of the network); and
+
+b) top k contexts (keeping the activations for the entire sequence and
+examining individual dimensions)
+
+The model used in this paper is a multimodal (text and visual)
+multi-task RNN (Imaginet). At each time step the next word is
+predicted from its current state, and an image vector is predicted
+from the model's final state. The model consists of two parallel GRUs
+with shared word embeddings.
+
+By pairing omission scores with POS and dependency relations, the
+authors show that the Visual pathway mostly focuses on noun and noun
+relations, while the textual pathway focuses more on function
+words. The textual pathway has in general a more uniform omission
+score distribution, the visual model peaks on content words (albeit
+not particularly on verbs, as the authors discuss).
+
+3 Strengths and Weaknesses [Help]
+
+Strengths:
+: 
+* neat idea and novel contribution (pairing activations with POS and
+dependency relations to gain further insights into RNNs)
+
+* extensive analysis; especially going beyond lexical cues (Section
+5.3), where the authors train a regression model to predict omission
+scores on a token level, showing that the network not just learns to
+rely on word types, but it also learns position information
+(particularly important for the textual pathway)
+
+Weaknesses:
+: 
+* the description of the top K contexts is not clear, how are top k
+contexts calculated? do you find the most active token (highest
+absolute value of hidden activation in the sequence?) and then use
+the words around it to find the trigrams? (what is meant by 'each
+unit' in Section6.1? Is the matrix actually R^nxd?); maybe you sort
+by column, not row?
+
+* Section 6.5 shows that the visual pathway focuses more on dependency
+relations, supporting the finding in the omission score part;
+however, the section is again very dense and Figure 7 only shows
+activations for Visual; I'd suggest to drop this part from the paper
+
+4 Substantive Revisions Required [Help]
+
+Complete this section if either 'Revise and Resubmit' or 'Reject' has
+been recommended.
+
+Revisions to be Required:
+: 
+- clearer description of how top k contexts are obtained
+
+Revisions to be Encouraged:
+: 
+- figures are not readable in BW printing; in addition to color, use
+  different symbols/line types
+- Related Work: clarify better how you differ from Li et al. (2015),
+  you mention "for a comparative analysis": I interpret this as "we
+  analyze a multi-task learning model (visual / textual) and compare
+  what the two parts learn"
+
+5 Minor Revisions Required [Help]
+: 
+- Sec 2, clarify: what are 'negative strings'? (non-words?)
+- Before equation 9 missing lambda ("weighted by \lambda"; what lambda
+  is used in the exp.?)
+- it would be helpful to include a figure of model (as in original
+  paper)
+- "two minor modifications" but the first seems you also use cosine,
+  was that different previously, not clear.
+- specify interaction features in regression model
+- 5.3.1. why also 'water'?
+
+6 Typographic Errors [Help]
+: 
+- abstract: sentence "Based on.." way too long
+- mention multi-task in abstract
+- introduction: "this includes various" > "This can be applied to
+  various"
+- from linguistic point (missing det)
+- generally avoid long sentences (e.g., last paragraph sec 5.1,
+  sec. 5.2)
+- rephrase: "this model we refer to as SUM"
+- windowsizes
+
+------------------------------------------------------
+
+------------------------------------------------------
+Reviewer F:
+
+2 What is this paper about? [Help]
+: 
+This paper presents an analysis of the information learned by
+recurrent neural networks. It focuses on one particular multimodal,
+multitask model, Imaginet, which is trained on images and their
+descriptions. The paper proposes a new measure, the omission score,
+which indicates how sensitive the network is to individual words in
+the input. Using this method, the authors show that the textual module
+of the model pays attention to function words, while the visual module
+pays attention to content words. The authors then propose a second
+method which can be used to analyze which n-grams the hidden units of
+the model are sensitive to; this results in meaningful syntactic and
+semantic clusters. This time, there is no clear-cut distinction
+between textual and visual modules.
+
+3 Strengths and Weaknesses [Help]
+
+Strengths:
+: 
+This paper makes a useful methodological contribution by proposing
+methods to analyze the knowledge learned by neural networks, which is
+an underexplored area, at least for networks trained on text data. The
+paper is well written, and technically sound.
+
+Weaknesses:
+: 
+The results of the paper are unfortunately fairly obvious. From the
+omission score analysis, we learn that a model trained to predict the
+next word in a sentence is sensitive to function words, while a word
+trained to associate a sentence with an image representation (which is
+essentially the output of the object detector) pays more attention to
+nouns. This is not at all surprising, it just reflects the different
+training objectives. The fact that the hidden units respond to
+specific n-grams is also expected; most people already believe that
+RNNs are good at approximating n-gram models; the question is, can
+they model more elaborate syntactic structures?
+
+This is really the question that the authors should explore, e.g., by
+testing whether the model can learn long distance dependencies,
+agreement, recursion, word order. An example for recent work that goes
+in this direction is Linzen et al. (2016).
+
+Another limitation of the paper is its focus on one particular model,
+which happens to be the authors'. To be convincing, the paper would
+really need to compare results for this model with results for other
+architectures, showing that the proposed analysis methods offer real
+discriminative power.
+
+Related to this, it is not clear to me why a multimodal, multitask
+model such as Imaginet is required for the work presented
+here. Surely, similar results would be obtained if two unimodal RNN
+models were trained, one on image descriptions to predict the next
+word, and one on image representations to predict the description.
+
+4 Substantive Revisions Required [Help]
+
+Complete this section if either 'Revise and Resubmit' or 'Reject' has
+been recommended.
+
+Revisions to be Required:
+: 
+(1) Add comparisons to unimodal models (or at least to unimodal
+versions of Imaginet).
+
+(2) Add comparisons to other architectures (LSTMs, etc.)
+
+(3) Sec. 5.3.2 should be cut. You don't have connected discourse, just
+isolated sentences. So it is purely speculative to say these
+observations are due to topic/content structure, rather than just to
+linear order.
+
+(4) The paper relies too much on impressionistic evaluation based on
+suggestive figures, rather than on quantitative analysis. An exception
+is Fig. 8. There should be more of this type of analysis.
+
+(5) The authors should discuss the recent paper by Linzen et
+al. (2016). This is a good example for an attempt to evaluate
+linguistic structure learned by an RNN that goes beyond individual
+words or n-grams.
+
+Reference
+
+Assessing the Ability of LSTMs to Learn Syntax-Sensitive
+Dependencies. Tal Linzen, Emmanuel Dupoux, Yoav Goldberg. To appear in
+TACL 2016. https://arxiv.org/abs/1611.01368v1
+
+Revisions to be Encouraged:
+:
+
+5 Minor Revisions Required [Help]
+:
+
+6 Typographic Errors [Help]
+:
+
+------------------------------------------------------
 \end{verbatim}
-
-This point was raised by both \emph{Reviewer A} and \emph{Reviewer B} and 
-had a major impact on our paper.
-Sections \ref{sec:intro}, \ref{sec:computeomission}, 
-\ref{sec:beyondlexical} and \ref{sec:reprdim} include new paragraphs
-aimed to make it clear in which settings the described methods could be meaningfully applied.
-
-\begin{itemize}
-  \item In Section \ref{sec:intro} ``Introduction'' on page
-    \pageref{generalintro}, we point out that we applied the introduced methods
-   only to {\sc Imaginet}, 
-  but that we also discuss their generalizability and offer pointers
-  on how to apply them in other settings.
-  \item On page \pageref{edit:introomissiongeneral} after briefly introducing the 
-   \emph{omission score} technique we point out that it can be applied to
-   other architectures too. The next paragraph on page
-   \pageref{edit:introtopkgeneral}, after introducing the
-   \emph{top-K-contexts} method itself, discusses its
-  applicability to other settings. 
-  \item We added a paragraph to the end of Section \ref{sec:computeomission} on
-  page \pageref{edit:omitgeneral} that points out in a bit more detail 
-  how the \emph{omission} technique can be generalized to other
-  RNN variants, recursive neural networks and CNNs. 
-  We hope that it will be clear to the reader that the analysis 
-  provided in Section \ref{sec:omitimaginet} can be applied to the 
-  same models as the \emph{omission} method.
-  \item We also added text in the first paragraph of Section \ref{sec:beyondlexical}
-  on page \pageref{edit:beyonlexicalgeneral} to clarify in which settings 
-  the described methods could be meaningfully applied.
-  \item In Section \ref{sec:reprdim} we explain that the provided 
-  analysis can be straight-forwardly applied to other RNN variants and 
-  suggest how to apply it to multi-layer architectures. We also point
-  out that it is likely to be limited to uni-directional models.
-\end{itemize}
-
-\begin{verbatim}
-The core results for content words and functions 
-words seem like the expected ones given their 
-linguistic content. In particular, the omission
-score method seems to show that the networks learn 
-to identify discriminating content words, and that
-changing or removing those words has a dramatic effect 
-on final predictions. This seems like a prerequisite 
-(given the linguistic facts) for learning an effective classifier. 
-Regular train/dev/test experimental analyses already
-show that these models are effective. Could you further 
-articulate what we learn via the omission score method 
-that we don't get from standard evaluations? 
-\end{verbatim}
-
-As mentioned above, the fact that the {\sc Visual} network learns to
-identify discriminating content words is only one of several results
-in our paper. Even this basic finding is not as obvious as it
-may seem: in the revised version, the last paragraph in Section \ref{sec:omitimaginet} on page 
-\pageref{edit:generality} discusses that the relatively low omission scores
-for verbs for the {\sc Visual} pathway are somewhat unexpected as one 
-would expect high omission
-scores for content words in general. We go on to discuss that the caption to
-image prediction setting 
-seems to lead to more task specific representations than expected.
-
-Furthermore, much of the paper is devoted to investigating patterns
-over and above attention paid to particular lexical items: for
-example in our experiment in Section~\ref{sec:macro} we focus on the
-effect of grammatical role and position in the sentence, {\it while
-  controling for word-type.} (See also the comment below).
-
-We believe these insights go well beyond what standard dev/test evaluations
-can tell us.
-
-\begin{verbatim}
-
-The connections with POS tags and dependency relations are intriguing.
-Could you, though, clarify the significance of these findings? My concern is
-that they just reflect the fact that the content-word/function-word
-distinction is highly correlated with POS and dependency relation.
-\end{verbatim}
-
-This comment really helped us restructure Section \ref{sec:macro}, with the aim
-to describe and highlight our main findings in a more clear manner. 
-Firstly in the new version of Section \ref{sec:omitimaginet} we do not emphasize so much
-the content- vs. function-word distinction and point out the relatively
-low omission of verbs in case of {\sc Visual}. Furthermore, in the new version
-we dedicate Section \ref{sec:beyondlexical} to raise the question about 
-the correlation between word-types, dependency relations and position in the 
-sentence and describe how we disentangle these factors using linear
-regression modeling. Concretely, we show
-that the omission scores both in case of {\sc Visual} and {\sc Textual} do not
-only depend on the word types, but on the grammatical function 
-and position of the tokens, while controling for word types.
-To further illustrate this point, we introduce a non-sequential baseline model {\sc Sum}, 
-in which we replace the RNN with a
-vector-summation operation over learned word-embeddings.
-
-\begin{verbatim}
-I had trouble understanding the information structure section. 
-I agree with the core insight about old/topic before new/comment 
-in languages like English. I had trouble relating this to what 
-I would expect from the RNN. For instance, the structure of the 
-RNN seems to ensure that information nearer the end of the sequence
-will have a larger impact. Is the Textual omission score result due 
-to this factor or is it truly due to something about language? 
-Conversely, for Visual, I guess I would have anticipated exactly 
-the opposite of what is reported here: the comment part should
-be the most important, corresponding to the new and visually 
-salient elements, rather than those that can fade into the background.
-\end{verbatim}
-
-Motivated by this and the previous comment in the revised version
-sections \ref{sec:gramfunc} and \ref{subsec:information-struct} use the same methodology
-for the analysis of the significance of grammatical function and position in the sentence.
-We hope this modification makes our methodology and the interpretation of our 
-results more straight-forward. Also for clarity, we streamlined the visualization
-of comparing how the position of tokens affect the omission scores of the different models
-- Figure \ref{fig:posrqs} on page \pageref{fig:posrqs}. 
-The current plot shows the same effect for {\sc Textual}
-and similar for {\sc Visual} as before, but hopefully in a more interpretable
-way. Furthermore, we have added a short discussion 
-on topic vs. comment in the image-description setting on page 
-\pageref{edit:topiccomment}, which we agree was definitely needed.
-
-\begin{verbatim}
-
->> Reviewer A <<
-
-One question about the omission score: the lower the score is, the low the
-similarity is. Thus, when dropping the important words, the similarity
-should go down -- so a lower omission score. It is not clear to me why
-dropping important words will lead to a higher omission score.
-
-\end{verbatim}
-
-Thank you for pointing this mistake out, we actually
-compute the cosine distance $1-cos(x,y)$ not cosine.
-We modified Equation \ref{eg:omit} on page \pageref{eg:omit} accordingly.
-
-\begin{verbatim}
-First, why is the architecture of
-Chrupala et al (2015) selected only? Is it representative enough for general
-RNNs? What about LSTM and its variants for language modelling task? How
-would they be different? It would be good to see such discussions. 
-\end{verbatim}
-
-Section \ref{sec:intro} (``Introduction'')
-discusses the reason why we choose this particular architecture
-on page \pageref{explainimaginet}. The model contains 
-two separate RNNs: a language model and a
-model that predicts image-vectors given sentences. 
-These ``pathways'' are trained on the same training data and 
-share word-embeddings. We choose such a setup as we
-are interested in 1) analyzing the hidden activation patterns 
-(rather than word embeddings), and 2) developing methods that allow for comparing models.
-As described in the first answer to the 
-Editor's first question, not discussing the generality and
-applicability of our methods was definitely a shortcoming of the paper and
-in the new version sections \ref{sec:computeomission},
-\ref{sec:beyondlexical}, \ref{sec:reprdim} discuss
-how general our techniques are and how to apply them in new settings.
-
-\begin{verbatim}
-Second, most of the paper presents qualitative results, without rigorous
-quantitative justifications. It would be nice, however, if the authors could
-collect some manual annotations and conduct automatic evaluations based on
-such annotations. For example, such annotations could include human's
-judgements on how important each word (or word class, syntactic class) in
-the sentence are etc.
-\end{verbatim}
-Regarding the collection of manual annotations, this is a great
-suggestion and we also have been planning 
-to assess the relationship between our results and human 
-performance. Conducting such experiments with 
-reliable results is labor- and resource-intensive and for the time 
-being we consider such an experiment to be out of the scope of our current
-paper, but as a great direction for our future work. 
-We discuss this idea in the new version of Section 
-\ref{conclusion} ``Conclusion''.
-
-Regarding qualitative results, we believe the paper offers a good
-balance of quantitative and qualitative analyses, even though our
-quantitative results do not take the form of accuracy scores. Our
-quantitative results include:
-\begin{itemize}
- \item We introduce omission scores in Section \ref{sec:computeomission}
- with the aim to quantify the importance of words to 
- the final prediction.
- \item Section \ref{sec:omitimaginet} compares {\sc Visual} and {\sc Textual} 
- with respect to the difference between their omission score distributions
- over POS tags and deprel categories.
- \item The new Section \ref{sec:beyondlexical} fits linear models with 
- various sets of predictor variables to predict omission score as outcome and
- assesses the contribution of these variables. 
- The analysis in sections \ref{sec:gramfunc}, 
- \ref{subsec:information-struct} also utilize these models.
- \item In Section \ref{seq:describe} we compare 
- {\sc Visual} and {\sc Textual}
- by measuring the strength of association between the 
- binned activation values
- of the hidden units and the token/deprel n-grams in the corpus with mutual
- information. 
-\end{itemize}
-
-\begin{verbatim}
-Finally, although the visualisation is helpful 
-and it might help the readers understand the 
-underlying activities of the RNN, little is known about why. 
-I understand it is just an empirical paper, however, it would 
-be helpful to at least share with the readers some high level 
-insights so that the readers could learn more from this paper. 
-One question that I have in mind after reading this paper is: 
-these results related to RNN presented here are expected. 
-It's good that RNN is able to capture such things. 
-However, how do I make use of the results here in my research? 
-How does this have some impact on my work?
-\end{verbatim}
-
-RNN variants proved to be powerful tools for modeling various aspects of human
-language and for NLP applications, and yet there has been little work on showing 
-what these models learn about language. In our paper we explore the 
-learned representations from a linguistic point of view, in the spirit
-of basic research.  
-Engineering applications are not the focus of this paper, and so in
-the body of the paper we try not to speculate on the relevance of our
-findings to practical NLP.
-
-That said, using our omission score results
-we do observe that the task of predicting images 
-from descriptions on MS-COCO does
-not promote the learning of representations for verbs. 
-Furthermore, one can observe a recency
-effect in the meaning representation of {\sc Textual}, 
-which probably also leads to too task specific representations. 
-Thus, we do hope that our methods can provide help for such
-engineering research in the future.
-
-\begin{verbatim}
-The authors claimed at the end of the paper that 
-"The goal of our paper is to propose novel methods
-for the encoding of linguistic knowledge in RNNs" -- I am not sure if I am
-convinced that this is goal of this work.  It is largely an empirical paper
-that tries to visualise/understand the RNN which was previously regarded as
-a blackbox, but does not really "propose" methods for encoding linguistic
-knowledge in RNN.
-\end{verbatim}
-
-This was an unfortunate mistake to leave this typo in the final version;
-we meant to write ``The goal of our paper is to propose novel methods
-for the ANALYSIS of the encoding of linguistic knowledge in RNNs''. We fixed it and 
-thank you for noticing it!
-
-\begin{verbatim}
- >> Reviewer B <<
-
-The main criticism I would have of this paper is that it 
-sets out as a providing generic new method for analysing RNNs 
-with text input, amongst other modalities. 
-I think this is overstating the contents of the paper: 
-the methods of analysis are fairly tied to the nature of the model
-(unidirectional RNN without attention) and broad task 
-(producing a final representation that can be used for classification or generation). 
-It would have been interesting to the same analyses performed on other, 
-related models (bi-directional RNN, word-based log-bilinear models, etc), 
-and on RNN tasks with other objectives (machine translation, language modelling, 
-sequence labelling). I realise it's a tall order to add this to the current work,
-but perhaps re-stating the scope of the work 
-(providing an example of how to perform further qualitative 
-and quantitative analysis of a an existing model trained
- on a specific task) 
- and briefly discussing how to adapt the experiments to other domains would be good.
-\end{verbatim}
-
-We really appreciated this comment! As described in our first answer, 
-this observation had a large impact on our paper. We did add 
-discussions in sections \ref{sec:intro}, \ref{sec:computeomission} 
-,\ref{sec:beyondlexical}, \ref{sec:reprdim} on how to apply our 
-introduced methods to different types of models.
-We hope the Reviewer agrees that the changes we made based on 
-this comment made our paper much stronger.
-
-\begin{verbatim}
-Likewise, a discussion comparing and contrasting these methods
-of analysis with other (complementary?) visualisation methods 
-for RNNs with attention such at Badhanau et al. 2015 or 
-Rocktaschel et al. 2015 [ICLR 2016] (which could be cited here) would be welcome.
-\end{verbatim}
-
-This was also a great suggestion and we did include a 
-discussion of the attention-weight 
-visualization on page \pageref{edit:attention} in 
-Section \ref{sec:related} based on the suggested works
-``Related Work''. This has made the section much more complete.
-
-\begin{verbatim}
-It would be nice to get a clear statement of the objective being optimised
-against in eq 9.
-\end{verbatim}
-
-The two loss functions are now described by equations \ref{eq:lossce}, and 
-\ref{eq:losscos} and the multi-task loss by Equation \ref{eq:losscombo}.
-
-\begin{verbatim}
-Figure 2 a little hard to read.
-\end{verbatim}
-
-We were considering ways to make the metioned figure less cluttered, but
-we are already reducing clutter by
-only including labels with frequency at least 500. We feel increasing
-this threshold further or merging categories would risk 
-removing too much information and end up with a misleading picture. 
-To help the understanding of the figure the main trends are explained in the text.
-
-\begin{verbatim}
-Figure 6 very hard to read.
-\end{verbatim}
-
-We agree that both the figure and the methodology in the previous
-version concerning the models' sensitivity to position in
-sentences was  not easy to interpret. Section 
-\ref{subsec:information-struct} reports new results 
-regarding the significance of the position of tokens along
-with a less cluttered Figure \ref{fig:posrqs}.
-
-\begin{verbatim}
-Section 5.1 "textual is more sensitive to ..." -> it looks like textual is
-more sensitive to most words being removed to some extent. I don't know if
-it's fair to single out these categories. Instead, just contrast more
-clearly with the heavy focus on content words (mainly NNs) for Visual.
-\end{verbatim}
-
-For clarity in the revised version of the paper we separated the
-introduction of the omission scores and the results based on them. 
-Section \ref{sec:computeomission} motivates omission scores and describes
-how to compute them and Section \ref{sec:omitimaginet} describes the
-distribution of the omission scores over POS-tags and deprels for 
-{\sc Textual} and {\sc Visual}. On page \ref{edit:textualomission} 
-in Section \ref{sec:omitimaginet} we point out that {\sc Visual} is 
-only sensitive to a couple of word-types, 
-while {\sc Textual} has a flatter omission score distribution in general.
-Only after this observation we go on to single out a couple of
-categories for a more detailed comparison.
-
-\begin{verbatim}
-Section 5.2, model 1 vs model 2 <- is the a significant difference in
-number of features here? Model 2 seems like it would have more features, and
-intuitively, more info == better model (modulo feature quality), so quick
-comment on this here would be good.
-\end{verbatim}
-
-We re-structured the sections analyzing the impact of position 
-and grammatical function. In the new version sections
-\ref{sec:gramfunc} and \ref{subsec:information-struct} 
-apply the same methodology to explore both factors. 
-Instead of commenting  on the difference in the number of features, in the new
-version for the comparison of {\sc Visual} and {\sc Textual} 
-we introduce a non-sequential baseline model - {\sc Sum}, and apply
-the same regression models with the same features to this baseline.
-Adding position and deprel variables in the linear model
-to predict omission scores does slightly increase the $R^2$ 
-even in case of {\sc Sum} - probably
-due to unseen words -, but much more so for {\sc Visual} and {\sc
-  Textual}.
-
-\begin{verbatim}
-Section 5.3 is very tied to structure of the model use in this paper
-(would expect different results for bidirection RNN, recursive neural net).
-Comment on this here would be good.
-\end{verbatim}
-
-The first paragraph in section Section \ref{sec:beyondlexical}
-on page \pageref{edit:beyonlexicalgeneral} argues that the analysis
-presented regarding the significance of word position are tied to RNNs. 
-We also added a footnote on page \pageref{edit:foot} 
-pointing out that recursive neural networks
-and CNNs are not affected by left-to-right and right-to-left word order
-in the way RNNs and bi-RNNs are. We agree that the results 
-would be probably different for bidirectional RNNs, but these 
-architectures are still affected by word-order.
-
-\begin{verbatim}
-Not a fan of qualitative analysis as done in Table 1/2. Too easy
-over-interpret, but I realise this is personal taste. Would not be offended
-if this is left in, but would suggest scrapping this to create space for
-discussions suggested above.
-\end{verbatim}
-
-Both in the Section \ref{sec:intro} and Section \ref{sec:reprdim}
-we point out that the provided results are based on qualitative methods.
-We agree that Tables \ref{tab:contexts}, \ref{tab:syntax} 
-and also Figure \ref{fig:dimheat} are easy 
-to over-interpret. This is arguably a danger with 
-qualitative analysis in general, but we do believe that it is interesting
-to complement our quantitative analysis with a detailed exploration
-of the function of individual hidden units from a closer look.
-
-\begin{verbatim}
- >> Reviewer C <<
-
-The paper builds heavily on previous work "IMAGINET", is densely described
-but lacking details in certain sections. For example, I find it rather
-challenging to understand the following:
-
-- the process of image vector prediction is vague (eqn 7). Does it mean
-prediction at the pixel level of an entire image, or a blob or principal
-image segment of interest? Also, how does the image retrieval shown in
-Figure 1 operate?
-
-\end{verbatim}
-
-To make these points more clear:
-
-\begin{itemize}
-	\item On page \pageref{edit:dumdumeddy} we 
-added a footnote ``Representing the full image, extracted from 
-a pre-trained Convolutional Neural Network''.
-	\item On page \pageref{edit:retrievalexplain} in Section 
-	\ref{sec:computeomission} we added an explanation about the retrieval
-	process.
-\end{itemize}
-
-\begin{verbatim}
-- Figure 6 would benefit from showing the aggregated POS groupings such as
-"VBN,VBG,VB" all grouped into "VERB" etc. for benefit of clarity in a small
-graph (too cluttered)
-\end{verbatim}
-
-Thank you for pointing it out, we hope that the reviewer agrees that 
-the new Figure \ref{fig:posrqs} is less cluttered and more intuitive.
-
-\end{document}

### 2017-05-19 src/depparse.py
Dependency parsing.

@@ -1,0 +1,45 @@
+import json
+import csv
+import spacy
+import en_core_web_md
+import logging
+
+coco_path = '/home/gchrupal/reimaginet/data/coco'
+data_path = '/home/gchrupal/cl-resubmit/data/'
+
+def main():
+    logging.getLogger().setLevel('INFO')
+    nlp = en_core_web_md.load()
+    split = 'val'
+    sent = sentences(json.load(open(coco_path + '/dataset.json')), split=split)
+    writer = csv.writer(open(data_path + '/depparse_coco_val.csv',"w"))
+    writer.writerow(["sentid", "position", "word", "postag", "dep", "head"])
+    for sent_i in sent:
+        logging.info("Parsing: {} {}".format(sent_i['sentid'], ' '.join(sent_i['tokens'])))
+        postag, label, head = parse(nlp, sent_i['tokens'])
+        for i in range(len(sent_i['tokens'])):
+            writer.writerow([sent_i['sentid'],
+                            i,
+                            sent_i['tokens'][i],
+                            postag[i],
+                            label[i],
+                            head[i]])
+
+
+def parse(nlp, tokens):
+    doc = spacy.tokens.Doc(nlp.vocab, words=tokens)
+    nlp.tagger(doc)
+    nlp.parser(doc)
+    head =  [word.head.i for word in doc]
+    label = [word.dep_ for word in doc]
+    postag = [word.tag_ for word in doc]
+    return postag, label, head
+
+def sentences(data, split='val'):
+    for image in data['images']:
+        if image['split'] == split:
+            for sentence in image['sentences']:
+                yield sentence
+
+if __name__ == '__main__':
+    main()

### 2017-05-19 src/omission.py
Omission stuff

@@ -1,0 +1,47 @@
+import json
+import csv
+import logging
+import imaginet.data_provider as dp
+import imaginet.defn.lm_visual as D
+moe
+from scipy.spatial.distance import cosine
+
+
+root = '/home/gchrupal/reimaginet/'
+data_path = '/home/gchrupal/cl-resubmit/data/'
+model_path = root + "/run-lm_visual-1/model.r.e6.zip"
+def main():
+    logging.getLogger().setLevel('INFO')
+    logging.info("Loading data")
+    prov = dp.getDataProvider(dataset='coco', root=root, audio_kind=None)
+    sent = list(prov.iterSentences(split='val'))
+    sent_id = [ sent_i['sentid'] for sent_i in sent]
+    sent_tok = [ sent_i['tokens'] for sent_i in sent]
+    logging.info("Loading imaginet model")
+    model = D.load(model_path)
+    writer = csv.writer(open(data_path + '/omission_coco_val.csv',"w"))
+    writer.writerow(["sentid", "position", "word", "omission_v","omission_t"])
+    for i in  range(len(sent)):
+        logging.info("Processing: {}".format(sent_id[i]))
+        O_v = omission(model, sent_tok[i], task=model.visual)
+        O_t = omission(model, sent_tok[i], task=model.lm)
+        for j in range(len(sent_tok[i])):
+            writer.writerow([sent_id[i],
+                            j,
+                            sent_tok[i][j],
+                            O_v[j],
+                            O_t[j]])
+
+def omission(model, toks, task=None):
+    if task is None:
+        task = model.visual
+    orig = task.states(model, [toks], task=model.visual)[0]
+    omit = task.states(model,
+                      [ toks[:i] + toks[i+1:] for i in range(len(toks))],
+                      task=model.visual)
+    return [ cosine(orig[-1], omit_i[-1]) for omit_i in omit]
+
+
+
+if __name__ == '__main__':
+    main()

### 2017-05-19 src/depparse.py
Depparse and omission.

@@ -13,17 +13,18 @@ def main():
     split = 'val'
     sent = sentences(json.load(open(coco_path + '/dataset.json')), split=split)
     writer = csv.writer(open(data_path + '/depparse_coco_val.csv',"w"))
-    writer.writerow(["sentid", "position", "word", "postag", "dep", "head"])
+    writer.writerow(["sentid", "position", "word", "postag", "postag_c", "dep", "head"])
     for sent_i in sent:
         logging.info("Parsing: {} {}".format(sent_i['sentid'], ' '.join(sent_i['tokens'])))
-        postag, label, head = parse(nlp, sent_i['tokens'])
+        postag, postag_c, label, head = parse(nlp, sent_i['tokens'])
         for i in range(len(sent_i['tokens'])):
             writer.writerow([sent_i['sentid'],
-                            i,
-                            sent_i['tokens'][i],
-                            postag[i],
-                            label[i],
-                            head[i]])
+                             i,
+                             sent_i['tokens'][i],
+                             postag[i],
+                             postag_c[i],
+                             label[i],
+                             head[i]])
 
 
 def parse(nlp, tokens):
@@ -33,7 +34,8 @@ def parse(nlp, tokens):
     head =  [word.head.i for word in doc]
     label = [word.dep_ for word in doc]
     postag = [word.tag_ for word in doc]
-    return postag, label, head
+    postag_c = [word.pos_ for word in doc ]
+    return postag, postag_c, label, head
 
 def sentences(data, split='val'):
     for image in data['images']:

### 2017-05-19 src/omission.py

@@ -2,14 +2,14 @@ import json
 import csv
 import logging
 import imaginet.data_provider as dp
-import imaginet.defn.lm_visual as D
-moe
+import imaginet.defn.lm_visual_vanilla as D
 from scipy.spatial.distance import cosine
-
+import imaginet.task
 
 root = '/home/gchrupal/reimaginet/'
 data_path = '/home/gchrupal/cl-resubmit/data/'
-model_path = root + "/run-lm_visual-1/model.r.e6.zip"
+model_path = root + "/run-lm_visual_vanilla-1/model.r.e7.zip"
+
 def main():
     logging.getLogger().setLevel('INFO')
     logging.info("Loading data")
@@ -35,10 +35,10 @@ def main():
 def omission(model, toks, task=None):
     if task is None:
         task = model.visual
-    orig = task.states(model, [toks], task=model.visual)[0]
-    omit = task.states(model,
+    orig = imaginet.task.states(model, [toks], task=task)[0]
+    omit = imaginet.task.states(model,
                       [ toks[:i] + toks[i+1:] for i in range(len(toks))],
-                      task=model.visual)
+                      task=task)
     return [ cosine(orig[-1], omit_i[-1]) for omit_i in omit]
 
 

### 2017-05-19 src/analysis.R
Viz of omission score distros.

@@ -1,0 +1,30 @@
+library(dplyr)
+library(ggplot2)
+
+dep <- read.csv("../data/depparse_coco_val.csv", header=TRUE)
+omi <- read.csv("../data/omission_coco_val.csv", header=TRUE)
+
+omi_v <- omi %>% select(-omission_t) %>% rename(score=omission_v) %>% mutate(pathway="visual")
+omi_t <- omi %>% select(-omission_v) %>% rename(score=omission_t) %>% mutate(pathway="textual")
+
+data_v <- merge(dep, omi_v) 
+data_t <- merge(dep, omi_t) 
+dep_lev <- (data_v %>% group_by(dep) %>% summarize(median_score=median(score)) %>% 
+              data.frame %>% arrange(median_score))$dep 
+
+
+data <- rbind(data_t, data_v) %>% mutate(dep = factor(dep, levels=dep_lev))
+
+filter_by_count <- function(dat, col, count) {
+  dat[dat[,col] %in% names(table(dat[,col]))[table(dat[,col]) > count],]
+}
+MINCOUNT<-500
+
+ggplot(filter_by_count(data, "dep", MINCOUNT)) +
+       aes(x=dep, y=score, color=pathway) +
+  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
+  coord_flip() +
+  theme(text=element_text(size=15))
+
+
+data_v %>% filter(dep=="attr")

### 2017-05-20 doc/categories.tex
Plots.

@@
   \begin{tabular}{cc}
-  \includegraphics[scale=0.55]{omission-stat/imaginet-omission-pos-boxplot.png} &
-  \includegraphics[scale=0.55]{omission-stat/imaginet-omission-deprel-boxplot.png} \\
+  \includegraphics[scale=0.55]{imaginet-omission-pos-boxplot.png} &
+  \includegraphics[scale=0.55]{imaginet-omission-dep-boxplot.png} \\
   \end{tabular}
 
-\caption{Distribution of omission scores for POS (left) and deprel labels
+\caption{Distribution of omission scores for POS (left) and dependency labels
   (right), for the {\sc Textual} and {\sc Visual} pathways. Only labels which occur at least 500 times are included.}
@@
   \begin{tabular}{cc}
-  \includegraphics[scale=0.55]{omission-stat/omission-pos-boxplot.png} &
-  \includegraphics[scale=0.55]{omission-stat/omission-deprel-boxplot.png} \\  
+  \includegraphics[scale=0.55]{imaginet-omission-ratio-pos-boxplot.png} &
+  \includegraphics[scale=0.55]{imaginet-omission-ratio-dep-boxplot.png} \\  
   \end{tabular}

### 2017-05-20 src/analysis.R
Analysis script.

@@ -11,20 +11,57 @@ data_v <- merge(dep, omi_v)
 data_t <- merge(dep, omi_t) 
 dep_lev <- (data_v %>% group_by(dep) %>% summarize(median_score=median(score)) %>% 
               data.frame %>% arrange(median_score))$dep 
+postag_lev <- (data_v %>% group_by(postag) %>% summarize(median_score=median(score)) %>%
+              data.frame %>% arrange(median_score))$postag
 
-
-data <- rbind(data_t, data_v) %>% mutate(dep = factor(dep, levels=dep_lev))
+data <- rbind(data_t, data_v) %>% 
+  mutate(dep = factor(dep, levels=dep_lev), 
+         postag = factor(postag, levels=postag_lev))
 
 filter_by_count <- function(dat, col, count) {
   dat[dat[,col] %in% names(table(dat[,col]))[table(dat[,col]) > count],]
 }
 MINCOUNT<-500
 
+
+# FIGURE 2
+ggplot(filter_by_count(data, "postag", MINCOUNT)) +
+  aes(x=postag, y=score, color=pathway) +
+  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
+  coord_flip() +
+  theme(text=element_text(size=15)) +
+  xlab("Part of Speech") +
+  ggsave(file="../doc/imaginet-omission-pos-boxplot.png",
+         width=5, height=5)
+   
+
 ggplot(filter_by_count(data, "dep", MINCOUNT)) +
        aes(x=dep, y=score, color=pathway) +
   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
   coord_flip() +
-  theme(text=element_text(size=15))
+  theme(text=element_text(size=15)) +
+  xlab("Dependency") +
+  ggsave(file="../doc/imaginet-omission-dep-boxplot.png",
+         width=5, height=5)
 
 
-data_v %>% filter(dep=="attr")
+# FIGURE 3
+ratio <- filter(data, pathway=='visual')$score / filter(data, pathway=='textual')$score
+data_r <- filter(data, pathway=="visual") %>% mutate(ratio=ratio, 
+                                                    postag=reorder(postag, ratio, median),
+                                                    dep=reorder(dep, ratio, median))
+ggplot(filter_by_count(data_r,"postag",MINCOUNT) %>% filter,
[…]

### 2017-05-20 src/ridge.py
Ridge models.

@@ -1,0 +1,81 @@
+from sklearn.linear_model import Ridge
+import pandas as pd
+import numpy
+import sys
+import logging
+
+from sklearn.feature_extraction import DictVectorizer
+from sklearn.linear_model import Ridge
+
+def choose(ds, keys):
+        return  ( dict([(key,d[key]) for key in keys ]) for d in ds )
+
+def dataset(data, score='score'):   
+    dicts = data.T.to_dict().values()
+    encoder = dict(word=DictVectorizer(), dep=DictVectorizer(), situation=DictVectorizer(), full=DictVectorizer())
+    X_word   = encoder['word'].fit_transform(choose(dicts, ['word']))
+    X_dep = encoder['dep'].fit_transform(choose(dicts, ['word', 'dep', 'word:dep']))
+    X_situation = encoder['situation'].fit_transform(choose(dicts, ['word','situation','word:situation']))
+    X_full = encoder['full'].fit_transform(choose(dicts, ['word','dep','word:dep','situation','word:situation']))
+    return dict(Y=numpy.array(data[score]), encoder=encoder, word=X_word, dep=X_dep, situation=X_situation, full=X_full)
+    
+def ridge_alphas(D, ix=100000):
+    Y = D['Y']
+    scores = dict(word=[], dep=[], situation=[], full=[])
+    alphas = [ 2**n for n in range(-2,6) ]
+    
+    def score(alpha, modeltyp):
+        model = Ridge(alpha=alpha)
+        model.fit(D[modeltyp][:ix,:], Y[:ix])
+        scores[modeltyp].append(model.score(D[modeltyp][ix:,:], Y[ix:]))
+        
+    for alpha in alphas:
+        for modeltyp in ['word','dep','situation','full']:
+            score(alpha, modeltyp)
+    return (alphas, scores)
+
+def predscore(kind, scores, out):
+        for key in scores.keys():
+            out.write("{}\t{}\t{}\n".format(kind, key, round(max(scores[key]),3)))
+            
+def situation(row):
+    if row['position'] == 0:
+        return "first"
+    elif row['position'] == 1:
+        return "second"
+    elif row['position'] == 2:
+        return "third"
+    elif row['position'] + 1 == row['length']:
+        return "last"
+    elif row['position'] + 1 == row['length'] - 1:
+        return "penult"
+    elif row['position'] + 1 == row['length'] - 2:
+        return "antepenult"
+    else:
+        return "middle"            
+    
[…]

### 2017-05-22 src/depparse.py
Added plain LM model and omission scores for it.

@@ -15,7 +15,8 @@ def main():
     writer = csv.writer(open(data_path + '/depparse_coco_val.csv',"w"))
     writer.writerow(["sentid", "position", "word", "postag", "postag_c", "dep", "head"])
     for sent_i in sent:
-        logging.info("Parsing: {} {}".format(sent_i['sentid'], ' '.join(sent_i['tokens'])))
+        #logging.info("Parsing: {} {}".format(sent_i['sentid'], ' '.join(sent_i['tokens'])))
+        logging.info("Parsing: {} {}".format(sent_i['sentid'], sent_i['raw']))
         postag, postag_c, label, head = parse(nlp, sent_i['tokens'])
         for i in range(len(sent_i['tokens'])):
             writer.writerow([sent_i['sentid'],

### 2017-05-22 src/omission.py

@@ -8,7 +8,8 @@ import imaginet.task
 
 root = '/home/gchrupal/reimaginet/'
 data_path = '/home/gchrupal/cl-resubmit/data/'
-model_path = root + "/run-lm_visual_vanilla-1/model.r.e7.zip"
+model_im_path = root + "/run-lm_visual_vanilla-1/model.r.e7.zip"
+model_lm_path = root + "/run-lm-0/model.r.e6.zip"
 
 def main():
     logging.getLogger().setLevel('INFO')
@@ -18,19 +19,23 @@ def main():
     sent_id = [ sent_i['sentid'] for sent_i in sent]
     sent_tok = [ sent_i['tokens'] for sent_i in sent]
     logging.info("Loading imaginet model")
-    model = D.load(model_path)
+    model_im = D.load(model_im_path)
+    logging.info("Loading plain LM model")
+    model_lm = imaginet.task.load(model_lm_path)
     writer = csv.writer(open(data_path + '/omission_coco_val.csv',"w"))
-    writer.writerow(["sentid", "position", "word", "omission_v","omission_t"])
+    writer.writerow(["sentid", "position", "word", "omission_v","omission_t","omission_lm"])
     for i in  range(len(sent)):
         logging.info("Processing: {}".format(sent_id[i]))
-        O_v = omission(model, sent_tok[i], task=model.visual)
-        O_t = omission(model, sent_tok[i], task=model.lm)
+        O_v = omission(model_im, sent_tok[i], task=model_im.visual)
+        O_t = omission(model_im, sent_tok[i], task=model_im.lm)
+        O_lm = omission(model_lm, sent_tok[i], task=model_lm.task)
         for j in range(len(sent_tok[i])):
             writer.writerow([sent_id[i],
                             j,
                             sent_tok[i][j],
                             O_v[j],
-                            O_t[j]])
+                            O_t[j],
+                            O_lm[j]])
 
 def omission(model, toks, task=None):
     if task is None:

### 2017-05-22 doc/categories.tex
Include LM omission scores.

@@
   \begin{tabular}{cc}
-  \includegraphics[scale=0.55]{imaginet-omission-pos-boxplot.png} &
-  \includegraphics[scale=0.55]{imaginet-omission-dep-boxplot.png} \\
+  \includegraphics[scale=0.6]{imaginet-omission-pos-boxplot.png} &
+  \includegraphics[scale=0.6]{imaginet-omission-dep-boxplot.png} \\
   \end{tabular}
@@
   \caption{Distributions of log ratios of omission scores of {\sc Textual} to {\sc Visual} per
+    POS (left) and deprel labels (right). Only labels which occur at least 500 times are included.}
+\label{fig:omission-imaginet-ratio}
+\end{figure*}
+
+\begin{figure*}[t]
+  \centering
+  \hspace*{-0.2in}
+  \setlength{\tabcolsep}{0pt}
+  \begin{tabular}{cc}
+  \includegraphics[scale=0.55]{imaginet-omission-quotient-pos-boxplot.png} &
+  \includegraphics[scale=0.55]{imaginet-omission-quotient-dep-boxplot.png} \\  
+  \end{tabular}
+  \caption{Distributions of log ratios of omission scores of {\sc LM} to {\sc Textual} per
     POS (left) and deprel labels (right). Only labels which occur at least 500 times are included.}

### 2017-05-22 src/analysis.R
Added vectorsum and LM to figure 5.

@@ -90,3 +90,18 @@ ggplot(filter_by_count(data_q,"dep",MINCOUNT) %>% filter,
   theme(text=element_text(size=15)) +
   xlab("Dependency")
 ggsave(file="../doc/imaginet-omission-quotient-dep-boxplot.png", width=5, height=5)
+
+data_lr <- read.table("../data/ridge_scores.txt", header=TRUE) 
+levels(data_lr$predictors)[levels(data_lr$predictors)=="situation"] <- "position"
+data_lr <- data_lr %>% mutate(predictors = reorder(predictors, R2), model=factor(model, levels=c("sum", "LM", "textual", "visual")))
+  
+for (model in levels(data_lr$model)) {
+  data_lr[data_lr$model==model,"score"] = data_lr[data_lr$model==model,"R2"]/data_lr[data_lr$model==model & data_lr$predictors=="word","R2"]
+  
+}
+ggplot(data_lr, aes(x=model, color=predictors, y=score)) + 
+  #geom_bar(stat="identity", position="dodge") +  
+  geom_boxplot() +
+  theme(aspect.ratio=2/3, text=element_text(size=25)) +
+  ylab("R Squared relative to word")
+ggsave("../doc/position-new.png")

### 2017-05-22 src/ridge.py

@@ -10,7 +10,7 @@ from sklearn.linear_model import Ridge
 def choose(ds, keys):
         return  ( dict([(key,d[key]) for key in keys ]) for d in ds )
 
-def dataset(data, score='score'):   
+def dataset(data, score='score'):
     dicts = data.T.to_dict().values()
     encoder = dict(word=DictVectorizer(), dep=DictVectorizer(), situation=DictVectorizer(), full=DictVectorizer())
     X_word   = encoder['word'].fit_transform(choose(dicts, ['word']))
@@ -18,17 +18,17 @@ def dataset(data, score='score'):
     X_situation = encoder['situation'].fit_transform(choose(dicts, ['word','situation','word:situation']))
     X_full = encoder['full'].fit_transform(choose(dicts, ['word','dep','word:dep','situation','word:situation']))
     return dict(Y=numpy.array(data[score]), encoder=encoder, word=X_word, dep=X_dep, situation=X_situation, full=X_full)
-    
+
 def ridge_alphas(D, ix=100000):
     Y = D['Y']
     scores = dict(word=[], dep=[], situation=[], full=[])
     alphas = [ 2**n for n in range(-2,6) ]
-    
+
     def score(alpha, modeltyp):
         model = Ridge(alpha=alpha)
         model.fit(D[modeltyp][:ix,:], Y[:ix])
         scores[modeltyp].append(model.score(D[modeltyp][ix:,:], Y[ix:]))
-        
+
     for alpha in alphas:
         for modeltyp in ['word','dep','situation','full']:
             score(alpha, modeltyp)
@@ -37,7 +37,7 @@ def ridge_alphas(D, ix=100000):
 def predscore(kind, scores, out):
         for key in scores.keys():
             out.write("{}\t{}\t{}\n".format(kind, key, round(max(scores[key]),3)))
-            
+
 def situation(row):
     if row['position'] == 0:
         return "first"
@@ -52,8 +52,8 @@ def situation(row):
     elif row['position'] + 1 == row['length'] - 2:
         return "antepenult"
     else:
-        return "middle"            
-    
+        return "middle"
+
 # Read and preprocess data
 def main():
     logging.getLogger().setLevel('INFO')
@@ -63,19 +63,29 @@ def main():
     data = pd.merge(depparse, omit)
 
     length = data.groupby("sentid")["position"].max()+1
-    data = data.merge(pd.DataFrame(dict(sentid=length.index, length=length.asobject)))
[…]

### 2017-05-22 src/omission.py
Additional models in omission.py and ridge.py

@@ -10,6 +10,7 @@ root = '/home/gchrupal/reimaginet/'
 data_path = '/home/gchrupal/cl-resubmit/data/'
 model_im_path = root + "/run-lm_visual_vanilla-1/model.r.e7.zip"
 model_lm_path = root + "/run-lm-0/model.r.e6.zip"
+model_sum_path = root + "/run-vectorsum-0/model.r.e7.zip"
 
 def main():
     logging.getLogger().setLevel('INFO')
@@ -22,20 +23,28 @@ def main():
     model_im = D.load(model_im_path)
     logging.info("Loading plain LM model")
     model_lm = imaginet.task.load(model_lm_path)
+    logging.info("Loading vectorsum model")
+    model_sum = imaginet.task.load(model_sum_path)
     writer = csv.writer(open(data_path + '/omission_coco_val.csv',"w"))
-    writer.writerow(["sentid", "position", "word", "omission_v","omission_t","omission_lm"])
+    writer.writerow(["sentid", "position", "word",
+                     "omission_v",
+                     "omission_t",
+                     "omission_lm",
+                     "omission_sum"])
     for i in  range(len(sent)):
         logging.info("Processing: {}".format(sent_id[i]))
         O_v = omission(model_im, sent_tok[i], task=model_im.visual)
         O_t = omission(model_im, sent_tok[i], task=model_im.lm)
         O_lm = omission(model_lm, sent_tok[i], task=model_lm.task)
+        O_sum = omission(model_sum, sent_tok[i], task=model_sum.task)
         for j in range(len(sent_tok[i])):
             writer.writerow([sent_id[i],
                             j,
                             sent_tok[i][j],
                             O_v[j],
                             O_t[j],
-                            O_lm[j]])
+                            O_lm[j],
+                            O_sum[j]])
 
 def omission(model, toks, task=None):
     if task is None:

### 2017-05-22 src/ridge.py

@@ -10,7 +10,7 @@ from sklearn.linear_model import Ridge
 def choose(ds, keys):
         return  ( dict([(key,d[key]) for key in keys ]) for d in ds )
 
-def dataset(data, score='score'):   
+def dataset(data, score='score'):
     dicts = data.T.to_dict().values()
     encoder = dict(word=DictVectorizer(), dep=DictVectorizer(), situation=DictVectorizer(), full=DictVectorizer())
     X_word   = encoder['word'].fit_transform(choose(dicts, ['word']))
@@ -18,17 +18,17 @@ def dataset(data, score='score'):
     X_situation = encoder['situation'].fit_transform(choose(dicts, ['word','situation','word:situation']))
     X_full = encoder['full'].fit_transform(choose(dicts, ['word','dep','word:dep','situation','word:situation']))
     return dict(Y=numpy.array(data[score]), encoder=encoder, word=X_word, dep=X_dep, situation=X_situation, full=X_full)
-    
+
 def ridge_alphas(D, ix=100000):
     Y = D['Y']
     scores = dict(word=[], dep=[], situation=[], full=[])
     alphas = [ 2**n for n in range(-2,6) ]
-    
+
     def score(alpha, modeltyp):
         model = Ridge(alpha=alpha)
         model.fit(D[modeltyp][:ix,:], Y[:ix])
         scores[modeltyp].append(model.score(D[modeltyp][ix:,:], Y[ix:]))
-        
+
     for alpha in alphas:
         for modeltyp in ['word','dep','situation','full']:
             score(alpha, modeltyp)
@@ -37,7 +37,7 @@ def ridge_alphas(D, ix=100000):
 def predscore(kind, scores, out):
         for key in scores.keys():
             out.write("{}\t{}\t{}\n".format(kind, key, round(max(scores[key]),3)))
-            
+
 def situation(row):
     if row['position'] == 0:
         return "first"
@@ -52,8 +52,8 @@ def situation(row):
     elif row['position'] + 1 == row['length'] - 2:
         return "antepenult"
     else:
-        return "middle"            
-    
+        return "middle"
+
 # Read and preprocess data
 def main():
     logging.getLogger().setLevel('INFO')
@@ -63,19 +63,21 @@ def main():
     data = pd.merge(depparse, omit)
 
     length = data.groupby("sentid")["position"].max()+1
-    data = data.merge(pd.DataFrame(dict(sentid=length.index, length=length.asobject)))
[…]

### 2017-05-22 doc/categories.tex

@@
   \caption{Distributions of log ratios of omission scores of {\sc Textual} to {\sc Visual} per
-    POS (left) and deprel labels (right). Only labels which occur at least 500 times are included.}
+    POS (left) and dependency labels (right). Only labels which occur at least 500 times are included.}
 \label{fig:omission-imaginet-ratio}
@@
   \caption{Distributions of log ratios of omission scores of {\sc LM} to {\sc Textual} per
-    POS (left) and deprel labels (right). Only labels which occur at least 500 times are included.}
+    POS (left) and dependency labels (right). Only labels which occur at least 500 times are included.}
 \label{fig:omission-imaginet-ratio}
@@
 
-We fit four linear models which predict the omission 
+We fit four L2-penalized linear regression models which predict the omission 
 scores per token with the following predictor variables:
@@
 \begin{enumerate}
-	\item {\sc LM word}: word type
-	\item {\sc LM deprel}: word type, dependency label and their interaction 
-	\item {\sc LM position}: word type, position (binned as first, second, third, middle,
+	\item {\sc LR word}: word type
+	\item {\sc LR deprel}: word type, dependency label and their interaction 
+	\item {\sc LR position}: word type, position (binned as first, second, third, middle,
 	antepenult, penult, last) and their interaction
-	\item {\sc LM full}: word type, dependency label, position, word-deprel interaction, 
-	word-position interaction
+	\item {\sc LR full}: word type, dependency label, position, word:dependency interaction, 
+	word:position interaction
 \end{enumerate}
@@
 over word-embeddings and thus does not have access to sequential cues. This
-model we refer to as {\sc Sum}.
+mposodel we refer to as {\sc Sum}.
 
 Figure~\ref{fig:rsquared} shows the increase in $R^2$  for the models 
-per pathway relative to {\sc LM word}. The plot reveals that 
+per pathway relative to {\sc LR word}. The plot reveals that 
 adding additional information on top of lexical features in case of {\sc Sum} 
@@
 the unseen words in the held out set. More interestingly, the
-relative increase in $R^2$ between {\sc LM word} and {\sc LM full} in
+relative increase in $R^2$ between {\sc LR word} and {\sc LR full} in
 case of {\sc Visual} is larger, suggesting that 
@@
   \includegraphics[scale=0.35]{omission-stat/position-new.png}
-\caption{Proportion of variance in omission scores explained by the models
- for {\sc Sum}, {\sc Visual} and {\sc Textual} pathways.}
+\caption{Proportion of variance in omission scores explained by the
+  linear regression models
+ for {\sc Sum}, {\sc Visual} and {\sc Textual} pathways, relative to
+ regressing on word identity only. }
 \label{fig:rsquared}

### 2017-05-22 src/analysis.R
Analysis.

@@ -93,13 +93,14 @@ ggsave(file="../doc/imaginet-omission-quotient-dep-boxplot.png", width=5, height
 
 data_lr <- read.table("../data/ridge_scores.txt", header=TRUE) 
 levels(data_lr$predictors)[levels(data_lr$predictors)=="situation"] <- "position"
-data_lr <- data_lr %>% mutate(predictors = reorder(predictors, R2), model=factor(model, levels=c("sum", "LM", "textual", "visual")))
+data_lr <- data_lr %>% mutate(predictors = factor(predictors, levels=c("word","position","dep","full")), model=factor(model, levels=c("sum", "LM", "textual", "visual")))
   
 for (model in levels(data_lr$model)) {
   data_lr[data_lr$model==model,"score"] = data_lr[data_lr$model==model,"R2"]/data_lr[data_lr$model==model & data_lr$predictors=="word","R2"]
   
 }
-ggplot(data_lr, aes(x=model, color=predictors, y=score)) + 
+ggplot(data_lr,#  %>% filter(model %in% c("sum","visual")), 
+       aes(x=model, color=predictors, y=score)) + 
   #geom_bar(stat="identity", position="dodge") +  
   geom_boxplot() +
   theme(aspect.ratio=2/3, text=element_text(size=25)) +

### 2017-05-22 doc/categories.tex
Ridge plot.

@@
 	\item {\sc LR word}: word type
-	\item {\sc LR deprel}: word type, dependency label and their interaction 
-	\item {\sc LR position}: word type, position (binned as first, second, third, middle,
-	antepenult, penult, last) and their interaction
+	\item {\sc LR dep}: word type, dependency label and their interaction 
+	\item {\sc LR position}: word type, position (binned as {\sc first, second, third, middle,
+	antepenult, penult, last}) and their interaction
 	\item {\sc LR full}: word type, dependency label, position, word:dependency interaction, 
@@
 \centering
-  \includegraphics[scale=0.35]{omission-stat/position-new.png}
+  \includegraphics[scale=0.35]{position-new.png}
 \caption{Proportion of variance in omission scores explained by the
@@
  for {\sc Sum}, {\sc Visual} and {\sc Textual} pathways, relative to
- regressing on word identity only. }
+ regressing on word identity and position only. }
 \label{fig:rsquared}

### 2017-05-22 src/analysis.R

@@ -96,7 +96,7 @@ levels(data_lr$predictors)[levels(data_lr$predictors)=="situation"] <- "position
 data_lr <- data_lr %>% mutate(predictors = factor(predictors, levels=c("word","position","dep","full")), model=factor(model, levels=c("sum", "LM", "textual", "visual")))
   
 for (model in levels(data_lr$model)) {
-  data_lr[data_lr$model==model,"score"] = data_lr[data_lr$model==model,"R2"]/data_lr[data_lr$model==model & data_lr$predictors=="word","R2"]
+  data_lr[data_lr$model==model,"score"] = data_lr[data_lr$model==model,"R2"]/data_lr[data_lr$model==model & data_lr$predictors=="position","R2"]
   
 }
 ggplot(data_lr,#  %>% filter(model %in% c("sum","visual")), 
@@ -104,5 +104,5 @@ ggplot(data_lr,#  %>% filter(model %in% c("sum","visual")),
   #geom_bar(stat="identity", position="dodge") +  
   geom_boxplot() +
   theme(aspect.ratio=2/3, text=element_text(size=25)) +
-  ylab("R Squared relative to word")
+  ylab('R Squared relative to Position')
 ggsave("../doc/position-new.png")

### 2017-05-23 doc/models.tex
expanding the description of Imaginet

@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.25]{imaginet.pdf} 
+\caption{Structure of {\sc Imaginet}, taken from Chrupała et al. (2015).}
+\label{fig:imaginet}
+\end{center}
+\end{figure}
+
 {\sc Imaginet} introduced in \namecite{chrupala2015learning} is a
 multi-modal GRU network architecture that learns visually grounded
-meaning representations from textual and visual input. It consists of
-two GRU pathways, {\sc Textual} and {\sc Visual}, with a shared
-word-embedding matrix. The inputs to the model are pairs of image descriptions 
-and their corresponding images. Each sentence is mapped to two
-sequences of hidden states, one by {\sc Visual} and the other by {\sc
-  Textual}:
+meaning representations from textual and visual input.  It acquires 
+linguistic knowledge via comprehension and production, by receiving a
+description of a scene and trying to visualise it through predicting a visual
+representation for the textual description, while concurrently predicting 
+the next word in the sequence.
+
+Figure~\ref{fig:imaginet} shows the structure of {\sc Imaginet}. As can
+be seen from the figure, the model consists of two GRU pathways, 
+{\sc Textual} and {\sc Visual}, with a shared word-embedding matrix. 
+The inputs to the model are pairs of image descriptions and their 
+corresponding images. The {\sc Textrual} pathway predicts the next 
+word at each position in the sequence of words in each caption, whereas the 
+{\sc Visual} pathway predicts a visual representation of the image that depicts the 
+scene described by the caption after the final word is received.
+
+Formally, each sentence is mapped to two sequences of hidden states, 
+one by {\sc Visual} and the other by {\sc Textual}:
 \vspace{-.2cm}
@@
 %
-\vspace{-.2cm}
-
-\begin{equation}
-\label{eq:lossce}
- L^{T}(\theta) = {-} \frac{1}{\tau}\sum_{t=1}^\tau \log p(S_t|S_{1:t})
-\end{equation}
-
-\begin{equation}
-\label{eq:losscos}
- L^V(\theta) =  1 - \frac{\hat{\mathbf{i}} \cdot \mathbf{i}}{\|
-     \hat{\mathbf{i}} \| \| \mathbf{i} \|}
-\end{equation}
-
-\begin{equation}
-\label{eq:losscombo}
-     L = \alpha L^T + (1-\alpha)L^{V}
-\end{equation}
+%\vspace{-.2cm}
+\begin{align}
+%\begin{equation}
+%\label{eq:lossce}
+&L^{T}(\theta) = {-} \frac{1}{\tau}\sum_{t=1}^\tau \log p(S_t|S_{1:t}) \\
+%\end{equation}
+%
+%\vspace{-.2cm}
+%\begin{equation}
+%\label{eq:losscos}
+&L^V(\theta) =  1 - \frac{\hat{\mathbf{i}} \cdot \mathbf{i}}{\| \hat{\mathbf{i}} \| \| \mathbf{i} \|} \\
+%\end{equation}
+%
+%\vspace{-.2cm}
+%\begin{equation}
+%\label{eq:losscombo}
+&L = \alpha L^T + (1-\alpha)L^{V}
+\end{align}
 %L^V(\theta) &= \frac{1}{K}\sum_{k=1}^K(\hat{i}_k-i_k)^2
@@
 
+\noindent
 For more details about the {\sc Imaginet} model and 

### 2017-05-23 doc/response.tex
expanding the description of Imaginet
adding place holders in the response letter, fixing minor issues.

@@
-\pagenumbering{gobble}
 \newpage
+\pagenumbering{roman}
+
+\noindent
+Dear Dr.~Merlo,
+\newline
+
+Thank you for the very helpful suggestions and insightful questions
+posed in this review.  We apologize for the delay getting this
+revision back to you and very much appreciate the extension you 
+gave us to resubmit the manuscript.
+
+We have revised parts of the article and have made clarifications
+throughout the text in response to yours and the reviewers' comments.
+Specifically with regard to the two main issues you raised, we have...
+\todo[inline]{explain the major revisions}
+
+We believe that the revisions to the paper address your collective
+concerns, and we thank you all again for the comments that have helped
+us to improve the paper.
+\newline
+
+\noindent
+Sincerely,
+\newline
+
+\noindent
+Ákos Kádár, Grzegorz Chrupała and Afra Alishahi
+\newline
 
 \begin{verbatim}
+Dear Ákos Kádár:
+
 We have reached a decision regarding your submission to Computational
@@
 
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}
+We have expanded Section~\ref{sec:imaginet} with a conceptual
+description of {\sc Imaginet} and its learning goals, and added a graphical
+representation of the structure of the model for more clarity (Figure~\ref{fig:imaginet}).
+\end{quote}
+\begin{verbatim}
+
 - Also in terms of clarity, the graphs are not described in enough
@@
   in the context of the specific graph.
-
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{explain the updated graph}
+\end{quote}
+\begin{verbatim}
 - Figure 7 is confusing -- what are the different colors between,
@@
   very confusing. Please fix this.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{this figure should be removed.}
+\end{quote}
+\begin{verbatim}
 
@@
 
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:} probably reorganise this section \todo[inline]{explain changes}
+\end{quote}
+\begin{verbatim}
+
 ~ a related (unpublished, but available on arxiv) work is
 https://arxiv.org/abs/1608.04207
+
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{add to related work}
+\end{quote}
+\begin{verbatim}
 
@@
 
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{explain reorganisations and major changes.}
+\end{quote}
+\begin{verbatim}
+
 6 Typographic Errors [Help]
@@
   (Elman 1990)" should be "of Elman (1990)").
+
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:} the typo and the citation inconsistencies are fixed.
+\end{quote}
+\begin{verbatim}
 
@@
 
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{explain restructuring of section 6}
+\end{quote}
+\begin{verbatim}
+
 * Section 6.5 shows that the visual pathway focuses more on dependency
@@
 activations for Visual; I'd suggest to drop this part from the paper
+
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{this section should probably be removed}
+\end{quote}
+\begin{verbatim}
 
@@
 - clearer description of how top k contexts are obtained
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{explain}
+\end{quote}
+\begin{verbatim}
 
@@
   different symbols/line types
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{update graphs?}
+\end{quote}
+\begin{verbatim}
+
 - Related Work: clarify better how you differ from Li et al. (2015),
@@
   what the two parts learn"
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{add to related work}
+\end{quote}
+\begin{verbatim}
 
@@
 - Sec 2, clarify: what are 'negative strings'? (non-words?)
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  we have reworded `positive' and `negative' as `grammatical' and `ungrammatical' to avoid confusion.
+\end{quote}
+\begin{verbatim}
 - Before equation 9 missing lambda ("weighted by \lambda"; what lambda
   is used in the exp.?)
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{good point! what is the value of lambda?}
+\end{quote}
+\begin{verbatim}
 - it would be helpful to include a figure of model (as in original
   paper)
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  We have included an illustration of the structure of the model in Figure~\ref{fig:imaginet}.
+\end{quote}
+\begin{verbatim}
 - "two minor modifications" but the first seems you also use cosine,
   was that different previously, not clear.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{Grzegorz?}
+\end{quote}
+\begin{verbatim}
 - specify interaction features in regression model
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{specify interaction features}
+\end{quote}
+\begin{verbatim}
 - 5.3.1. why also 'water'?
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{explain}
+\end{quote}
+\begin{verbatim}
 
@@
 - windowsizes
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  we have fixed these issues, thanks for bringing them to our attention.
+\end{quote}
+\begin{verbatim}
 
@@
 they model more elaborate syntactic structures?
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{address}
+\end{quote}
+\begin{verbatim}
 
@@
 in this direction is Linzen et al. (2016).
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{discuss linzen et al. in related work}
+\end{quote}
+\begin{verbatim}
 
@@
 discriminative power.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{address}
+\end{quote}
+\begin{verbatim}
 
@@
 word, and one on image representations to predict the description.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{address}
+\end{quote}
+\begin{verbatim}
 
@@
 versions of Imaginet).
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{explain new experiments}
+\end{quote}
+\begin{verbatim}
 
 (2) Add comparisons to other architectures (LSTMs, etc.)
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{hmm.}
+\end{quote}
+\begin{verbatim}
 
@@
 linear order.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{explain the rewritten version.}
+\end{quote}
+\begin{verbatim}
 
@@
 is Fig. 8. There should be more of this type of analysis.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{address.}
+\end{quote}
+\begin{verbatim}
 
@@
 words or n-grams.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  \todo[inline]{add this to related work.}
+\end{quote}
+\begin{verbatim}
 

### 2017-05-23 doc/categories.tex
ridge
adding place holders in the response letter, fixing minor issues.

@@
 For other RNN architectures such as GRUs, LSTMs \label{edit:omitgeneral}
-and their bi-directional variants to measure the contribution
+and their bi-directional variants, measuring the contribution
 of tokens to their predictions the $\mathrm{omission}$ 
@@
 at the last time step used for prediction. Furthermore, the technique 
-can be applied in general to other architectures, which
+can be applied in general to other architectures which
 embed variable length linguistic expressions to the same fixed dimensional
@@
 for Recurrent Neural Networks as they compute their representations sequentially and are not
-limited by fixed windowsizes.\footnote{CNNs with multi-word filters
+limited by fixed window sizes.\footnote{CNNs with multi-word filters
 and tree-structured recursive neural networks do not incrementaly build representations
@@
 where the GRU in the {\sc Visual} pathway is replaced by a simple vector-summation operation
-over word-embeddings and thus does not have access to sequential cues. This
-mposodel we refer to as {\sc Sum}.
+over word-embeddings and thus does not have access to sequential cues. We refer to this
+model as {\sc Sum}.
 
 Figure~\ref{fig:rsquared} shows the increase in $R^2$  for the models 
-per pathway relative to {\sc LR word}. The plot reveals that 
+per pathway relative to {\sc LR position}. The plot reveals that 
 adding additional information on top of lexical features in case of {\sc Sum} 

### 2017-05-23 doc/intro.tex
minor edits
adding place holders in the response letter, fixing minor issues.

@@
   sentence, and predicting the next word in the same sentence. 
-  Based on our proposed method to estimate the amount of contribution
-  of individual tokens in the input to the final prediction of the networks   
-  we show that the image prediction pathway: a) is sensitive to
+  We propose a method for estimating the amount of contribution
+  of individual tokens in the input to the final prediction of the networks.   
+  We show that the image prediction pathway: a) is sensitive to
   the information structure of the sentence, b) pays
@@
 Recurrent neural networks (RNNs) were introduced by
-%\cite{elman1990finding} 
+\namecite{elman1990finding} 
 as a connectionist architecture with the
@@
 As our case study we picked the {\sc Imaginet} model introduced by \label{explainimaginet}
-\cite{chrupala2015learning}. It is a multi-task, multi-modal
+\namecite{chrupala2015learning}. It is a multi-task, multi-modal
 architecture consisting of two Gated-Recurrent Unit (GRU)
@@
 to the final prediction of models that compute distributed representations 
-of sentences. This includes various RNN architectures, Recursive Neural Networks 
+of sentences. This can be applied to various RNN architectures, Recursive Neural Networks 
 and Convolutional Neural Networks.\label{edit:introomissiongeneral}

### 2017-05-23 doc/related.tex
adding place holders in the response letter, fixing minor issues.

@@
 networks on strings generated by small 
-deterministic regular grammars with the objective to recognize positive and reject negative 
+deterministic regular grammars with the objective to recognize grammatical and reject ungrammatical 
 strings, and develops the \emph{dynamic state partitioning} technique to extract the learned 

### 2017-05-23 doc/categories.tex
More fussing with fig:rsquared and around.

@@
 	\item {\sc LR word}: word type
-	\item {\sc LR dep}: word type, dependency label and their interaction 
-	\item {\sc LR position}: word type, position (binned as {\sc first, second, third, middle,
+	\item {\sc LR +dep}: word type, dependency label and their interaction 
+	\item {\sc LR +pos}: word type, position (binned as {\sc first, second, third, middle,
 	antepenult, penult, last}) and their interaction
@@
 
-\noindent We split the whole MSCOCO validation set into two parts, fit the 
+\noindent We split the 5-thousand-image portion of MSCOCO validation data into two parts, fit the 
 models (regularized via L2 penalty) on the first part, and compute the 
@@
 
-Figure~\ref{fig:rsquared} shows the increase in $R^2$  for the models 
-per pathway relative to {\sc LR position}. The plot reveals that 
-adding additional information on top of lexical features in case of {\sc Sum} 
-does increase the explained variance slightly, which is probably due to 
-the unseen words in the held out set. More interestingly, the
-relative increase in $R^2$ between {\sc LR word} and {\sc LR full} in
-case of {\sc Visual} is larger, suggesting that 
-the omission scores in case of {\sc Visual} do not only depend on the words themselves,
-but their grammatical function and position in sentences.  The raw  $R^2$
-scores show that for the {\sc Textual} model the word-type predicts the omission-score
-to much smaller degree compared to {\sc Visual} (0.38 against 0.48), and the 
-increase in $R^2$ shows that the position of the word adds considerable 
-amount of information.
+\begin{table}
+  \centering
+  \caption{Proportion of variance in omission scores explained by
+    linear regression.}
+    \begin{tabular}{l|rrrr}
+               & word   & +pos  & +dep  & full \\\hline
+     sum       & 0.654  & 0.661 & 0.670 & 0.670 \\
+     LM        & 0.358  & 0.586 & 0.415 & 0.601 \\
+     textual   & 0.364  & 0.703 & 0.451 & 0.715 \\
+     visual    & 0.490  & 0.506 & 0.515 & 0.523 \\
+    \end{tabular}
+    \label{tab:lr-r2}
+\end{table}
+
+Table~\ref{tab:lr-r2} shows the proportion of variance $R^2$ in omission
+scores explained by the linear regression with the different predictors.
+Figure~\ref{fig:rsquared} offers a different view of the data, showing
+the increase or decrease in $R^2$ for the models per pathway relative to {\sc LR
+  +pos}.  The results reveal that adding additional information on
+top of lexical features in case of {\sc Sum} does increase the
+explained variance slightly, which is probably due to the unseen words
+in the held out set. More interestingly, there is a sizeable
+increase in $R^2$ between {\sc LR +pos} and {\sc LR full} in
+case of {\sc Visual}, suggesting that 
+the omission scores in case of {\sc Visual} depend on the words'
+grammatical function in sentences, even after controling for word
+identity and linear position.
+
+The raw $R^2$ scores show that for the {\sc Textual} and {\sc LM}
+models the word-type predicts the omission-score to much smaller
+degree compared to {\sc Visual}, and the increase in $R^2$ shows that
+the position of the word adds considerable amount of information for
+these language models: this is not surprising considering that the omission
+scores are measured with respect to the final activation state. 
+Less predictably, for the {\sc Visual} model,
+dependency labels are more informative than linear position,
+hinting at the importance of syntactic structure beyond linear order.
+
+Overall, when regressing on word identities, word position and
+dependency labels, the {\sc Visual} model's omission scores are the
+hardest to predict of the four models suggesting that the model may be
+encoding additional structural features not captured by these
+predictors.
 
@@
 
-In order to find out some of the specific configurations leading to
-increase in $R^2$ in case of {\sc Visual}, we next considered all word types with
+In order to find out some of the specific syntactic configurations leading to
+increase in $R^2$ between {\sc LR word} and {\sc LR dep} predictors
+in case of {\sc Visual}, we next considered all word types with
 occurrence counts of at least 100 and ranked them according to how much
-better, on average, {\sc LM deprel} predicted their omission scores 
-compared to {\sc LM word}. Figure~\ref{fig:top_words} shows the per-dependency 
+better, on average, {\sc LR dep} predicted their omission scores 
+compared to {\sc LM +pos}.
+
+Figure~\ref{fig:top_words} shows the per-dependency 
 omission-score distributions for the five top ranked words plus the word {\it  water}. 

### 2017-05-23 src/analysis.R

@@ -92,17 +92,21 @@ ggplot(filter_by_count(data_q,"dep",MINCOUNT) %>% filter,
 ggsave(file="../doc/imaginet-omission-quotient-dep-boxplot.png", width=5, height=5)
 
 data_lr <- read.table("../data/ridge_scores.txt", header=TRUE) 
-levels(data_lr$predictors)[levels(data_lr$predictors)=="situation"] <- "position"
-data_lr <- data_lr %>% mutate(predictors = factor(predictors, levels=c("word","position","dep","full")), model=factor(model, levels=c("sum", "LM", "textual", "visual")))
+levels(data_lr$predictors)[levels(data_lr$predictors)=="situation"] <- "+pos"
+levels(data_lr$predictors)[levels(data_lr$predictors)=="dep"] <- "+dep"
+
+data_lr <- data_lr %>% mutate(predictors = factor(predictors, levels=c("word","+pos","+dep","full")), 
+                              model=factor(model, levels=c("sum", "LM", "textual", "visual")))
   
 for (model in levels(data_lr$model)) {
-  data_lr[data_lr$model==model,"score"] = data_lr[data_lr$model==model,"R2"]/data_lr[data_lr$model==model & data_lr$predictors=="position","R2"]
+  data_lr[data_lr$model==model,"score"] = data_lr[data_lr$model==model,"R2"]/data_lr[data_lr$model==model & data_lr$predictors=="+pos","R2"]
   
 }
-ggplot(data_lr,#  %>% filter(model %in% c("sum","visual")), 
+ggplot(data_lr, 
        aes(x=model, color=predictors, y=score)) + 
   #geom_bar(stat="identity", position="dodge") +  
   geom_boxplot() +
   theme(aspect.ratio=2/3, text=element_text(size=25)) +
-  ylab('R Squared relative to Position')
+  ylab('R Squared relative to +pos')
 ggsave("../doc/position-new.png")
+

### 2017-05-23 src/ridge.py
Working on fig:top_words

@@ -19,20 +19,22 @@ def dataset(data, score='score'):
     X_full = encoder['full'].fit_transform(choose(dicts, ['word','dep','word:dep','situation','word:situation']))
     return dict(Y=numpy.array(data[score]), encoder=encoder, word=X_word, dep=X_dep, situation=X_situation, full=X_full)
 
-def ridge_alphas(D, ix=100000):
+def tune(D, ix=100000):
     Y = D['Y']
     scores = dict(word=[], dep=[], situation=[], full=[])
     alphas = [ 2**n for n in range(-2,6) ]
-
+    predict = dict(word=[], dep=[], situation=[], full=[])
     def score(alpha, modeltyp):
         model = Ridge(alpha=alpha)
         model.fit(D[modeltyp][:ix,:], Y[:ix])
         scores[modeltyp].append(model.score(D[modeltyp][ix:,:], Y[ix:]))
+        predict[modeltyp].append(model.predict(D[modeltyp][ix:,:]))
 
     for alpha in alphas:
         for modeltyp in ['word','dep','situation','full']:
             score(alpha, modeltyp)
-    return (alphas, scores)
+
+    return {'alphas':alphas, 'scores':scores, 'predict': predict }
 
 def predscore(kind, scores, out):
         for key in scores.keys():
@@ -72,20 +74,30 @@ def main():
     data_t = dataset(data, score='omission_t')
     data_lm = dataset(data, score='omission_lm')
     data_sum = dataset(data, score='omission_sum')
+    IX = 100000
     logging.info("Training Ridge models on visual")
-    alphas, scores_v = ridge_alphas(data_v)
+    result_v = tune(data_v, ix=IX)
     logging.info("Training Ridge models on textual")
-    alphas, scores_t = ridge_alphas(data_t)
+    result_t = tune(data_t, ix=IX)
     logging.info("Training Ridge models on LM")
-    alpahs, scores_lm = ridge_alphas(data_lm)
+    result_lm = tune(data_lm, ix=IX)
     logging.info("Training Ridge models on vectorsum")
-    alphas, scores_sum =ridge_alphas(data_sum)
+    result_sum = tune(data_sum, ix=IX)
     with open("../data/ridge_scores.txt","w") as out:
         out.write("model\tpredictors\tR2\n")
-        predscore("visual", scores_v, out)
-        predscore("textual", scores_t, out)
-        predscore("LM", scores_lm, out)
-        predscore("sum", scores_sum, out)
[…]

### 2017-05-24 src/analysis.R
Rest of section 5.

@@ -110,3 +110,26 @@ ggplot(data_lr,
   ylab('R Squared relative to +pos')
 ggsave("../doc/position-new.png")
 
+
+top_words_er <- read.csv("../data/top7_words_er.csv") %>%
+  mutate(dep = reorder(dep, omission_v))
+ggplot(top_words_er %>% mutate(wordxdep=paste(word,dep)) %>% filter_by_count("wordxdep",5) %>% filter_by_count("dep", 70),
+       aes(x=word, y=omission_v, color=dep)) +
+  geom_boxplot(position = "dodge") +
+  theme(text=element_text(size=25)) +
+  ylab("omission score")
+ggsave(file="../doc/top_words.pdf", width=12, height=8)
+
+data_coef<- read.table(file="../data/position_coef.txt", header=TRUE)
+values <- c("first","second","third","middle","antepenult","penult","last")
+data_coef <- data_coef %>% mutate(coef = factor(coef, levels=values, ordered=TRUE),
+                    model=factor(model, levels=c("sum","LM","textual","visual")))
+ggplot(data_coef, aes(x=as.numeric(coef), y=value, color=model)) + geom_point() + geom_line() +
+  xlab("Position") +
+  ylab("Coefficient") +
+  scale_x_continuous(breaks=1:length(values), 
+                     labels=values) +
+  theme(aspect.ratio=2/3, text=element_text(size=25))
+ggsave("../doc/position-coef.pdf")
+
+

### 2017-05-24 src/ridge.py

@@ -24,17 +24,17 @@ def tune(D, ix=100000):
     scores = dict(word=[], dep=[], situation=[], full=[])
     alphas = [ 2**n for n in range(-2,6) ]
     predict = dict(word=[], dep=[], situation=[], full=[])
+    models = dict(word=[], dep=[], situation=[], full=[])
     def score(alpha, modeltyp):
         model = Ridge(alpha=alpha)
         model.fit(D[modeltyp][:ix,:], Y[:ix])
         scores[modeltyp].append(model.score(D[modeltyp][ix:,:], Y[ix:]))
         predict[modeltyp].append(model.predict(D[modeltyp][ix:,:]))
-
+        models[modeltyp].append(model)
     for alpha in alphas:
         for modeltyp in ['word','dep','situation','full']:
             score(alpha, modeltyp)
-
-    return {'alphas':alphas, 'scores':scores, 'predict': predict }
+    return {'alphas':alphas, 'scores':scores, 'predict': predict, 'models': models }
 
 def predscore(kind, scores, out):
         for key in scores.keys():
@@ -55,6 +55,22 @@ def situation(row):
         return "antepenult"
     else:
         return "middle"
+
+def coef(D, model):
+    coef_full = D['encoder']['full'].inverse_transform(model.coef_)[0]
+    for which in ['first','second','third','middle','antepenult','penult','last']:
+        yield (which, round(coef_full['situation={}'.format(which)],3))
+
+
+def topwords(valid):
+    data_freq = valid.groupby("word").filter(lambda x: len(x) > 100)
+    def er(row):
+        hi = abs(row["omission_v_pred_word"]-row["omission_v"])
+        lo = abs(row["omission_v_pred_dep"]-row["omission_v"])
+        return hi-lo
+    data_freq["er"] = data_freq.apply(er, axis=1)
+    top7_er = data_freq.groupby("word")["er"].mean().sort_values()[-7:].index
+    return data_freq[data_freq["word"].isin(top7_er)][["word","dep","omission_v"]]
 
 # Read and preprocess data
 def main():
@@ -97,6 +113,19 @@ def main():
     valid['omission_v_pred_word'] = result_v['predict']['word'][best_word]
     valid['omission_v_pred_dep'] = result_v['predict']['dep'][best_dep]
     valid.to_csv("../data/ridge_predict.csv", index=False)
+    #valid = pd.read_csv("../data/ridge_predict.csv")
+    logging.info("Saving ridge coefficients")
+    with open("../data/position_coef.txt","w") as out:
[…]

### 2017-05-24 doc/categories.tex
section 5

@@
 \centering
-  \includegraphics[scale=0.35]{position-new.png}
+  \includegraphics[scale=0.35]{position-new.pdf}
 \caption{Proportion of variance in omission scores explained by the
@@
   \centering
-  \includegraphics[scale=0.35]{omission-stat/top_words.pdf}
+  \includegraphics[scale=0.35]{top_words.pdf}
   \caption{Distribution of omission scores per deprel label for the selected word types.}
@@
 occurrence counts of at least 100 and ranked them according to how much
-better, on average, {\sc LR dep} predicted their omission scores 
-compared to {\sc LM +pos}.
+better, on average, {\sc LR +dep} predicted their omission scores 
+compared to {\sc LM word}.
 
@@
 \centering
- \includegraphics[scale=0.4]{omission-stat/position-coef.png}
+ \includegraphics[scale=0.4]{position-coef.pdf}
  \caption{Coefficients on the y-axis of {\sc LM full} corresponding to the

### 2017-05-24 src/analysis.R

@@ -108,7 +108,7 @@ ggplot(data_lr,
   geom_boxplot() +
   theme(aspect.ratio=2/3, text=element_text(size=25)) +
   ylab('R Squared relative to +pos')
-ggsave("../doc/position-new.png")
+ggsave("../doc/position-new.pdf")
 
 
 top_words_er <- read.csv("../data/top7_words_er.csv") %>%

### 2017-05-24 src/analysis.R
Fix messed up plots

@@ -108,7 +108,7 @@ ggplot(data_lr,
   geom_boxplot() +
   theme(aspect.ratio=2/3, text=element_text(size=25)) +
   ylab('R Squared relative to +pos')
-ggsave("../doc/position-new.pdf")
+ggsave("../doc/position-new.pdf", width=10, height=10)
 
 
 top_words_er <- read.csv("../data/top7_words_er.csv") %>%
@@ -130,6 +130,6 @@ ggplot(data_coef, aes(x=as.numeric(coef), y=value, color=model)) + geom_point() 
   scale_x_continuous(breaks=1:length(values), 
                      labels=values) +
   theme(aspect.ratio=2/3, text=element_text(size=25))
-ggsave("../doc/position-coef.pdf")
+ggsave("../doc/position-coef.pdf", width=10, height=10)
 
 

### 2017-05-26 doc/related.tex
related work updated

@@
 
-\namecite{li2015convergent} train a Convolutional Neural Networks (CNN)
-with different random initializations
-%based on the architecture introduced in \cite{krizhevsky2012imagenet} 
-on the ImageNet dataset \cite{krizhevsky2012imagenet}. For each layer
+\namecite{li2015convergent} use a Convolutional Neural Networks (CNN)
+based on the architecture of \namecite{krizhevsky2012imagenet}, 
+and train it on the ImageNet dataset using different random initializations. For each layer
 in all networks they store the activation values produced on the
@@
 
+While this paper was under review, a number of articles appeared which also 
+investigate linguistic representations in LSTM architectures. In an 
+approach similar to ours, \citet{li2016understanding} study the contribution of 
+individual input tokens as well as hidden units and word embedding dimensions 
+by erasing them from the representation and analyzing how this affects the model.
+\citet{adi2016fine} take an alternative approach by introducing prediction tasks 
+to analyze information encoded in sentence embeddings about sentence length, 
+sentence content and word order. Finally, \namecite{linzen2016assessing} examine 
+the acquisition of long-distance syntactic regularities such as number agreement in
+an different versions of an LSTM model with different objectives (number prediction, 
+grammaticality judgment, and language modelling).
+
 In general, there has been a growing interest within computer vision

### 2017-05-29 src/examples.py
examples.py

@@ -1,0 +1,51 @@
+import imaginet.defn.lm_visual_vanilla as D
+import imaginet.data_provider as dp
+import numpy
+import imaginet.evaluate as evaluate
+import logging
+import urllib
+import json
+
+root = '/home/gchrupal/reimaginet/'
+
+def main():
+    logging.getLogger().setLevel('INFO')
+    logging.info("Loading imaginet model")
+    model_im_path = root + "/run-lm_visual_vanilla-1/model.r.e7.zip"
+    model_im = D.load(model_im_path)
+
+    #sent = list(prov.iterSentences(split='val'))
+    logging.info("Building ranker")
+    ranker = Ranker(model_im)
+
+    ex1a = "a baby sits on a bed laughing with a laptop computer open".split()
+    ex1b = "a sits on a bed laughing with a laptop computer open".split()
+    hits = ranker.top1([ex1a, ex1b])
+    dest = "/home/gchrupal/cl-resubmit/doc/"
+    logging.info("Writing images")
+    for hit in hits:
+        logging.info("Writing {}/{}.jpg".format(dest, hit))
+        urllib.urlretrieve("http://mscoco.org/images/{}".format(hit), "{}/{}.jpg".format(dest, hit))
+
+class Ranker:
+    def __init__(self, model):
+        self.model = model
+        prov = dp.getDataProvider(dataset='coco', root=root, audio_kind=None)
+        images = list(prov.iterImages(split='val'))
+        self.img_fs = model.scaler.transform(numpy.array([ img['feat'] for img in images ], dtype='float32'))
+        data = json.load(open("/home/gchrupal/reimaginet/data/coco/dataset.json"))
+        COCOID = {}
+        for img in data['images']:
+            COCOID[img['imgid']] = img['cocoid']
+        self.IDS = [COCOID[img['imgid']] for img in images]
+
+    def top1(self, sents):
+        pred = D.encode_sentences(self.model, sents)
+        distances = evaluate.Cdist(batch_size=2**13)(pred, self.img_fs)
+        result = []
+        for row in distances:
+            result.append(self.IDS[numpy.argsort(row)[0]])
+        return result
+
+if __name__ == '__main__':
+    main()

### 2017-05-29 doc/categories.tex
Changed figures.

@@
 
+% \begin{figure*}[t]
+% \hspace*{-0.5in}
+% \setlength{\tabcolsep}{0.01pt}
+%     \begin{tabular}{cccc}
+%     \includegraphics[scale=0.14]{new_omission_examples/textual1} &
+%     \includegraphics[scale=0.14]{new_omission_examples/visual1} &
+%     \includegraphics[scale=0.33]{new_omission_examples/img1} & 
+%     \includegraphics[scale=0.28]{new_omission_examples/img1_omit} \\
+%     \includegraphics[scale=0.16]{new_omission_examples/textual2} &
+%     \includegraphics[scale=0.16]{new_omission_examples/visual2} &
+%     \includegraphics[scale=0.28]{new_omission_examples/img2} & 
+%     \includegraphics[scale=0.30]{new_omission_examples/img2_omit}
+%     \end{tabular}
+%     \caption{The omission scores of {\sc Visual} and {\sc Textual} for two example sentences, 
+% and the best retrieved images for the original sentence (left) and the sentence with the most 
+% important word removed (right). 
+% %Top row example caption is {\it a bottle next to a windowsill with
+% %  light coming through},  bottom row: {\it a train on a track next to
+% %  many bushes}.
+% }
+%     \label{fig:omissionex}
+% \end{figure*}
+
 \begin{figure*}[t]
-\hspace*{-0.5in}
-\setlength{\tabcolsep}{0.01pt}
-    \begin{tabular}{cccc}
-    \includegraphics[scale=0.14]{new_omission_examples/textual1} &
-    \includegraphics[scale=0.14]{new_omission_examples/visual1} &
-    \includegraphics[scale=0.33]{new_omission_examples/img1} & 
-    \includegraphics[scale=0.28]{new_omission_examples/img1_omit} \\
-    \includegraphics[scale=0.16]{new_omission_examples/textual2} &
-    \includegraphics[scale=0.16]{new_omission_examples/visual2} &
-    \includegraphics[scale=0.28]{new_omission_examples/img2} & 
-    \includegraphics[scale=0.30]{new_omission_examples/img2_omit}
-    \end{tabular}
-    \caption{The omission scores of {\sc Visual} and {\sc Textual} for two example sentences, 
-and the best retrieved images for the original sentence (left) and the sentence with the most 
-important word removed (right). 
-%Top row example caption is {\it a bottle next to a windowsill with
-%  light coming through},  bottom row: {\it a train on a track next to
-%  many bushes}.
-}
+  \centering
+  \includegraphics[scale=0.5]{omissionex.pdf}
+  \caption{Omission scores for an example sentence for {\sc
+      LM} and the two pathways, {\sc Textual} and {\sc Visual}, of
+    {\sc Imaginet.}}
   \label{fig:omissionex}
+\end{figure*}
+
+\begin{figure*}[t]
+  \centering
+  \includegraphics[scale=0.25]{85826.jpg}
+  \includegraphics[scale=0.25]{60596.jpg}
+  \caption{Images retrieved for an example sentence (left) and the version of
+    it with the  second word omitted (right).}
+  \label{fig:omissionexpic}
 \end{figure*}

### 2017-05-29 doc/related.tex
related work updated

--- 2017-05-23 doc/related.tex
@@
 for a deeper understanding of the activation patterns of RNNs, but focus on models with 
-modern architectures trained on large scale data sets. More specifically, they train Long Short-Term Memory networks (LSTM) \cite{hochreiter1997long} 
+modern architectures trained on large scale data sets. More specifically, they train Long 
+Short-Term Memory networks (LSTM) \cite{hochreiter1997long} 
 for phrase level sentiment analysis 
@@
 
+While this paper was under review, a number of articles appeared which also 
+investigate linguistic representations in LSTM architectures. In an 
+approach similar to ours, \namecite{li2016understanding} study the contribution of 
+individual input tokens as well as hidden units and word embedding dimensions 
+by erasing them from the representation and analyzing how this affects the model.
+\namecite{adi2016fine} take an alternative approach by introducing prediction tasks 
+to analyze information encoded in sentence embeddings about sentence length, 
+sentence content and word order. Finally, \namecite{linzen2016assessing} examine 
+the acquisition of long-distance dependencies through the study of number agreement 
+in different variations of an LSTM model with different objectives (number prediction, 
+grammaticality judgment, and language modeling). Their results show that such dependencies 
+can be captured with very high accuracy when the model receives a strong supervision 
+signal (that is, whether the subject is plural or singular), but simple language models still capture 
+the majority of test cases. All these models focus on text-based tasks and do not take other 
+modalities such as visual input into account.
+
 In general, there has been a growing interest within computer vision

### 2017-05-29 src/analysis.R
Git more messing with section 5.

@@ -4,20 +4,24 @@ library(ggplot2)
 dep <- read.csv("../data/depparse_coco_val.csv", header=TRUE)
 omi <- read.csv("../data/omission_coco_val.csv", header=TRUE)
 
-omi_v  <- omi %>% select(-omission_t, -omission_lm) %>% rename(score=omission_v) %>% mutate(pathway="visual")
-omi_t  <- omi %>% select(-omission_v, -omission_lm) %>% rename(score=omission_t) %>% mutate(pathway="textual")
-omi_lm <- omi %>% select(-omission_t, -omission_v)  %>% rename(score=omission_lm) %>% mutate(pathway="lm")
+omi_v  <- omi %>% select(-omission_t, -omission_lm, -omission_sum) %>% rename(score=omission_v) %>% mutate(pathway="visual")
+omi_t  <- omi %>% select(-omission_v, -omission_lm, -omission_sum) %>% rename(score=omission_t) %>% mutate(pathway="textual")
+omi_lm <- omi %>% select(-omission_t, -omission_v, -omission_sum)  %>% rename(score=omission_lm) %>% mutate(pathway="lm")
+omi_sum <- omi %>% select(-omission_v, -omission_t, -omission_lm)  %>% rename(score=omission_sum) %>% mutate(pathway="sum")
+
+
 
 data_v <- merge(dep, omi_v) 
 data_t <- merge(dep, omi_t) 
 data_lm <- merge(dep, omi_lm)
+data_sum <- merge(dep, omi_sum)
 
 dep_lev <- (data_v %>% group_by(dep) %>% summarize(median_score=median(score)) %>% 
               data.frame %>% arrange(median_score))$dep 
 postag_lev <- (data_v %>% group_by(postag) %>% summarize(median_score=median(score)) %>%
               data.frame %>% arrange(median_score))$postag
 
-data <- rbind(data_t, data_v) %>% rbind(data_lm) %>% 
+data <- rbind(data_t, data_v) %>% rbind(data_lm) %>% rbind(data_sum) %>%
   mutate(dep = factor(dep, levels=dep_lev), 
          postag = factor(postag, levels=postag_lev))
 
@@ -27,8 +31,23 @@ filter_by_count <- function(dat, col, count) {
 MINCOUNT<-500
 
 
-# FIGURE 2
-ggplot(filter_by_count(data, "postag", MINCOUNT)) +
+# figure fig:omissionex
[…]

### 2017-05-29 doc/categories.tex
Git more messing with section 5.
Missing file.

@@
   \centering
-  \includegraphics[scale=0.5]{omissionex.pdf}
-  \caption{Omission scores for an example sentence for {\sc
-      LM} and the two pathways, {\sc Textual} and {\sc Visual}, of
-    {\sc Imaginet.}}
+  \includegraphics[scale=0.3]{omissionex.pdf}
+  \caption{Omission scores for an example sentence for {\sc LM} and
+    the two pathways, {\sc Textual} and {\sc Visual}, of {\sc
+      Imaginet.}}
   \label{fig:omissionex}
@@
   \includegraphics[scale=0.25]{60596.jpg}
-  \caption{Images retrieved for an example sentence (left) and the version of
-    it with the  second word omitted (right).}
+  \caption{Images retrieved for the example sentence {\it a baby sits
+      on a bed laughing with a laptop computer open} (left) and the
+    same sentence with the second word omitted (right).}
   \label{fig:omissionexpic}
@@
 from a linguistic point of view and focus
-on the high level understanding of what types of words the networks
+on the high level understanding of what parts of input sentence the networks
 pay most attention to. Furthermore, we investigate if the networks
-learn to assign appropriate amounts of importance to tokens depending
+learn to assign differential amounts of importance to tokens depending
 on their position and grammatical function in the sentences.\label{edit:whyposdep}
@@
 of dependency relations and part-of-speech categories and compares {\sc Visual}
-and {\sc Textual}. Lastly Section \ref{sec:beyondlexical} investigates the extent to which
-the importance of words for the different pathways depend on the words themselves,
+and {\sc Textual}. Section \ref{sec:beyondlexical} investigates the extent to which
+the importance of words for the different networks depend on the words themselves,
 their position and their grammatical function in the sentences.

### 2017-05-29 doc/response.tex
response letter updated

@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{this figure should be removed.}
+\textsc{Author Response:}  Following the recommendation of the reviewers,
+we shortened Section~\ref{sec:micro} and removed this figure.
 \end{quote}
@@
   value? are negative activations not important?
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  It is in fact the absolute value of the activations 
+that is taken into account. We have changed the description of the top-k-contexts 
+method to clear this issue.
+\end{quote}
+\begin{verbatim}
 
@@
   not clear what a "unit" is and how it is chosen.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  A unit is one neuron in the final hidden layer 
+of the network. We have made this clear in the revised text.
+\end{quote}
+\begin{verbatim}
 
@@
   appreciate an elaboration on this point (or a removal).
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  We removed that sentence and added an expanded discussion of the generalizability of our techniques to the last section.
+\end{quote}
+\begin{verbatim}
 
@@
   not clear from the description of the method.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:} \todo[inline]{This needs explanation and I don't have an answer (yet).}
+\end{quote}
+\begin{verbatim}
 
@@
   function.
+\end{verbatim}  
+\begin{quote}
+\textsc{Author Response:}  We have reorganized Section~\ref{sec:micro} and 
+tried to motivate this particular experiment better. The idea is to 
+come up with techniques for singling out units with a specific type
+of function, such as encoding grammatical function. Therefore we 
+use the coefficients of the logistic regression model as pointers 
+to units of this type.
+\end{quote}
+\begin{verbatim}
 
@@
 \begin{quote}
-\textsc{Author Response:} probably reorganise this section \todo[inline]{explain changes}
+\textsc{Author Response:} It was there, believe us! But we have reorganized 
+Section~\ref{sec:micro} and removed this experiment for the sake of clarity.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{add to related work}
+\textsc{Author Response:}  We have expanded the discussion of related work 
+(Section~\ref{sec:related}) and included this work in addition to a few others which 
+came out while our submission was under review.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{explain reorganisations and major changes.}
+\textsc{Author Response:}  We hope we have addressed all your comments and concerned.
+We did not remove the experiments in Section 6.3, but as we mentioned before we rewrote and restructured Section~\ref{sec:micro} and tried to put this experiment in context and clear our motivation.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{explain restructuring of section 6}
+\textsc{Author Response:}  You are absolutely right about the description of 
+the top-k-context method being confusing. We have rewritten this part to make the description of the method more clear. Furthermore, we have reorganized 
+Section~\ref{sec:micro} to make it more coherent. 
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{this section should probably be removed}
+\textsc{Author Response:}  Following your and others' suggestion, we have removed this section from the revised manuscript.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{explain}
+\textsc{Author Response:}  As mentioned above, we have rewritten Section~\ref{sec:micro} to make the description of the methods and the experiments more clear.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{discuss linzen et al. in related work}
+\textsc{Author Response:} We have expanded the related work 
+(Section~\ref{sec:related}) with a discussion of Linzen et al. (2016) 
+and a few other articles that came out while our submission was under review.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{add this to related work.}
+\textsc{Author Response:}  As mentioned before, we have included a discussion of
+this paper in our related work.
 \end{quote}

### 2017-05-29 doc/response.tex
related work and response letter updated

@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{add to related work}
+\textsc{Author Response:} We have edited our related work section to make the comparison more 
+clear. In addition to taking into account tasks that rely on different modalities (language as well as vision), 
+we focus more on learning structural properties of the input, whereas Li et al. (2015) focus on single
+tokens.
 \end{quote}

### 2017-05-29 doc/categories.tex
Mostly adding stuff about LM

@@
 on their position and grammatical function in the sentences.\label{edit:whyposdep}
-In Section \ref{sec:computeomission} we introduce \emph{omission scores};
-a metric to measure the contribution of each token to the prediction of the networks.
-Section \ref{sec:omitimaginet} aggregates the omission scores in terms
-of dependency relations and part-of-speech categories and compares {\sc Visual}
-and {\sc Textual}. Section \ref{sec:beyondlexical} investigates the extent to which
+In Section~\ref{sec:computeomission} we introduce \emph{omission scores};
+a metric to measure the contribution of each token to the prediction
+of the networks and in Section~\ref{sec:omitimaginet} we analyze how
+these scores are distributed over dependency relations and
+part-of-speech categories. 
+In Section~\ref{sec:beyondlexical} we investigate the extent to which
 the importance of words for the different networks depend on the words themselves,
-their position and their grammatical function in the sentences.
+their sequential position and their grammatical function in the sentences.
 % MAYBE THIS WOULD BE TOO MUCH
@@
 
-In both pathways of {\sc Imaginet} the full sentences are represented by the
+In all the models the full sentences are represented by the
 activation vector at the end-of-sentence symbol
@@
 \noindent Figure~\ref{fig:omissionex} demonstrates the omission
-measure for the {\sc Visual} and {\sc Textual} pathways for two
-example captions. For both captions the $\mathrm{omission}$ scores are
-plotted along with the first image retrieved by {\sc Visual} for the
+measure for the {\sc LM}, {\sc Visual} and {\sc Textual} networks for an
+example captions. The $\mathrm{omission}$ scores are
+plotted for the
 full sentence and for the sentence with the word with the highest
-$\mathrm{omission}(i,S)$ removed. The images are retrieved from the
-validation set of MS-COCO by: 1) computing the image representation of the 
-given sentence with {\sc Visual}; 2) extracting the CNN features for the 
-images from the set; and 3) finding the image that minimizes the cosine distance
-to the query.\label{edit:retrievalexplain} 
-In the first example, the omission
-scores for {\sc Visual} suggest that the model interpreted {\it  pizza} as the most important
-word in the sequence and returned an image that depicts a pizza on a plate. 
-Removing the word {\it pizza} promotes {\it salad} and {\it bowl} as
-the main theme of the sentence and the model retrieves an image with a dining table 
-with salad-like dishes. 
-For the second example, the omission scores for {\sc
-  Visual} show that the model paid attention mostly to {\it baby} and
-{\it bed} and slightly to {\it laptop} and retrieved an image depicting a baby 
-sitting on a bed with a laptop.
-Removing the word {\it baby} leads
-to an image that depicts an adult male laying on a bed. Figure~\ref{fig:omissionex}
-also shows that in contrast to {\sc Visual}, in both examples {\sc
-  Textual} distributes its attention more evenly across time steps
-instead of focusing on the types of words related to the corresponding
-visual scene.
+$\mathrm{omission}(i,S)$ removed. 
+Figure~\ref{fig:omissionexpic} shows the images retrieved by {\sc
+  Visual} for the full
+caption and for the one with the word {\sc baby} omitted. 
+The images are retrieved from the validation set of MS-COCO by: 1)
+computing the image representation of the given sentence with {\sc
+  Visual}; 2) extracting the CNN features for the images from the set;
+and 3) finding the image that minimizes the cosine distance to the
+query.\label{edit:retrievalexplain} 
+The omission scores for {\sc Visual} show that the model paid attention
+mostly to {\it baby} and {\it bed} and slightly to {\it laptop} and
+retrieved an image depicting a baby sitting on a bed with a laptop.
+Removing the word {\it baby} leads to an image that depicts an adult
+male laying on a bed. Figure~\ref{fig:omissionex} also shows that in
+contrast to {\sc Visual}, {\sc Textual} distributes
+its attention more evenly across time steps instead of focusing on the
+types of words related to the corresponding visual scene. The peaks
+for {\sc LM} are the same as for {\sc Textual}, but the variance of
+the omission scores is higher, suggesting that the unimodal language
+model is more sensitive overall to input perturbations than {\sc Textual}.
 
-For other RNN architectures such as GRUs, LSTMs \label{edit:omitgeneral}
+For other RNN architectures such as LSTMs \label{edit:omitgeneral}
 and their bi-directional variants, measuring the contribution
@@
 \caption{Distribution of omission scores for POS (left) and dependency labels
-  (right), for the {\sc Textual} and {\sc Visual} pathways. Only labels which occur at least 500 times are included.}
+  (right), for the {\sc Textual} and {\sc Visual} pathways and for
+  {\sc LM}. Only labels which occur at least 500 times are included.}
 \label{fig:omission-imaginet}
@@
 
+The $\mathrm{omission}$ scores can be used not only to
+estimate the importance of individual words, but also of syntactic
+categories. We estimate the salience of each syntactic category by
+accumulating the omission scores for all words in that category. We
+tag every word in a sentence with the part-of-speech (POS) category
+and the dependency relation (deprel) label of its incoming arc. For
+example, for the sentence \emph{the black dog}, we get ({\it
+  the},~DT,~det), 
+({\it black},~JJ,~amod), ({\it dog},~NN,~root). 
+Both POS tagging and dependency parsing are performed 
+using the  dependency parser from the Spacy package.\footnote{Available at
+  \url{https://spacy.io/}.} 
+%The POS tags used are the Penn Treebank tags and the dependencies are 
+%the Stanford basic dependencies.
+
+Figure~\ref{fig:omission-imaginet} shows the distribution of omission
+scores per POS and dependency label for the two pathways of {\sc
+  Imaginet} and for {\sc LM}.  The general trend is that for the {\sc
+  Visual} pathway, the omission scores are high for a small subset of
+labels - corresponding mostly to nouns, less so for adjectives and
+even less for verbs - and low for the rest (mostly function words and
+various types of verbs). For {\sc Textual} the differences
+are \label{edit:textualomission} smaller, and the pathway seems to be
+sensitive to the omission of most types of words.  For {\sc LM} the
+distribution over categories is also relatively uniform, but the omission scores are higher
+overall than for {\sc Textual}.
+
+Figure~\ref{fig:omission-imaginet-ratio} compares the two pathways of
+{\sc Imaginet} directly using the log of the ratio of the {\sc Visual}
+to {\sc Textual} omission scores, and plots the distribution of this
+ratio for different POS and dependency labels.  Log ratios above zero
+indicate stronger association with the {\sc Visual} pathway and below
+zero with the {\sc Textual} pathway. We see that in relative terms,
+{\sc Visual} is more sensitive to adjectives (JJ), nouns (NNS, NN) and
+numerals (CD), and {\sc Textual} to the prepositions (TO, IN),
+determiners (DT, WDT) and pronouns (PRP) as well as some types of
+verbs (VBZ, VBP).
+
+This picture is complemented by the analysis of the
+relative importance of dependency relations: {\sc Visual} pays most
+attention to the relations {\sc nsubjpass, attr, amod, nsubj, ROOT,
+  compound, dobj, nummod}
+whereas {\sc Textual} is more sensitive to {\sc det, mark, prep, aux,
+  auxpass, cc, poss, advmod, prt, relcl}.
+As expected, {\sc Visual} is more focused on grammatical
+functions typically filled by semantically contentful words, while
+{\sc Textual} distributes its attention more uniformly and 
+attends relatively more to purely grammatical functions.
+
+It is worth noting, however, the relatively low omission scores for verbs in case \label{edit:generality}
+of {\sc Visual}. One might expect that the task of image prediction from
+descriptions requires general language understanding and so high omission
+scores for all content words in general, however, the results
+suggest that this setting is not optimal for learning useful representations of verbs,
+which possibly leads to representations that are too task specific 
+and not transferable across tasks.
+
 \begin{figure*}[t]
@@
     POS (left) and dependency labels (right). Only labels which occur at least 500 times are included.}
-\label{fig:omission-imaginet-ratio}
+\label{fig:omission-imaginet-quotient}
 \end{figure*}
 
-The $\mathrm{omission}$ scores can be used not only to
-estimate the importance of individual words, but also of syntactic
-categories. We estimate the salience of each syntactic category by
-accumulating the omission scores for all words in that category. We
-tag every word in a sentence with the part-of-speech (POS) category
-and the dependency relation (deprel) label of its incoming arc. For
-example, for the sentence \emph{the black dog}, we get ({\it
-  the},~DT,~det), 
-({\it black},~JJ,~amod), ({\it dog},~NN,~root). 
-Both POS tagging and dependency parsing are performed jointly
-using the TurboParser dependency parser \cite{martins2013turning}.\footnote{Available at
-  \url{https://github.com/andre-martins/TurboParser}{github.com/andre-martins/TurboParser}.} The POS tags used are the Penn Treebank tags and the dependencies are 
-the Stanford basic dependencies.
-
-Figure~\ref{fig:omission-imaginet} shows the distribution of omission
-scores per POS and deprel label for the two pathways of {\sc Imaginet}. 
-The general trend is that for the {\sc Visual} pathway, the omission scores are
-high for a small subset of labels - corresponding mostly to nouns, less so for
-adjectives and even less for verbs - and low for the rest (mostly function words and
-various types of verbs). For {\sc Textual} the differences are \label{edit:textualomission}
-smaller, and the pathway seems to be sensitive to the omission of most
-types of words.
-
-Figure~\ref{fig:omission-imaginet-ratio} compares the
-two pathways directly using the log of the ratio of the {\sc Visual}
-to {\sc Textual} omission scores, and plots the distribution of this
-ratio for different POS and deprel labels.  Log ratios above zero
-indicate stronger association with the {\sc Visual} pathway and below
-zero with the {\sc Textual} pathway. We see that in relative terms,
-{\sc Visual} is more sensitive to nouns (NNP, NNS, NN), numerals (CD)
-and adjectives (JJ), and {\sc Textual} to the prepositions (TO, IN),
-some types of verbs (VBZ, VBP), determiners (DET, WDT) and
-particles (RP). This picture is complemented by the analysis of the
-relative importance of dependency relations: {\sc Visual} pays most
-attention to the relations {\sc nsubjpass, nsubj, pobj, root, dobj,
-  nn, conj, dep, amod} and {\sc num}, whereas {\sc Textual} is more
-sensitive to {\sc det, mark, aux, auxpass, prt, prep, poss, cc} and
-{\sc cop}.
-
-As expected, {\sc Visual} is more focused on grammatical
-functions typically filled by semantically contentful words, while
-{\sc Textual} distributes its attention more uniformly and 
-attends relatively more to purely grammatical functions. 
-It is worth noting, however, the relatively low omission scores for verbs in case \label{edit:generality}
-of {\sc Visual}. One might expect that the task of image prediction from
-descriptions requires general language understanding and so high omission
-scores for all content words in general, however, the results
-suggest that this setting is not optimal for learning useful representations of verbs,
-which possibly leads to representations that are too task specific 
-and not transferable across tasks.
+Figure~\ref{fig:omission-imaginet-quotient} shows a similar analysis
+contrasting {\sc LM} with the {\sc Textual} pathway of {\sc
+  Imaginet}. The first observation is that the range of values of the
+log ratios is narrow, indicating that the differences between these
+two networks regarding which grammatical categories they are sensitive
+to is less pronounced than when comparing {\sc Visual} to {\sc
+  Textual}. While the size of the effect is weak, there also seems to
+be a tendency for the {\sc Textual} model to pay relatively more
+attention to lexical and less to structural dependencies, compared to
+{\sc LM}: it may be that the {\sc Visual} pathway pulls {\sc Textual}
+in this direction by sharing word embeddings with it.
 

### 2017-05-29 doc/experiments.tex
Mostly adding stuff about LM

@@
 we explore the kinds of linguistic regularities the
-{\sc Textual} and {\sc Visual} pathways of {\sc Imaginet} learn from
+networks learn from
 word-level input. Section~\ref{sec:macro} presents the macro-level
@@
 the linguistic features encoded by individual hidden units. 
-In both sections we report our findings based on the {\sc Imaginet} model
-and discuss the generalizabilty of our methods to other architectures.
-\label{edit:experimentsgeneral}
-For all the experiments, we trained {\sc Imaginet} on the training portion of the
+In both sections we report our findings based on the {\sc Imaginet}
+model, discuss the generalizabilty of our methods to other architectures.
+\label{edit:experimentsgeneral}, and whenever appropriate compare to
+our two other models, {\sc LM} and {\sc Sum}.  For all the
+experiments, we trained {\sc Imaginet} on the training portion of the
 MSCOCO image-caption dataset \cite{lin2014microsoft}, and analyzed the
-representations of the sentences in the validation set. The target
-image representations were extracted from the pre-softmax layer of the
-16-layer CNN of \namecite{simonyan2014very}.
+representations of the sentences in the validation set corresponding
+to 5000 randomly chosen images. The target image representations were
+extracted from the pre-softmax layer of the 16-layer CNN of
+\namecite{simonyan2014very}.
 

### 2017-05-29 src/analysis.R

@@ -51,6 +51,7 @@ ggplot(data %>% filter(pathway != "sum") %>% filter_by_count("postag", MINCOUNT)
   aes(x=postag, y=score, color=pathway) +
   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
   coord_flip() +
+  theme(legend.title=element_blank()) +
   #theme(text=element_text(size=15)) +
   xlab("Part of Speech") +
   ggsave(file="../doc/imaginet-omission-pos-boxplot.png",
@@ -61,6 +62,7 @@ ggplot(data %>% filter(pathway != "sum") %>% filter_by_count("dep", MINCOUNT)) +
        aes(x=dep, y=score, color=pathway) +
   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
   coord_flip() +
+  theme(legend.title=element_blank()) +
   #theme(text=element_text(size=15)) +
   xlab("Dependency") +
   ggsave(file="../doc/imaginet-omission-dep-boxplot.png",
@@ -81,7 +83,7 @@ ggplot(filter_by_count(data_r,"postag",MINCOUNT) %>% filter,
   xlab("Part of Speech")
 ggsave(file="../doc/imaginet-omission-ratio-pos-boxplot.png", width=5, height=5)
 
-ggplot(filter_by_count(data_r,"dep",MINCOUNT) %>% filter,
+ggplot(filter_by_count(data_r,"dep",MINCOUNT),
        aes(x=dep, y=log(ratio))) +
   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
   coord_flip() +
@@ -94,7 +96,7 @@ data_q <- filter(data, pathway=="lm") %>% mutate(ratio=quotient,
                                                      postag=reorder(postag, ratio, median),
                                                      dep=reorder(dep, ratio, median))
 
-ggplot(filter_by_count(data_q,"postag",MINCOUNT) %>% filter,
+ggplot(filter_by_count(data_q,"postag",MINCOUNT),
        aes(x=postag, y=log(ratio))) +
   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
   coord_flip() +

### 2017-05-29 doc/related.tex

@@
 and visualize local semantic compositionality in RNNs. 
-In comparison, the present work focuses more on exploring structure 
-learning in RNNs and on developing methods for a comparative analysis between RNNs.
+In comparison, the present work goes beyond the importance of single words and focuses more on exploring structure 
+learning in RNNs, as well as on developing methods for a comparative analysis between RNNs that are focused on different modalities (language versus vision).
 

### 2017-05-30 src/mutual.py
Mutual information stuff.

@@ -1,0 +1,41 @@
+import pandas as pd
+import numpy as np
+import imaginet.data_provider as dp
+from scipy.stats.mstats import mquantiles
+from sklearn.metrics import mutual_info_score
+import csv
+import logging
+
+def ngram(seq, order):
+    pad = ['*--*']*(order-1)+seq
+    return [ ' '.join(tup) for tup in zip(*[pad[i:] for i in range(order)]) ]
+
+
+def quantize(v):
+    return np.digitize(v, mquantiles(v, prob=np.linspace(0,1,20)))
+
+def main():
+    logging.getLogger().setLevel('INFO')
+    logging.info("Loading input data")
+    root = '/home/gchrupal/reimaginet/'
+    prov = dp.getDataProvider(dataset='coco', root=root, audio_kind=None)
+    ids = np.array([senti['sentid'] for senti in prov.iterSentences(split='val')])
+    data = pd.read_csv("/home/gchrupal/cl-resubmit/data/depparse_coco_val.csv")
+    sent_data = dict(senti for senti in data.groupby(['sentid']))
+    with open("/home/gchrupal/cl-resubmit/data/mutual.csv", "w") as out:
+        writer = csv.writer(out)
+        writer.writerow(["pathway", "condition", "order", "dimension", "mi"])
+        for net in ["visual", "textual", "lm", "sum"]:
+            states = np.load("/home/gchrupal/cl-resubmit/data/states_{}.npy".format(net), encoding='bytes')
+            S = quantize(np.vstack([state[:-1,:] for state in states]).T)
+            for typ in ['word', 'dep']:
+                for order in [1,2,3]:
+                    logging.info("Computing scores for {}:{}:{}".format(net, typ, order))
+                    context = [item for i in ids for item in ngram(list(sent_data[i][typ]), order) ]
+                    #print(context[:10])
+                    for i in range(1024):
+                        mi = mutual_info_score(S[i], context)
+                        writer.writerow([net, typ, order, i, mi])
+
+if __name__ == '__main__':
+    main()

### 2017-05-30 src/omission.py

@@ -5,12 +5,20 @@ import imaginet.data_provider as dp
 import imaginet.defn.lm_visual_vanilla as D
 from scipy.spatial.distance import cosine
 import imaginet.task
+import numpy
 
 root = '/home/gchrupal/reimaginet/'
 data_path = '/home/gchrupal/cl-resubmit/data/'
 model_im_path = root + "/run-lm_visual_vanilla-1/model.r.e7.zip"
 model_lm_path = root + "/run-lm-0/model.r.e6.zip"
 model_sum_path = root + "/run-vectorsum-0/model.r.e7.zip"
+
+
+def dump_activations(sent, models):
+    for name, model, task in models:
+        logging.info("Writing activations for {}".format(name))
+        states = imaginet.task.states(model, sent, task=task)
+        numpy.save(data_path + "states_{}.npy".format(name), states)
 
 def main():
     logging.getLogger().setLevel('INFO')
@@ -25,6 +33,11 @@ def main():
     model_lm = imaginet.task.load(model_lm_path)
     logging.info("Loading vectorsum model")
     model_sum = imaginet.task.load(model_sum_path)
+    dump_activations([sent_i['tokens'] for sent_i in sent ],
+                           [("visual", model_im, model_im.visual ),
+                            ("textual", model_im, model_im.lm),
+                            ("lm", model_lm, model_lm.task),
+                            ("sum", model_sum, model_sum.task)])
     writer = csv.writer(open(data_path + '/omission_coco_val.csv',"w"))
     writer.writerow(["sentid", "position", "word",
                      "omission_v",

### 2017-05-30 doc/categories.tex

@@
 be a tendency for the {\sc Textual} model to pay relatively more
-attention to lexical and less to structural dependencies, compared to
+attention to content and less to function words, compared to
 {\sc LM}: it may be that the {\sc Visual} pathway pulls {\sc Textual}
 in this direction by sharing word embeddings with it.
+
+Most of our findings up to this point conform reasonably well to prior
+expectations about effects that  particular learning objectives should
+have. This fact serves to validate our methods. In the next section we
+go on to investigate less straightforward patterns.
 
@@
 
-Models that utilize the sequential structure of natural-language 
+Models that utilize the sequential structure of language 
 have the capacity to interpret the same word-type differently depending on
@@
 pathways are differentially sensitive to content vs.\ function
-words. This may be either just due to purely lexical features or the model 
+words. In principle, this may be either just due to purely lexical features or the model 
 may actually learn to pay more attention to the same word type in appropriate
@@
 of sentences in a left-to-right or right-to-left fashion. 
-Bi-directional RNNs, however, are affected by word-order and can potentially
+Bi-directional RNNs, are sensitive by word-order and can potentially
 learn to handle the same word in different positions differently. \label{edit:foot}}
@@
 scores per token with the following predictor variables: 
-
 \begin{enumerate}
@@
 \noindent We split the 5-thousand-image portion of MSCOCO validation data into two parts, fit the 
-models (regularized via L2 penalty) on the first part, and compute the 
+models on the first part, and compute the 
 proportion of variance explained on the second part. 
-For comparison we also train a version of {\sc Imaginet},
-where the GRU in the {\sc Visual} pathway is replaced by a simple vector-summation operation
-over word-embeddings and thus does not have access to sequential cues. We refer to this
-model as {\sc Sum}.
+For comparison we also use the {\sc Sum} model which composes word
+embeddings via summation, and uses the same loss function as {\sc
+  Visual}. This model is unable to encode information about word
+order, and thus is a good baseline here as we investigate the
+sensitivity of the networks to positional and structural cues.
 
@@
 Figure~\ref{fig:rsquared} offers a different view of the data, showing
-the increase or decrease in $R^2$ for the models per pathway relative to {\sc LR
-  +pos}.  The results reveal that adding additional information on
+the increase or decrease in $R^2$ for the models relative to {\sc
+  LR~+pos}.  
+The results reveal that adding additional information on
 top of lexical features in case of {\sc Sum} does increase the
-explained variance slightly, which is probably due to the unseen words
+explained variance slightly, which is most likely due to the unseen words
 in the held out set. More interestingly, there is a sizeable
-increase in $R^2$ between {\sc LR +pos} and {\sc LR full} in
+increase in $R^2$ between {\sc LR~+pos} and {\sc LR~full} in
 case of {\sc Visual}, suggesting that 
 the omission scores in case of {\sc Visual} depend on the words'
-grammatical function in sentences, even after controling for word
-identity and linear position.
+grammatical function in sentences, {\it even after controling for word
+identity and linear position.}
 
@@
 these language models: this is not surprising considering that the omission
-scores are measured with respect to the final activation state. 
-Less predictably, for the {\sc Visual} model,
+scores are measured with respect to the final activation state, and
+given the fact that in a language model the recent history is most
+important for accurate prediction. Less expectedly, for the {\sc Visual} model,
 dependency labels are more informative than linear position,
@@
   linear regression models
- for {\sc Sum}, {\sc Visual} and {\sc Textual} pathways, relative to
+ for {\sc Sum}, {\sc LM}, {\sc Visual} and {\sc Textual}, relative to
  regressing on word identity and position only. }
@@
 In order to find out some of the specific syntactic configurations leading to
-increase in $R^2$ between {\sc LR word} and {\sc LR dep} predictors
+increase in $R^2$ between the {\sc LR~word} and {\sc LR~dep} predictors
 in case of {\sc Visual}, we next considered all word types with
 occurrence counts of at least 100 and ranked them according to how much
-better, on average, {\sc LR +dep} predicted their omission scores 
-compared to {\sc LM word}.
+better, on average, {\sc LR~+dep} predicted their omission scores 
+compared to {\sc LM~word}.
 
 Figure~\ref{fig:top_words} shows the per-dependency 
-omission-score distributions for the five top ranked words plus the word {\it  water}. 
+omission-score distributions for seven top ranked words.
 There are clear and large differences in how these words
@@
 they share their contribution with the first, often more important,
-member of the conjunction (e.g. {\it A cow and its baby eating
-  grass}).
+member of the conjunction, for example: {\it A cow and its baby eating
+  grass}. 
+% The pattern for {\sc nn} (nominal modifier) is a bit more complicated: for
+% four of the words shown (as well as for most other words not shown in
+% the figure), the score is very low in this grammatical function--presumably 
+% because most words contribute less to the sentence meaning when used  
+% as modifiers than as heads (e.g.\ {\it a clock tower}). However,
+% for the words {\it zebra} and {\it water}, omission scores are high
+% when they act as a nominal modifier {\sc nn}. This appears due to two reasons:
+% \begin{enumerate}
+% \item For {\it zebra}, there are frequent erroneous parses such as {\it
+%     zebra/{\sc nn} browsing/{\sc root}} instead of {\it zebra/{\sc nsubj} browsing/{\sc
+%     root}}. The network does not make this mistake, and treats these
+%     occurrences of {\it zebra} according to its importance as {\sc nsubj}.
+% \item {\it Water} as a modifier often changes the meaning of its head in
+%  a visually salient way: e.g.\ {\it water fall, water balloon, water scene, water skiing}, and
+%   thus the network learns that this particular word is important in the modifier position.
+% \end{enumerate}
 
-The pattern for {\sc nn} (nominal modifier) is a bit more complicated: for
-four of the words shown (as well as for most other words not shown in
-the figure), the score is very low in this grammatical function--presumably 
-because most words contribute less to the sentence meaning when used  
-as modifiers than as heads (e.g.\ {\it a clock tower}). However,
-for the words {\it zebra} and {\it water}, omission scores are high
-when they act as a nominal modifier {\sc nn}. This appears due to two reasons:
-\begin{enumerate}
-\item For {\it zebra}, there are frequent erroneous parses such as {\it
-    zebra/{\sc nn} browsing/{\sc root}} instead of {\it zebra/{\sc nsubj} browsing/{\sc
-    root}}. The network does not make this mistake, and treats these
-    occurrences of {\it zebra} according to its importance as {\sc nsubj}.
-\item {\it Water} as a modifier often changes the meaning of its head in
- a visually salient way: e.g.\ {\it water fall, water balloon, water scene, water skiing}, and
-  thus the network learns that this particular word is important in the modifier position.
-\end{enumerate}
-
-\subsubsection{Sensitivity to information structure}
+\subsubsection{Sensitivity to linear structure}
 \label{subsec:information-struct}
@@
 adding extra information about the position of words 
-explains more of the variance in case of {\sc Visual} and especially {\sc Textual}.
+explains more of the variance in case of {\sc Visual} and especially
+{\sc Textual} and {\sc LM}.
 Figure \ref{fig:posrqs} shows the coefficients corresponding to the
-position variables in {\sc LM full}. Since the omission scores 
+position variables in {\sc LR~full}. Since the omission scores 
 are measured at the end-of-sentence token, the expectation is that 
-for {\sc Textual}, as a language model,  
+for {\sc Textual} and {\sc LM}, as language models,  
 the words appearing closer to the end of the sentence would have a
 stronger effect on the omission scores. This seems to be confirmed by
-the plot as the coefficients up until the \emph{penultimate} are all negative.
+the plot as the coefficients for these two networks up until the
+\emph{antepenult} are all negative.
 
@@
 tokens in positions closer to the end of the sentence to be more important. 
-On the other hand in English the information structure (or pragmatic structure) 
-of a sentence is expressed via linear ordering: the {\sc topic} of a 
+On the other hand in English the information structure a single
+sentence is 
+expressed via linear ordering: the {\sc topic} of a 
 sentence appears sentence-initially, and the {\sc comment} follows. 
@@
 contain new information in these cases. In our setting of image captions 
-however, it is the {\sc topic} that typically contains the most important 
+however, sentences are not part of larger discourse, it is sentence
+initial material that typically contains the most important 
 objects depicted in the image, e.g. {\it {\underline{two zebras}} are grazing in tall grass on a savannah.}

### 2017-05-30 doc/models.tex

@@
 \section{Models}
+
+We apply our methods to the following models:
+\begin{itemize}
+\item {\sc Imaginet}: A multi-modal GRU network which consists of two
+  pathways, {\sc Visual} and {\sc Textual}, coupled via word
+  embeddings.
+\item {\sc LM}: A (unimodal) language model consisting of a Gated Recurrent
+  Unit (GRU) network.
+\item {\sc Sum}: A network with the same objective as the {\sc Visual}
+  pathway of {\sc Imaginet}, but which uses sum of word embeddings
+  instead of a GRU.
+\end{itemize}
+The rest of this sections gives more details for these models.
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.25]{imaginet.pdf} 
-\caption{Structure of {\sc Imaginet}, taken from Chrupała et al. (2015).}
-\label{fig:imaginet}
-\end{center}
-\end{figure}
-
 {\sc Imaginet} introduced in \namecite{chrupala2015learning} is a
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.2]{imaginet.pdf} 
+\caption{Structure of {\sc Imaginet}, adapted from \protect\cite{chrupala2015learning}.}
+\label{fig:imaginet}
+\end{center}
+\end{figure}
+
 Figure~\ref{fig:imaginet} shows the structure of {\sc Imaginet}. As can
@@
 The inputs to the model are pairs of image descriptions and their 
-corresponding images. The {\sc Textrual} pathway predicts the next 
+corresponding images. The {\sc Textual} pathway predicts the next 
 word at each position in the sequence of words in each caption, whereas the 
@@
 % better in the particular setting.
+
+\subsection{Unimodal language model}
+The model {\sc LM} is a language model analogous to the {\sc Textual}
+pathway of {\sc Imaginet} with the difference that its word embeddings
+are not shared, and its loss functions is the cross-entropy on word
+prediction. Using this model we remove the visual objective as a
+factor, as the model does not use the images corresponding to captions
+in any way.
+
+\subsection{Sum of word embeddings}
+The model {\sc Sum} is a stripped down version of the {\sc Visual}
+pathway, which does not share word embeddings, only uses the cosine
+loss function, and replaces the GRU network with a summation over word
+embeddings. It this removes the effect of word order from
+consideration. We use this model as a baseline in the sections which
+focus on language structure.

### 2017-05-30 src/analysis_mi.R
Mutual info

@@ -1,0 +1,56 @@
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+library(boot)
+
+data <- read.csv("../data/mutual.csv")
+
+
+log_ratio <- function(data, ix) {
+  w <- log(median(filter(data[ix,], pathway=="textual", condition=="word")$mi) / 
+        median(filter(data[ix,], pathway=="visual", condition=="word")$mi),base=exp(1))
+  d <- log(median(filter(data[ix,], pathway=="textual", condition=="dep")$mi) / 
+             median(filter(data[ix,], pathway=="visual", condition=="dep")$mi),base=exp(1))
+  return(c(w=w,d=d))
+}
+
+data_boot <- data.frame()
+for (i in 1:3) {  
+  res <- boot(data %>% filter(order==i), log_ratio, R=5000)
+  data_boot <- rbind(data_boot, data.frame(order=i, condition="word", log.ratio=res$t[,1]))
+  data_boot <- rbind(data_boot, data.frame(order=i, condition="dep", log.ratio=res$t[,2]))
+}
+ggplot(data=data_boot,
+       aes(x=as.factor(order), color=condition, y=log.ratio)) +
+  geom_boxplot() +
+  coord_flip() +
+  xlab("n-gram range") +
+  ylab(expression(log(MI[C]^T / MI[C]^V ))) +
+  theme(text=element_text(size=20)) +
+  ggsave(file="../doc/bootstrappedMI.pdf", width=6, height=4)
+
+
+# LM / Textual
+log_ratio_2 <- function(data, ix) {
+  w <- log(median(filter(data[ix,], pathway=="lm", condition=="word")$mi) / 
+             median(filter(data[ix,], pathway=="textual", condition=="word")$mi),base=exp(1))
+  d <- log(median(filter(data[ix,], pathway=="lm", condition=="dep")$mi) / 
+             median(filter(data[ix,], pathway=="textual", condition=="dep")$mi),base=exp(1))
+  return(c(w=w,d=d))
+}
+
+data_boot_2 <- data.frame()
+for (i in 1:3) {  
+  res <- boot(data %>% filter(order==i), log_ratio_2, R=5000)
+  data_boot_2 <- rbind(data_boot_2, data.frame(order=i, condition="word", log.ratio=res$t[,1]))
[…]

### 2017-05-30 doc/experiments.tex
experiments section edited.

@@
 
-In the following sections, we report a series of experiments in which
-we explore the kinds of linguistic regularities the
-networks learn from
-word-level input. Section~\ref{sec:macro} presents the macro-level
-analyses where we propose methods to analyze the final hidden activation 
-vectors of the recurrent pathways from linguistic point of view. 
-Section~\ref{sec:micro} reports exploratory experiments on
-the linguistic features encoded by individual hidden units. 
-In both sections we report our findings based on the {\sc Imaginet}
-model, discuss the generalizabilty of our methods to other architectures.
-\label{edit:experimentsgeneral}, and whenever appropriate compare to
-our two other models, {\sc LM} and {\sc Sum}.  For all the
-experiments, we trained {\sc Imaginet} on the training portion of the
+In the following section, we report a series of experiments in which
+we explore the kinds of linguistic regularities the networks learn from
+word-level input. Sections~\ref{sec:computeomission} to \ref{sec:beyondlexical} 
+present the analyses of the final hidden activation 
+vectors of the recurrent pathways from a linguistic point of view. 
+Sections~\ref{sec:syntacticdim} and \ref{seq:describe} report exploratory experiments on
+the linguistic features encoded by individual hidden dimensions. 
+In all these experiments we report our findings based on the {\sc Imaginet}
+model, whenever appropriate compare to our two other models {\sc LM} and {\sc Sum}, 
+and discuss the generalizabilty of our methods to other architectures.
+\label{edit:experimentsgeneral}
+For all the experiments, we trained the models on the training portion of the
 MSCOCO image-caption dataset \cite{lin2014microsoft}, and analyzed the

### 2017-05-30 doc/categories.tex
sections 4.1 and 4.2 edited.

@@
-\section{Analysis of hidden activation vectors}
-\label{sec:macro}
+%\section{Analysis of hidden activation vectors}
+%\label{sec:macro}
 
@@
   \includegraphics[scale=0.3]{omissionex.pdf}
-  \caption{Omission scores for an example sentence for {\sc LM} and
+  \caption{Omission scores for the example sentence {\it a baby sits
+      on a bed laughing with a laptop computer open} for {\sc LM} and
     the two pathways, {\sc Textual} and {\sc Visual}, of {\sc
@@
 %\label{sec:salience}
-In this section we propose novel techniques for interpreting the
-activation patterns of neural networks trained on language tasks
-from a linguistic point of view and focus
-on the high level understanding of what parts of input sentence the networks
-pay most attention to. Furthermore, we investigate if the networks
-learn to assign differential amounts of importance to tokens depending
-on their position and grammatical function in the sentences.\label{edit:whyposdep}
-In Section~\ref{sec:computeomission} we introduce \emph{omission scores};
-a metric to measure the contribution of each token to the prediction
-of the networks and in Section~\ref{sec:omitimaginet} we analyze how
-these scores are distributed over dependency relations and
-part-of-speech categories. 
-In Section~\ref{sec:beyondlexical} we investigate the extent to which
-the importance of words for the different networks depend on the words themselves,
-their sequential position and their grammatical function in the sentences.
+%In this section we propose novel techniques for interpreting the
+%activation patterns of neural networks trained on language tasks
+%from a linguistic point of view and focus
+%on the high level understanding of what parts of input sentence the networks
+%pay most attention to. Furthermore, we investigate if the networks
+%learn to assign differential amounts of importance to tokens depending
+%on their position and grammatical function in the sentences.\label{edit:whyposdep}
+%In Section~\ref{sec:computeomission} we introduce \emph{omission scores};
+%a metric to measure the contribution of each token to the prediction
+%of the networks and in Section~\ref{sec:omitimaginet} we analyze how
+%these scores are distributed over dependency relations and
+%part-of-speech categories. 
+%In Section~\ref{sec:beyondlexical} we investigate the extent to which
+%the importance of words for the different networks depend on the words themselves,
+%their sequential position and their grammatical function in the sentences.
 % MAYBE THIS WOULD BE TOO MUCH
@@
 \label{sec:computeomission}
+
+We propose a novel technique for interpreting the
+activation patterns of neural networks trained on language tasks
+from a linguistic point of view, and focus
+on the high-level understanding of what parts of input sentence the networks
+pay most attention to. Furthermore, we investigate if the networks
+learn to assign different amounts of importance to tokens depending
+on their position and grammatical function in the sentences.\label{edit:whyposdep}
 
@@
 \noindent Figure~\ref{fig:omissionex} demonstrates the omission
-measure for the {\sc LM}, {\sc Visual} and {\sc Textual} networks for an
-example captions. The $\mathrm{omission}$ scores are
+scores for the {\sc LM}, {\sc Visual} and {\sc Textual} models for an
+example caption. The omission scores are
 plotted for the
@@
   Visual} for the full
-caption and for the one with the word {\sc baby} omitted. 
+caption and for the one with the word {\it baby} omitted. 
 The images are retrieved from the validation set of MS-COCO by: 1)
@@
 and their bi-directional variants, measuring the contribution
-of tokens to their predictions the $\mathrm{omission}$ 
+of tokens to their predictions (or the omission scores)
 can be straight-forwardly computed using their hidden state 
@@
 can be applied in general to other architectures which
-embed variable length linguistic expressions to the same fixed dimensional
+map variable length linguistic expressions to the same fixed dimensional
 space and perform predictions based on these embeddings. 
@@
 
-The $\mathrm{omission}$ scores can be used not only to
+The omission scores can be used not only to
 estimate the importance of individual words, but also of syntactic
@@
 relative importance of dependency relations: {\sc Visual} pays most
-attention to the relations {\sc nsubjpass, attr, amod, nsubj, ROOT,
+attention to the relations {\sc nsubjpass, attr, amod, nsubj, root,
   compound, dobj, nummod}
@@
 suggest that this setting is not optimal for learning useful representations of verbs,
-which possibly leads to representations that are too task specific 
+which possibly leads to representations that are too task-specific 
 and not transferable across tasks.

### 2017-05-30 doc/categories.tex

@@
 limited by fixed window sizes.\footnote{CNNs with multi-word filters
-and tree-structured recursive neural networks do not incrementaly build representations
+and tree-structured recursive neural networks do not incrementally build representations
 of sentences in a left-to-right or right-to-left fashion. 
-Bi-directional RNNs, are sensitive by word-order and can potentially
+Bi-directional RNNs, on the other hand, are sensitive by word-order and can potentially
 learn to handle the same word in different positions differently. \label{edit:foot}}
@@
 scores per token with the following predictor variables: 
-\begin{enumerate}
+\begin{itemize}
 	\item {\sc LR word}: word type
@@
 	word:position interaction
-\end{enumerate}
+\end{itemize}
 
-\noindent We split the 5-thousand-image portion of MSCOCO validation data into two parts, fit the 
+\noindent We split the 5000-image portion of MSCOCO validation data into two equal parts, fit the 
 models on the first part, and compute the 
@@
                & word   & +pos  & +dep  & full \\\hline
-     sum       & 0.654  & 0.661 & 0.670 & 0.670 \\
-     LM        & 0.358  & 0.586 & 0.415 & 0.601 \\
-     textual   & 0.364  & 0.703 & 0.451 & 0.715 \\
-     visual    & 0.490  & 0.506 & 0.515 & 0.523 \\
+     {\sc Sum}       & 0.654  & 0.661 & 0.670 & 0.670 \\
+     {\sc LM}        & 0.358  & 0.586 & 0.415 & 0.601 \\
+     {\sc Textual}   & 0.364  & 0.703 & 0.451 & 0.715 \\
+     {\sc Visual}    & 0.490  & 0.506 & 0.515 & 0.523 \\
     \end{tabular}
@@
 
+\todo[inline]{I think we can improve the discuss of results below.}
 Table~\ref{tab:lr-r2} shows the proportion of variance $R^2$ in omission
@@
 The results reveal that adding additional information on
-top of lexical features in case of {\sc Sum} does increase the
+top of lexical features in the case of {\sc Sum} does increase the
 explained variance slightly, which is most likely due to the unseen words
@@
 increase in $R^2$ between {\sc LR~+pos} and {\sc LR~full} in
-case of {\sc Visual}, suggesting that 
-the omission scores in case of {\sc Visual} depend on the words'
-grammatical function in sentences, {\it even after controling for word
+the case of {\sc Visual}, suggesting that 
+the omission scores in the case of {\sc Visual} depend on the words'
+grammatical function in sentences, {\it even after controlling for word
 identity and linear position.}
@@
 In order to find out some of the specific syntactic configurations leading to
-increase in $R^2$ between the {\sc LR~word} and {\sc LR~dep} predictors
-in case of {\sc Visual}, we next considered all word types with
+an increase in $R^2$ between the {\sc LR~word} and {\sc LR~+dep} predictors
+in the case of {\sc Visual}, we next considered all word types with
 occurrence counts of at least 100 and ranked them according to how much
@@
 Figure~\ref{fig:top_words} shows the per-dependency 
-omission-score distributions for seven top ranked words.
+omission score distributions for seven top-ranked words.
 There are clear and large differences in how these words
@@
 adding extra information about the position of words 
-explains more of the variance in case of {\sc Visual} and especially
+explains more of the variance in the case of {\sc Visual} and especially
 {\sc Textual} and {\sc LM}.

### 2017-05-30 doc/experiments.tex

@@
 
-In the following section, we report a series of experiments in which
+In this section, we report a series of experiments in which
 we explore the kinds of linguistic regularities the networks learn from
-word-level input. Sections~\ref{sec:computeomission} to \ref{sec:beyondlexical} 
-present the analyses of the final hidden activation 
-vectors of the recurrent pathways from a linguistic point of view. 
-Sections~\ref{sec:syntacticdim} and \ref{seq:describe} report exploratory experiments on
-the linguistic features encoded by individual hidden dimensions. 
+word-level input. 
+%Sections~\ref{sec:computeomission} to \ref{sec:beyondlexical} 
+%present the analyses of the final hidden activation 
+%vectors of the recurrent pathways from a linguistic point of view. 
+%Sections~\ref{sec:syntacticdim} and \ref{seq:describe} report exploratory experiments on
+%the linguistic features encoded by individual hidden dimensions. 
+In Section~\ref{sec:computeomission} we introduce \emph{omission scores},
+a metric to measure the contribution of each token to the prediction
+of the networks and in Section~\ref{sec:omitimaginet} we analyze how
+these scores are distributed over dependency relations and
+part-of-speech categories. 
+In Section~\ref{sec:beyondlexical} we investigate the extent to which
+the importance of words for the different networks depend on the words themselves,
+their sequential position and their grammatical function in the sentences.
+\todo[inline]{add a short description of the experiments in sections 5.4 and 5.5.}
+
 In all these experiments we report our findings based on the {\sc Imaginet}

### 2017-05-30 doc/intro.tex

@@
 
+\todo[inline]{once we finalise our experiments, we should update the following paragraphs and get rid of the macro/micro distinction.}
 We report a thorough \emph{macro level} quantitative analysis to

### 2017-05-30 doc/models.tex

@@
 
-We apply our methods to the following models:
+In our analyses of the acquired linguist knowledge, we apply our methods to the following models:
 \begin{itemize}
-\item {\sc Imaginet}: A multi-modal GRU network which consists of two
-  pathways, {\sc Visual} and {\sc Textual}, coupled via word
-  embeddings.
-\item {\sc LM}: A (unimodal) language model consisting of a Gated Recurrent
-  Unit (GRU) network.
+\item {\sc Imaginet}: A multi-modal Gated Recurrent Unit (GRU) network consisting of two
+  pathways, {\sc Visual} and {\sc Textual}, coupled via word embeddings.
+\item {\sc LM}: A (unimodal) language model consisting of a GRU network.
 \item {\sc Sum}: A network with the same objective as the {\sc Visual}
@@
 \end{itemize}
-The rest of this sections gives more details for these models.
+The rest of this section gives a detailed description of these models.
 
@@
 information over multiple time steps.
-Specifically, the GRU computes the hidden state at current time step, $\mathbf{h}_{t}$, as the
+Specifically, the GRU computes the hidden state at current time step $\mathbf{h}_{t}$, as the
 linear combination of previous activation $\mathbf{h_{t-1}}$, and a new
@@
 compared to the original {\sc Imaginet}. As the performance of
-{\sc Visual} is measured on image-retrieval, which is based on cosine 
-distances we use cosine distance as the visual loss.
+{\sc Visual} is measured on image-retrieval which is based on cosine 
+distances, we use cosine distance as the visual loss.
 Moreover, we observe that using standardized image vectors, where each dimension 
-is transformed by subtracting the mean and dividing by standard deviation further
+is transformed by subtracting the mean and dividing by standard deviation, further
 improves performance.
@@
 pathway of {\sc Imaginet} with the difference that its word embeddings
-are not shared, and its loss functions is the cross-entropy on word
+are not shared, and its loss function is the cross-entropy on word
 prediction. Using this model we remove the visual objective as a
@@
 \subsection{Sum of word embeddings}
-The model {\sc Sum} is a stripped down version of the {\sc Visual}
+The model {\sc Sum} is a stripped-down version of the {\sc Visual}
 pathway, which does not share word embeddings, only uses the cosine
 loss function, and replaces the GRU network with a summation over word
-embeddings. It this removes the effect of word order from
+embeddings. This removes the effect of word order from
 consideration. We use this model as a baseline in the sections which

### 2017-05-30 src/analysis_mi.R
Minor edits in analysis_mi.R

@@ -4,7 +4,10 @@ library(reshape2)
 library(boot)
 
 data <- read.csv("../data/mutual.csv")
+ggplot(data, aes(x=pathway, y=mi, color=paste(condition, order))) + geom_boxplot() +
+  ggsave("../doc/raw_mutual.pdf")
 
+# Textual vs Visual
 
 log_ratio <- function(data, ix) {
   w <- log(median(filter(data[ix,], pathway=="textual", condition=="word")$mi) / 
@@ -14,6 +17,7 @@ log_ratio <- function(data, ix) {
   return(c(w=w,d=d))
 }
 
+
 data_boot <- data.frame()
 for (i in 1:3) {  
   res <- boot(data %>% filter(order==i), log_ratio, R=5000)
@@ -22,11 +26,12 @@ for (i in 1:3) {
 }
 ggplot(data=data_boot,
        aes(x=as.factor(order), color=condition, y=log.ratio)) +
-  geom_boxplot() +
+  geom_boxplot(notch = TRUE) +
   coord_flip() +
   xlab("n-gram range") +
   ylab(expression(log(MI[C]^T / MI[C]^V ))) +
   theme(text=element_text(size=20)) +
+  theme(legend.title=element_blank()) +
   ggsave(file="../doc/bootstrappedMI.pdf", width=6, height=4)
 
 
@@ -47,10 +52,12 @@ for (i in 1:3) {
 }
 ggplot(data=data_boot_2,
        aes(x=as.factor(order), color=condition, y=log.ratio)) +
-  geom_boxplot() +
+  geom_boxplot(notch = TRUE) +
   coord_flip() +
   xlab("n-gram range") +
   ylab(expression(log(MI[C]^LM / MI[C]^T ))) +
   theme(text=element_text(size=20)) +
+  theme(legend.title=element_blank()) +
   ggsave(file="../doc/bootstrappedMI2.pdf", width=6, height=4)
 
+data %>% filter(pathway=="textual", condition=="dep") %>% arrange(desc(mi)) %>% head

### 2017-05-30 doc/categories.tex

@@
 they share their contribution with the first, often more important,
-member of the conjunction, for example: {\it A cow and its baby eating
+member of the conjunction, for example in {\it A cow and its baby eating
   grass}. 
@@
  \includegraphics[scale=0.4]{position-coef.pdf}
- \caption{Coefficients on the y-axis of {\sc LM full} corresponding to the
+ \caption{Coefficients on the y-axis of {\sc LR full} corresponding to the
 position variables on the x-axis.}
@@
 tokens in positions closer to the end of the sentence to be more important. 
-On the other hand in English the information structure a single
-sentence is 
-expressed via linear ordering: the {\sc topic} of a 
+On the other hand in English the information structure of a single
+sentence is expressed via linear ordering: the {\sc topic} of a 
 sentence appears sentence-initially, and the {\sc comment} follows. 

### 2017-05-31 src/mutual.py
MI

@@ -22,20 +22,31 @@ def main():
     ids = np.array([senti['sentid'] for senti in prov.iterSentences(split='val')])
     data = pd.read_csv("/home/gchrupal/cl-resubmit/data/depparse_coco_val.csv")
     sent_data = dict(senti for senti in data.groupby(['sentid']))
-    with open("/home/gchrupal/cl-resubmit/data/mutual.csv", "w") as out:
-        writer = csv.writer(out)
-        writer.writerow(["pathway", "condition", "order", "dimension", "mi"])
-        for net in ["visual", "textual", "lm", "sum"]:
-            states = np.load("/home/gchrupal/cl-resubmit/data/states_{}.npy".format(net), encoding='bytes')
-            S = quantize(np.vstack([state[:-1,:] for state in states]).T)
-            for typ in ['word', 'dep']:
-                for order in [1,2,3]:
-                    logging.info("Computing scores for {}:{}:{}".format(net, typ, order))
-                    context = [item for i in ids for item in ngram(list(sent_data[i][typ]), order) ]
-                    #print(context[:10])
-                    for i in range(1024):
-                        mi = mutual_info_score(S[i], context)
-                        writer.writerow([net, typ, order, i, mi])
+    with open("/home/gchrupal/cl-resubmit/data/mutual_examples.csv", "w") as out_ex:
+        writer_ex = csv.writer(out_ex)
+        writer_ex.writerow(["pathway", "condition", "order", "dimension", "index","history","history_w"])
+        with open("/home/gchrupal/cl-resubmit/data/mutual.csv", "w") as out:
+            writer = csv.writer(out)
+            writer.writerow(["pathway", "condition", "order", "dimension", "mi"])
+            for net in ["visual", "textual", "lm", "sum"]:
+                states = np.load("/home/gchrupal/cl-resubmit/data/states_{}.npy".format(net), encoding='bytes')
+                S = quantize(np.vstack([state[:-1,:] for state in states]).T)
+                for typ in ['word', 'dep']:
+                    for order in [1,2,3]:
+                        logging.info("Computing scores for {}:{}:{}".format(net, typ, order))
+                        context_w = [item for i in ids for item in ngram(list(sent_data[i]['word']), order) ]
+                        context = [item for i in ids for item in ngram(list(sent_data[i][typ]), order) ]
+                        mi = []
+                        for i in range(1024):
+                            mi.append(mutual_info_score(S[i], context))
[…]

### 2017-05-31 doc/intro.tex
Models and intro

@@
 trained to predict image vectors given image descriptions, while the other
-pathway ({\sc Textual}) is a language model. This particular
+pathway ({\sc Textual}) is a language model, trained to sequentially predict each
+word in the descriptions. This particular
 architecture allows a comparative analysis of the hidden activations
-patterns between networks trained on two very different tasks, while
+patterns between networks trained on two different tasks, while
 keeping the training data and the word embeddings fixed. Recurrent neural
@@
 view of representing meaning conveyed via linguistic structure.
+For comparison, we also consider a standard standalone language model.
 
@@
 and Convolutional Neural Networks.\label{edit:introomissiongeneral}
-Based on these omission scores our experiments show that the image prediction 
-pathway in general pays special attention to words of syntactic categories that 
-carry semantic content. Furthermore, we observe that on top of
-lexical features it learns important aspects of the information structure of sentences,
-and treats the same input tokens differently depending on their
-grammatical function in the sentence. In contrast, the language model
-is more sensitive to the local syntactic characteristics of the input
-sentences.
+
+Based on these omission scores our experiments show that the image
+prediction pathway in general pays special attention to syntactic
+categories which carry semantic content, and especially nouns. More
+surprisingly, this pathway also learns learns to treat word types
+differently depending on their grammatical function and their position
+in the sequential structure of the sentence. In contrast, the textual
+pathway and the standalone language model are especially sensitive to
+the local syntactic characteristics of the input sentences.
 

### 2017-05-31 doc/models.tex
Models and intro

@@
 and cosine distance $L^V$ for the image
-predictions.
+predictions\footnote{Note that the original formulation in
+  \namecite{chrupala2015learning} uses mean squared error instead; as the performance of
+{\sc Visual} is measured on image-retrieval which is based on cosine 
+distances, we use cosine distance as the visual loss here.}, weighting them with the parameter $\alpha$ (set to $0.1$).
 %
@@
 \noindent
-For more details about the {\sc Imaginet} model and 
-its performance see \namecite{chrupala2015learning}.
-
-In our version of the model we add two minor modifications 
-compared to the original {\sc Imaginet}. As the performance of
-{\sc Visual} is measured on image-retrieval which is based on cosine 
-distances, we use cosine distance as the visual loss.
-Moreover, we observe that using standardized image vectors, where each dimension 
-is transformed by subtracting the mean and dividing by standard deviation, further
-improves performance.
+For more details about the {\sc Imaginet} model and its performance
+see \namecite{chrupala2015learning}. Note that we introduce a small change in
+the image representation: we observe that using standardized image
+vectors, where each dimension is transformed by subtracting the mean
+and dividing by standard deviation, improves performance.
 

### 2017-05-31 doc/categories.tex
discussion of results in section 4.3 expanded.

@@
 may actually learn to pay more attention to the same word type in appropriate
-contexts. This section investigates to what extent {\sc Visual} and {\sc Textual}
+contexts. This section investigates to what extent our models
 discriminate between occurrences of a given word in different positions and 
@@
 scores per token with the following predictor variables: 
-\begin{itemize}
+\begin{enumerate}
 	\item {\sc LR word}: word type
@@
 	word:position interaction
-\end{itemize}
+\end{enumerate}
 
-\noindent We split the 5000-image portion of MSCOCO validation data into two equal parts, fit the 
-models on the first part, and compute the 
-proportion of variance explained on the second part. 
+\noindent We use the 5000-image portion of MSCOCO validation data for 
+training and test. The captions contain about 260,000 words in total, of 
+which we use 100,000 to fit the regression models. We then use the rest 
+of the words to compute the proportion of variance explained by the models. 
 For comparison we also use the {\sc Sum} model which composes word
@@
 
-\todo[inline]{I think we can improve the discuss of results below.}
 Table~\ref{tab:lr-r2} shows the proportion of variance $R^2$ in omission
 scores explained by the linear regression with the different predictors.
-Figure~\ref{fig:rsquared} offers a different view of the data, showing
-the increase or decrease in $R^2$ for the models relative to {\sc
-  LR~+pos}.  
-The results reveal that adding additional information on
-top of lexical features in the case of {\sc Sum} does increase the
-explained variance slightly, which is most likely due to the unseen words
-in the held out set. More interestingly, there is a sizeable
-increase in $R^2$ between {\sc LR~+pos} and {\sc LR~full} in
-the case of {\sc Visual}, suggesting that 
-the omission scores in the case of {\sc Visual} depend on the words'
-grammatical function in sentences, {\it even after controlling for word
-identity and linear position.}
-
-The raw $R^2$ scores show that for the {\sc Textual} and {\sc LM}
-models the word-type predicts the omission-score to much smaller
-degree compared to {\sc Visual}, and the increase in $R^2$ shows that
-the position of the word adds considerable amount of information for
-these language models: this is not surprising considering that the omission
+The raw $R^2$ scores show that for the language models {\sc LM} and 
+{\sc Textual}, the word-type predicts the omission-score to much smaller 
+degree compared to {\sc Visual}. Moreover, adding information about 
+either the position or the dependency labels increases the explained variance for all models. 
+However, for the {\sc Textual} and {\sc LM} models the position of the word adds 
+considerable amount of information. This is not surprising considering that the omission
 scores are measured with respect to the final activation state, and
 given the fact that in a language model the recent history is most
-important for accurate prediction. Less expectedly, for the {\sc Visual} model,
-dependency labels are more informative than linear position,
-hinting at the importance of syntactic structure beyond linear order.
+important for accurate prediction.
+
+Figure~\ref{fig:rsquared} offers a different view of the data to, showing
+the increase or decrease in $R^2$ for the models relative to {\sc LR~+pos} 
+to emphasise the importance of syntactic structure beyond the position in the sentence.
+Interestingly, for the {\sc Visual} model, dependency labels are 
+more informative than linear position, hinting at the importance of syntactic 
+structure beyond linear order. There is a sizeable increase in $R^2$ between 
+{\sc LR~+pos} and {\sc LR~full} in the case of {\sc Visual}, suggesting that 
+the omission scores for {\sc Visual} depend on the words'
+grammatical function in sentences, {\it even after controlling for word
+identity and linear position.}  In contrast, adding additional information on
+top of lexical features in the case of {\sc Sum} increases the
+explained variance only slightly, which is most likely due to the unseen words
+in the held out set.
 
@@
 hardest to predict of the four models suggesting that the model may be
-encoding additional structural features not captured by these
-predictors.
+encoding additional structural features not captured by these predictors. 
+We will look more deeply into such potential features in the following sections.
 

### 2017-05-31 doc/response.tex
response letter updated.

@@
 
-We have revised parts of the article and have made clarifications
-throughout the text in response to yours and the reviewers' comments.
-Specifically with regard to the two main issues you raised, we have...
+We have substantially revised the article in response to yours and the reviewers' 
+comments. Specifically with regard to the two main issues you raised, we have 
+introduced a uni-modal language model (referred to as the {\sc LM} model in the text) 
+and reported the results of all the experiments on this model as well. Also following 
+your recommendation, we have removed the exploratory parts of (the old) Section 6, 
+and reorganized the paper by merging previous Sections 4, 5 and the remaining of 
+Section 6 into a single section (Section 4 in the revised manuscript) reporting all the 
+experiments. We have kept and expanded the quantitative analysis of the nature of 
+n-gram contexts that trigger individual dimensions in the hidden layers; this experiment is now reported in Section~\ref{sec:contexts}.
+
+Furthermore, we have made clarifications about the details of the techniques all 
+through the manuscript, expanded Introduction (Section~\ref{sec:intro}) to 
+emphasise our contributions, updated Related Work (Section~\ref{sec:related}) 
+to include relevant papers that came out while this manuscript was under review, and expanded the Discussion
+
 \todo[inline]{explain the major revisions}

### 2017-05-31 doc/response.tex
response letter updated.

--- 2017-05-29 doc/response.tex
@@
 
-We have revised parts of the article and have made clarifications
-throughout the text in response to yours and the reviewers' comments.
-Specifically with regard to the two main issues you raised, we have...
-\todo[inline]{explain the major revisions}
+We have substantially revised the article in response to your and the reviewers' 
+comments. Specifically with regard to the two main issues you raised, we have 
+introduced a uni-modal language model (referred to as the {\sc LM} model in the text) 
+and reported the results of all the experiments on this model as well. Also following 
+your recommendation, we have removed the exploratory parts of (the old) Section 6, 
+and reorganized the paper by merging previous Sections 4, 5 and the remaining of 
+Section 6 into a single section (Section 4 in the revised manuscript) containing all the 
+experiments. We have kept and expanded the quantitative analysis of the nature of 
+n-gram contexts that trigger individual dimensions in the hidden layers; this experiment is now reported in Section~\ref{sec:contexts}.
+
+Furthermore, we have made clarifications about the details of the techniques all 
+through the manuscript, expanded Introduction (Section~\ref{sec:intro}) to 
+emphasise our contributions, updated Related Work (Section~\ref{sec:related}) 
+to include relevant papers that came out while this manuscript was under review, and expanded the Discussion to discuss in more detail the relevance and generalizability of our techniques.
 

### 2017-05-31 doc/response.tex
response letter updated

@@
 \begin{quote}
-\textsc{Author Response:} \todo[inline]{This needs explanation and I don't have an answer (yet).}
+\textsc{Author Response:} We removed this section in the revised manuscript. We address the nature of contexts which trigger individual dimensions in more detail now in Section~\ref{sec:contexts}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  we have reworded `positive' and `negative' as `grammatical' and `ungrammatical' to avoid confusion.
+\textsc{Author Response:}  we have reworded `positive' and `negative' as `grammatical' 
+and `ungrammatical' to avoid confusion.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{good point! what is the value of lambda?}
+\textsc{Author Response:}  There was no $\lambda$ in that equation, but in 
+Section{sec:imaginer} we now report the value we use for the parameter 
+$\alpha$ in Equation~(\ref{eq:losscombo}).
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{Grzegorz?}
+\textsc{Author Response:}  You are right, the description of the model in Section~\ref{sec:imaginet} presents the version used in this paper, which uses cosine distance.
+The original Imaginet model uses mean square error instead; we clarified this issue in footnote~\ref{ft:imaginet}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{specify interaction features}
+\textsc{Author Response:}  We have specified the interaction features used in each regression model in Section~\ref{sec:beyond lexical}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{explain}
+\textsc{Author Response:}  We have removed the {\it water} from the set of examples in Figure~\ref{fig:top_words}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{address}
+\textsc{Author Response:}  Due to space and time limitations we have to 
+leave the comparison with other architectures for future work; however we 
+have added a section to Discussion to discuss the generalizability of our 
+techniques to other architectures in more detail.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{explain new experiments}
+\textsc{Author Response:}  In the revised manuscript we have introduced a uni-modal language model, and updated all are experiments to include this model as well.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{hmm.}
+\textsc{Author Response:}  As we mentioned before, we are very interested in 
+applying our proposed techniques to other tasks and architectures in future work. 
+For the sake of this manuscript, we have included a theoretical discussion of how these techniques can be implemented for a number of other architectures.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{explain the rewritten version.}
+\textsc{Author Response:}  Our explanation of the aim of this experiment in the original version might have been misleading; we have rewritten this section to make it clear that we are examining the information structure within a single sentence and not in a dialogue.
+We do believe that the results we report in this section are interesting and nontrivial, especially the fact that the position usually occupied by the subject of the sentence is the most important for the {\sc Visual} model despite the common recency bias in RNNs.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{address.}
+\textsc{Author Response:}  We have removed the exploratory parts of (the old) Section 6, and expanded the quantitative analysis of the nature of triggering contexts in Section~\ref{sec:contexts}. 
 \end{quote}

### 2017-05-31 doc/models.tex
response letter updated

@@
 {\sc Visual} is measured on image-retrieval which is based on cosine 
-distances, we use cosine distance as the visual loss here.}, weighting them with the parameter $\alpha$ (set to $0.1$).
+distances, we use cosine distance as the visual loss here.\label{ft:imaginet}}, weighting them with the parameter $\alpha$ (set to $0.1$).
 %
@@
 %\begin{equation}
-%\label{eq:losscombo}
+\label{eq:losscombo}
 &L = \alpha L^T + (1-\alpha)L^{V}

### 2017-05-31 src/analysis_mi.R
Raw MI

@@ -4,8 +4,12 @@ library(reshape2)
 library(boot)
 
 data <- read.csv("../data/mutual.csv")
-ggplot(data, aes(x=pathway, y=mi, color=paste(condition, order))) + geom_boxplot() +
-  ggsave("../doc/raw_mutual.pdf")
+ggplot(data %>% filter(pathway != "sum"), aes(x=pathway, y=mi, color=paste(condition, order))) + geom_boxplot() +
+  xlab(NULL) +
+  ylab("Mutual Information") +
+  theme(legend.title=element_blank()) +
+  theme(text=element_text(size=20)) 
+  ggsave("../doc/raw_mutual.pdf",width=6, height=4)
 
 # Textual vs Visual
 

### 2017-05-31 doc/response.tex
response letter updated

@@
 and reorganized the paper by merging previous Sections 4, 5 and the remaining of 
-Section 6 into a single section (Section 4 in the revised manuscript) containing all the 
-experiments. We have kept and expanded the quantitative analysis of the nature of 
-n-gram contexts that trigger individual dimensions in the hidden layers; this experiment is now reported in Section~\ref{sec:contexts}.
+Section 6 into a single section containing all the experiments (that is Section 4 in the 
+revised manuscript). We have kept and expanded the quantitative analysis of the nature of 
+n-gram contexts that trigger individual dimensions in the hidden layers; this experiment is 
+now reported in Section~\ref{sec:contexts}.
 
@@
 emphasise our contributions, updated Related Work (Section~\ref{sec:related}) 
-to include relevant papers that came out while this manuscript was under review, and expanded the Discussion to discuss in more detail the relevance and generalizability of our techniques.
+to include relevant papers that came out while this manuscript was under review, and expanded Discussion (Section~\ref{sec:conclusion}) to discuss in more detail the relevance and generalizability of our techniques.
 
@@
 \textsc{Author Response:}  Following the recommendation of the reviewers,
-we shortened Section~\ref{sec:micro} and removed this figure.
+we shortened the original Section 6 and removed this figure.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  It is in fact the absolute value of the activations 
-that is taken into account. We have changed the description of the top-k-contexts 
-method to clear this issue.
+\textsc{Author Response:}  It was in fact the absolute value of the activations 
+that was taken into account. In the revised manuscript, we have removed this 
+and other exploratory experiments and expanded the quantitative analyses.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  A unit is one neuron in the final hidden layer 
+\textsc{Author Response:}  A unit is one individual dimension in the final hidden layer 
 of the network. We have made this clear in the revised text.
@@
 \begin{quote}
-\textsc{Author Response:}  We removed that sentence and added an expanded discussion of the generalizability of our techniques to the last section.
+\textsc{Author Response:}  We removed this part and added an expanded discussion of the generalizability of our techniques to Discussion (Section~\ref{sec:conclusion}).
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:} We removed this section in the revised manuscript. We address the nature of contexts which trigger individual dimensions in more detail now in Section~\ref{sec:contexts}.
+\textsc{Author Response:} We have removed this section in the revised manuscript. We 
+address the nature of contexts which trigger individual dimensions and a systematic
+procedure for selecting the dimensions themselves in more detail now in 
+Section~\ref{sec:contexts}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  We have reorganized Section~\ref{sec:micro} and 
-tried to motivate this particular experiment better. The idea is to 
-come up with techniques for singling out units with a specific type
-of function, such as encoding grammatical function. Therefore we 
-use the coefficients of the logistic regression model as pointers 
-to units of this type.
+\textsc{Author Response:}  As mentioned before, we have reorganised 
+the experiments and remove the exploratory parts, including the old Section 6.3. 
+We address the issue of encoding grammatical functions in the revised manuscript in 
+Sections ~\ref{sec:beyondlexical} (specifically in \ref{sec:gramfunc}).
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:} It was there, believe us! But we have reorganized 
-Section~\ref{sec:micro} and removed this experiment for the sake of clarity.
+\textsc{Author Response:} It was there, believe us! But we have removed this 
+experiment from the revised version for the sake of clarity and coherence of our analyses.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  We hope we have addressed all your comments and concerned.
-We did not remove the experiments in Section 6.3, but as we mentioned before we rewrote and restructured Section~\ref{sec:micro} and tried to put this experiment in context and clear our motivation.
+\textsc{Author Response:}  We hope we have addressed all your comments and concerns. We have removed the experiments in (the old) Section 6.3 from the revised manuscript.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:} the typo and the citation inconsistencies are fixed.
+\textsc{Author Response:} The typo and the citation inconsistencies are fixed.
 \end{quote}
@@
 \textsc{Author Response:}  You are absolutely right about the description of 
-the top-k-context method being confusing. We have rewritten this part to make the description of the method more clear. Furthermore, we have reorganized 
-Section~\ref{sec:micro} to make it more coherent. 
+top k contexts being confusing. We have removed this part and reorganized 
+the experiments section (Section~\ref{sec:experiments}) to make it more understandable
+and coherent. 
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  As mentioned above, we have rewritten Section~\ref{sec:micro} to make the description of the methods and the experiments more clear.
+\textsc{Author Response:}  As mentioned above, we have removed this part and reorganized Section~\ref{sec:experiments} to make our analyses more concrete and clear.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{update graphs?}
+\textsc{Author Response:}  We have reformatted many of our figures to make them more readable.
 \end{quote}
@@
 \textsc{Author Response:}  There was no $\lambda$ in that equation, but in 
-Section{sec:imaginer} we now report the value we use for the parameter 
+Section~\ref{sec:imaginet} we report the value we use for the parameter 
 $\alpha$ in Equation~(\ref{eq:losscombo}).
@@
 \begin{quote}
-\textsc{Author Response:}  We have included an illustration of the structure of the model in Figure~\ref{fig:imaginet}.
+\textsc{Author Response:}  We have included an illustration of the structure of the 
+model in Figure~\ref{fig:imaginet}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  You are right, the description of the model in Section~\ref{sec:imaginet} presents the version used in this paper, which uses cosine distance.
-The original Imaginet model uses mean square error instead; we clarified this issue in footnote~\ref{ft:imaginet}.
+\textsc{Author Response:}  You are right, the description of the model in 
+Section~\ref{sec:imaginet} presents the version used in this paper, which uses 
+cosine distance. The original Imaginet model uses mean square error instead; 
+we clarified this issue in footnote~\ref{ft:imaginet}.  
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  We have specified the interaction features used in each regression model in Section~\ref{sec:beyond lexical}.
+\textsc{Author Response:}  We have specified the interaction features used in each regression model in Section~\ref{sec:beyondlexical}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  We have removed the {\it water} from the set of examples in Figure~\ref{fig:top_words}.
+\textsc{Author Response:}  We have removed {\it water} from the set of examples 
+in Figure~\ref{fig:top_words}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  we have fixed these issues, thanks for bringing them to our attention.
+\textsc{Author Response:}  We have fixed these issues, thanks for bringing them to our attention.
 \end{quote}
@@
 (Section~\ref{sec:related}) with a discussion of Linzen et al. (2016) 
-and a few other articles that came out while our submission was under review.
+and a few other relevant articles that came out while our submission was under review.
 \end{quote}
@@
 leave the comparison with other architectures for future work; however we 
-have added a section to Discussion to discuss the generalizability of our 
-techniques to other architectures in more detail.
+have expanded the discussion of the generalizability of our techniques to other
+architectures in Discussion (Section~\ref{sec:conclusion}).
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  In the revised manuscript we have introduced a uni-modal language model, and updated all are experiments to include this model as well.
+\textsc{Author Response:}  In the revised manuscript we have introduced a uni-modal language model, and updated all of our experiments to include this model as well.
 \end{quote}
@@
 applying our proposed techniques to other tasks and architectures in future work. 
-For the sake of this manuscript, we have included a theoretical discussion of how these techniques can be implemented for a number of other architectures.
+For the sake of this manuscript, we have included a theoretical discussion of how 
+these techniques can be implemented for a number of other architectures (see 
+Section~\ref{sec:conclusion}).
 \end{quote}

### 2017-05-31 doc/categories.tex
Boxplot explanation.

@@
 Both POS tagging and dependency parsing are performed 
-using the  dependency parser from the Spacy package.\footnote{Available at
+using the  \verb+en_core_web_md+ dependency parser from the Spacy package.\footnote{Available at
   \url{https://spacy.io/}.} 
@@
 scores per POS and dependency label for the two pathways of {\sc
-  Imaginet} and for {\sc LM}.  The general trend is that for the {\sc
+  Imaginet} and for {\sc LM}.\footnote{The boxplots in this and
+  subsequent figures are Tukey boxplots and should be interpreted as follows: the box extends
+from the 25th to the 75th percentile of the data; the line across the
+box is the 50th percentile, while the whiskers extend past the lower
+and upper quartile to $1.5\times$
+the interquartile range (i.e.\ 75th percentile - 25th percentile); the
+points are outliers.}  The general trend is that for the {\sc
   Visual} pathway, the omission scores are high for a small subset of

### 2017-05-31 doc/intro.tex
intro updated

@@
 
+\todo[inline]{revise abstract}
   We present novel methods for analyzing the activation patterns of
@@
 
-\todo[inline]{once we finalise our experiments, we should update the following paragraphs and get rid of the macro/micro distinction.}
-We report a thorough \emph{macro level} quantitative analysis to
-provide a linguistic interpretation of the networks' activation
-patterns. These experiments rely on a novel method we
-call \emph{omission score} to measure the importance of input tokens 
-to the final prediction of models that compute distributed representations 
-of sentences. This can be applied to various RNN architectures, Recursive Neural Networks 
-and Convolutional Neural Networks.\label{edit:introomissiongeneral}
+We report a thorough quantitative analysis to provide a linguistic 
+interpretation of the networks' activation patterns. We present a 
+series of experiments using a novel method we call \emph{omission score} 
+to measure the importance of input tokens to the final prediction of models 
+that compute distributed representations of sentences. 
+Furthermore, we introduce a more global measure for estimating the 
+informativeness of various types of n-gram contexts for each model.
+These techniques can be applied to various RNN architectures, Recursive 
+Neural Networks and Convolutional Neural Networks.
 
-Based on these omission scores our experiments show that the image
-prediction pathway in general pays special attention to syntactic
-categories which carry semantic content, and especially nouns. More
-surprisingly, this pathway also learns learns to treat word types
-differently depending on their grammatical function and their position
-in the sequential structure of the sentence. In contrast, the textual
-pathway and the standalone language model are especially sensitive to
-the local syntactic characteristics of the input sentences.
+Our experiments show that the image prediction pathway in general 
+pays special attention to syntactic categories which carry semantic 
+content, and particularly nouns. More surprisingly, this pathway also 
+learns to treat word types differently depending on their 
+grammatical function and their position in the sequential structure of 
+the sentence. In contrast, the textual pathway and the standalone 
+language model are especially sensitive to the local syntactic 
+characteristics of the input sentences.
+\todo[inline]{say something about the findings of the last experiment.}
 
@@
 
-We also perform a \emph{micro level} analysis to explore \label{edit:introtopkgeneral}
-the function of individual hidden units by introducing a method we dub \emph{top-k-contexts}.
-It involves identifying the sentential contexts which yield the 
-highest activation values and the analysis we present is applicable 
-to uni-directional RNN architectures.
-Through a qualitative examination of these contexts we
-observe that in case of {\sc Imaginet} some of the contexts represent a particular 
-syntactic category or dependency function, while others correspond to a semantic theme. 
-We identify cases where the hidden units encode characteristics of 
-the context that go beyond their lexical properties, and represent abstract 
-patterns that are useful for the network's task. 
-Furthermore, we explore and visualize units that carry their
-activation over to later time steps to 
-extract longer dependencies and more complex linguistic features. 
-We also quantitatively show that features encoded by the language model are more associated 
-with syntactic constructions than in case of the image prediction pathway.
+%We also perform a \emph{micro level} analysis to explore \label{edit:introtopkgeneral}
+%the function of individual hidden units by introducing a method we dub \emph{top-k-contexts}.
+%It involves identifying the sentential contexts which yield the 
+%highest activation values and the analysis we present is applicable 
+%to uni-directional RNN architectures.
+%Through a qualitative examination of these contexts we
+%observe that in case of {\sc Imaginet} some of the contexts represent a particular 
+%syntactic category or dependency function, while others correspond to a semantic theme. 
+%We identify cases where the hidden units encode characteristics of 
+%the context that go beyond their lexical properties, and represent abstract 
+%patterns that are useful for the network's task. 
+%Furthermore, we explore and visualize units that carry their
+%activation over to later time steps to 
+%extract longer dependencies and more complex linguistic features. 
+%We also quantitatively show that features encoded by the language model are more associated 
+%with syntactic constructions than in case of the image prediction pathway.
 

### 2017-05-31 doc/response.tex
why imaginet

@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{address}
+\textsc{Author Response:}  The motivation for using {\sc Imaginet} is laid
+out in the introduction. The main point is that this model consists of
+two loosely coupled pathways: One ({\sc Textual}) is a language
+model which is a type of network widely used and reasonably well
+understood. The other one ({\sc Visual}) is a visually-grounded encoder,
+a much less commonly used and studied type of network. By using {\sc
+  Imaginet} we can apply our methods to these two network types and
+compare the results head-to-head. Thus {\sc Textual} serves
+mostly to show that our methods give reasonably, expected answers,
+while with {\sc Visual} we can generate novel, previously unfamiliar
+insights. The multitask setup is not strictly
+necessary for this; in order to abstract away from the effect of multitask
+training on {\sc Textual} we have now added experiments on a
+standalone language model {\sc LM}.
+
+%\todo[inline]{address}
 \end{quote}

### 2017-05-31 src/analysis.R

@@ -137,6 +137,7 @@ top_words_er <- read.csv("../data/top7_words_er.csv") %>%
 ggplot(top_words_er %>% mutate(wordxdep=paste(word,dep)) %>% filter_by_count("wordxdep",5) %>% filter_by_count("dep", 70),
        aes(x=word, y=omission_v, color=dep)) +
   geom_boxplot(position = "dodge") +
+  theme(legend.title=element_blank()) +
   theme(text=element_text(size=25)) +
   ylab("omission score")
 ggsave(file="../doc/top_words.pdf", width=12, height=8)

### 2017-05-31 doc/intro.tex

@@
 \begin{abstract}
-
-\todo[inline]{revise abstract}
   We present novel methods for analyzing the activation patterns of
@@
   representations of the visual scene corresponding to an input
-  sentence, and predicting the next word in the same sentence. 
+  sentence, and predicting the next word in the same sentence, 
+  in addition to a standard standalone language model. 
   We propose a method for estimating the amount of contribution
   of individual tokens in the input to the final prediction of the networks.   
-  We show that the image prediction pathway: a) is sensitive to
-  the information structure of the sentence, b) pays
+  Using this method, we show that the image prediction pathway pays
   selective attention to lexical categories and grammatical functions
-  that carry semantic information, and c) learns to treat the same
-  input token differently depending on its grammatical functions in
-  the sentence. In contrast the language model is comparatively more
-  sensitive to words with a syntactic function. 
-  Furthermore, we propose methods to explore the function
-  of individual hidden units in RNNs and show that the two pathways
-  of the architecture in our case study contain
-  specialized units tuned to patterns informative
-  for the task, some of which can carry activations to later time
-  steps to encode long-term dependencies.
+  that carry semantic information, and also learns to treat word types 
+  differently depending on their grammatical function and their position 
+  in the sequential structure of the sentence. In contrast, the language 
+  models are comparatively more sensitive to words with a syntactic 
+  function. Further analysis of the most informative n-gram contexts for 
+  each model shows that in comparison to the image prediction pathway,
+  the language models react more strongly to abstract contexts 
+  that represent syntactic constructions. 
+%  Furthermore, we propose methods to explore the function
+%  of individual hidden units in RNNs and show that the two pathways
+%  of the architecture in our case study contain
+%  specialized units tuned to patterns informative
+%  for the task, some of which can carry activations to later time
+%  steps to encode long-term dependencies.
 
@@
 to a specific architecture trained on specific tasks, but also provide
-pointers about how to generalize the proposed analysis to other settings.\label{generalintro}
+pointers about how to generalize the proposed analysis to other settings.
 
@@
 language model are especially sensitive to the local syntactic 
-characteristics of the input sentences.
-\todo[inline]{say something about the findings of the last experiment.}
+characteristics of the input sentences.  Further analysis of the most 
+informative n-gram contexts for each model shows that while the image 
+prediction pathway is mostly sensitive to lexical (i.e., token n-gram) 
+contexts, the language models react more strongly to abstract contexts
+(i.e., dependency relation n-grams) that represent syntactic constructions.
 

### 2017-05-31 doc/conclusions.tex
discussion of generalizability was moved to the last section

@@
 \section{Conclusion}
-\label{conclusion}
+\label{sec:conclusion}
 
@@
 
+\subsection{Generalizing to other architectures}
+
+{\bf From 4.1 (computing omission scores):}
+
+For other RNN architectures such as LSTMs \label{edit:omitgeneral}
+and their bi-directional variants, measuring the contribution
+of tokens to their predictions (or the omission scores)
+can be straight-forwardly computed using their hidden state 
+at the last time step used for prediction. Furthermore, the technique 
+can be applied in general to other architectures which
+map variable length linguistic expressions to the same fixed dimensional
+space and perform predictions based on these embeddings. 
+This includes tree-structured Recursive Neural Network models such as the Tree-LSTM
+introduced in \namecite{kai2015treelstm} or the CNN architecture of \namecite{yoonneural2014} 
+for sentence classification. In both cases the pre-softmax activations can be extracted 
+from the models as the representations of the full and partial sentences.
+
+{\bf From 4.3 (beyond lexical cues):}
+The analysis we described here takes
+the omission scores as input data, therefore it can be potentially applied 
+to any architecture for which the omission scores can be computed. However,
+the presented analysis and results regarding word positions can only be meaningful
+for Recurrent Neural Networks as they compute their representations sequentially and are not
+limited by fixed window sizes.\footnote{CNNs with multi-word filters
+and tree-structured recursive neural networks do not incrementally build representations
+of sentences in a left-to-right or right-to-left fashion. 
+Bi-directional RNNs, on the other hand, are sensitive by word-order and can potentially
+learn to handle the same word in different positions differently. \label{edit:foot}}
+
+{\bf From old 6 (top-k-context), not sure if it's applicable to the mutual information measure:}
+This method can be straight-forwardly applied to various 
+RNN architectures such as Elman networks or LSTMs 
+as it only requires storing the activation values for hidden units and 
+their corresponding context. For architectures with $n$ hidden layers
+one could extract multiple activation vectors $M_i^{1}, \ldots, M_i^{n}$
+for each unit, and perform analysis on each of them separately.
+A limitation of the generalizability of our analysis is that in the case of 
+bi-directional architectures the interpretation of the features
+extracted by the RNNs that process the input tokens in the reversed order
+might be hard from a linguistic point of view.
+
+A limitation of the generalizability of our 
+analysis is that in the case of 
+bi-directional architectures the interpretation of the features
+extracted by the RNNs that process the input tokens in the reversed order
+might be hard from a linguistic point of view.
+
 % \iffalse

### 2017-05-31 doc/response.tex
parser

@@
 introduced a uni-modal language model (referred to as the {\sc LM} model in the text) 
-and reported the results of all the experiments on this model as well. Also following 
+and reported the results of the experiments on this model as well. Also following 
 your recommendation, we have removed the exploratory parts of (the old) Section 6, 
@@
 to include relevant papers that came out while this manuscript was under review, and expanded Discussion (Section~\ref{sec:conclusion}) to discuss in more detail the relevance and generalizability of our techniques.
+We reran all the experiments with a different, more reliable parser,
+which caused minor differences in the details of the figures.
 
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{address}
+\textsc{Author Response:}  
+The results from Figure~\ref{fig:omission-imaginet} are indeed
+unsurprising. We should expect any method which works well to recover
+such expected results. In addition, there are, however, other findings
+which we consider far from obvious: for example the fact that the
+{\sc Visual} pathway assigns very different importance to the same
+word type in different grammatical functions (see See
+Section~\ref{sec:beyondlexical}). We have tried to highlight the
+non-trivial insights more in the paper.
+
+%\todo[inline]{address}
 \end{quote}

### 2017-05-31 doc/categories.tex

@@
 
-For other RNN architectures such as LSTMs \label{edit:omitgeneral}
-and their bi-directional variants, measuring the contribution
-of tokens to their predictions (or the omission scores)
-can be straight-forwardly computed using their hidden state 
-at the last time step used for prediction. Furthermore, the technique 
-can be applied in general to other architectures which
-map variable length linguistic expressions to the same fixed dimensional
-space and perform predictions based on these embeddings. 
-This includes tree-structured Recursive Neural Network models such as the Tree-LSTM
-introduced in \namecite{kai2015treelstm} or the CNN architecture of \namecite{yoonneural2014} 
-for sentence classification. In both cases the pre-softmax activations can be extracted 
-from the models as the representations of the full and partial sentences.
+%For other RNN architectures such as LSTMs \label{edit:omitgeneral}
+%and their bi-directional variants, measuring the contribution
+%of tokens to their predictions (or the omission scores)
+%can be straight-forwardly computed using their hidden state 
+%at the last time step used for prediction. Furthermore, the technique 
+%can be applied in general to other architectures which
+%map variable length linguistic expressions to the same fixed dimensional
+%space and perform predictions based on these embeddings. 
+%This includes tree-structured Recursive Neural Network models such as the Tree-LSTM
+%introduced in \namecite{kai2015treelstm} or the CNN architecture of \namecite{yoonneural2014} 
+%for sentence classification. In both cases the pre-softmax activations can be extracted 
+%from the models as the representations of the full and partial sentences.
 
@@
 discriminate between occurrences of a given word in different positions and 
-grammatical functions. The analysis we described here takes \label{edit:beyonlexicalgeneral}
-the omission scores as input data, therefore it can be potentially applied 
-to any architecture for which the omission scores can be computed. However,
-the presented analysis and results regarding word positions can only be meaningful
-for Recurrent Neural Networks as they compute their representations sequentially and are not
-limited by fixed window sizes.\footnote{CNNs with multi-word filters
-and tree-structured recursive neural networks do not incrementally build representations
-of sentences in a left-to-right or right-to-left fashion. 
-Bi-directional RNNs, on the other hand, are sensitive by word-order and can potentially
-learn to handle the same word in different positions differently. \label{edit:foot}}
+grammatical functions.
+
+%The analysis we described here takes
+%the omission scores as input data, therefore it can be potentially applied 
+%to any architecture for which the omission scores can be computed. However,
+%the presented analysis and results regarding word positions can only be meaningful
+%for Recurrent Neural Networks as they compute their representations sequentially and are not
+%limited by fixed window sizes.\footnote{CNNs with multi-word filters
+%and tree-structured recursive neural networks do not incrementally build representations
+%of sentences in a left-to-right or right-to-left fashion. 
+%Bi-directional RNNs, on the other hand, are sensitive by word-order and can potentially
+%learn to handle the same word in different positions differently. \label{edit:foot}}
 

### 2017-05-31 doc/conclusions.tex
discussion

--- 2017-05-16 doc/conclusions.tex
@@
-\section{Conclusion}
-\label{conclusion}
+\section{Discussion}
+\label{sec:conclusion}
 
@@
 activations to later time steps to encode long-term dependencies.
+
+\subsection{Generalizing to other architectures}
+
+%{\bf From 4.1 (computing omission scores):}
+
+For other RNN architectures such as LSTMs \label{edit:omitgeneral}
+and their bi-directional variants, measuring the contribution
+of tokens to their predictions (or the omission scores)
+can be straight-forwardly computed using their hidden state 
+at the last time step used for prediction. Furthermore, the technique 
+can be applied in general to other architectures which
+map variable length linguistic expressions to the same fixed dimensional
+space and perform predictions based on these embeddings. 
+This includes tree-structured Recursive Neural Network models such as the Tree-LSTM
+introduced in \namecite{kai2015treelstm} or the CNN architecture of \namecite{yoonneural2014} 
+for sentence classification. 
+%In both cases the pre-softmax activations can be extracted 
+%from the models as the representations of the full and partial sentences.   
+%{\bf From 4.3 (beyond lexical cues):}
+However,
+the presented analysis and results regarding word positions can only be meaningful
+for Recurrent Neural Networks as they compute their representations sequentially and are not
+limited by fixed window sizes.
+% \footnote{CNNs with multi-word filters
+% and tree-structured recursive neural networks do not incrementally build representations
+% of sentences in a left-to-right or right-to-left fashion. 
+% Bi-directional RNNs, on the other hand, are sensitive by word-order and can potentially
+% learn to handle the same word in different positions differently. \label{edit:foot}}
+
+%{\bf From old 6 (top-k-context), not sure if it's applicable to the mutual information measure:}
+% This method can be straight-forwardly applied to various 
+% RNN architectures such as Elman networks or LSTMs 
+% as it only requires storing the activation values for hidden units and 
+% their corresponding context. For architectures with $n$ hidden layers
+% one could extract multiple activation vectors $M_i^{1}, \ldots, M_i^{n}$
+% for each unit, and perform analysis on each of them separately.
+% A limitation of the generalizability of our analysis is that in the case of 
+% bi-directional architectures the interpretation of the features
+% extracted by the RNNs that process the input tokens in the reversed order
+% might be hard from a linguistic point of view.
+
+A limitation of the generalizability of our 
+analysis is that in the case of 
+bi-directional architectures the interpretation of the features
+extracted by the RNNs that process the input tokens in the reversed order
+might be hard from a linguistic point of view.
+
+\subsection{Future directions}
 

### 2017-05-31 doc/intro.tex
minor

@@
   selective attention to lexical categories and grammatical functions
-  that carry semantic information, and also learns to treat word types 
+  that carry semantic information, and learns to treat word types 
   differently depending on their grammatical function and their position 

### 2017-05-31 doc/categories.tex
boxplot footnote

@@
 the interquartile range (i.e.\ 75th percentile - 25th percentile); the
-points are outliers.}  The general trend is that for the {\sc
+points are outliers.\label{ft:boxplots}}  The general trend is that for the {\sc
   Visual} pathway, the omission scores are high for a small subset of

### 2017-05-31 doc/categories.tex
Plots.

@@
 \label{subsec:omission-text-vis}
-\begin{figure*}[t]
+\begin{figure}[!htbp]
+\centering
 %\hspace*{-0.3in}
-\setlength{\tabcolsep}{0pt}
-  \begin{tabular}{cc}
-  \includegraphics[scale=0.6]{imaginet-omission-pos-boxplot.png} &
-  \includegraphics[scale=0.6]{imaginet-omission-dep-boxplot.png} \\
-  \end{tabular}
+%\setlength{\tabcolsep}{0pt}
+%  \begin{tabular}{cc}
+  \includegraphics[scale=0.9]{imaginet-omission-pos-boxplot.png}
+
+  \includegraphics[scale=0.9]{imaginet-omission-dep-boxplot.png} 
+%  \end{tabular}
 
@@
   (right), for the {\sc Textual} and {\sc Visual} pathways and for
-  {\sc LM}. Only labels which occur at least 500 times are included.}
+  {\sc LM}. Only labels which occur at least 1250 times are included.}
 \label{fig:omission-imaginet}
-\end{figure*}
+\end{figure}
 
@@
   \caption{Distributions of log ratios of omission scores of {\sc Textual} to {\sc Visual} per
-    POS (left) and dependency labels (right). Only labels which occur at least 500 times are included.}
+    POS (left) and dependency labels (right). Only labels which occur at least 1250 times are included.}
 \label{fig:omission-imaginet-ratio}
@@
 zero with the {\sc Textual} pathway. We see that in relative terms,
-{\sc Visual} is more sensitive to adjectives (JJ), nouns (NNS, NN) and
-numerals (CD), and {\sc Textual} to the prepositions (TO, IN),
-determiners (DT, WDT) and pronouns (PRP) as well as some types of
-verbs (VBZ, VBP).
+{\sc Visual} is more sensitive to adjectives (JJ), nouns (NNS, NN),
+numerals (CD) and participles (VBN), and {\sc Textual} to determiners
+(DT), pronouns (PRP), prepositions (IN) as well finite verbs
+(VBZ, VBP).
 
@@
 relative importance of dependency relations: {\sc Visual} pays most
-attention to the relations {\sc nsubjpass, attr, amod, nsubj, root,
+attention to the relations {\sc amod, nsubj, root,
   compound, dobj, nummod}
-whereas {\sc Textual} is more sensitive to {\sc det, mark, prep, aux,
-  auxpass, cc, poss, advmod, prt, relcl}.
+whereas {\sc Textual} is more sensitive to {\sc det, prep, aux, cc, poss, advmod, prt, relcl}.
 As expected, {\sc Visual} is more focused on grammatical
@@
   \caption{Distributions of log ratios of omission scores of {\sc LM} to {\sc Textual} per
-    POS (left) and dependency labels (right). Only labels which occur at least 500 times are included.}
+    POS (left) and dependency labels (right). Only labels which occur at least 1250times are included.}
 \label{fig:omission-imaginet-quotient}

### 2017-05-31 doc/related.tex
related

@@
 by erasing them from the representation and analyzing how this affects the model.
+They focus on text-only tasks and do not take other 
+modalities such as visual input into account.
 \namecite{adi2016fine} take an alternative approach by introducing prediction tasks 
 to analyze information encoded in sentence embeddings about sentence length, 
-sentence content and word order. Finally, \namecite{linzen2016assessing} examine 
+sentence content and word order. 
+Finally, \namecite{linzen2016assessing} examine 
 the acquisition of long-distance dependencies through the study of number agreement 
@@
 signal (that is, whether the subject is plural or singular), but simple language models still capture 
-the majority of test cases. All these models focus on text-based tasks and do not take other 
-modalities such as visual input into account.
+the majority of test cases. While they focus on an in-depth analysis of a single phenomenon, in our work
+we are interested in methods which make it possible to uncover a broad variety of patterns of bebavior in RNNs.
 

### 2017-05-31 doc/intro.tex
final edits on intro

@@
   RNNs from a linguistic point of view and explore the types of 
-  linguistic structure they learn. As a case study, we use a multi-task gated
+  linguistic structure they learn. As a case study, we use a standard 
+  standalone language model, and a multi-task gated
   recurrent network architecture consisting of two parallel pathways with
-  shared word embeddings trained on predicting the
+  shared word embeddings; The {\sc Visual} pathway is trained on predicting the
   representations of the visual scene corresponding to an input
-  sentence, and predicting the next word in the same sentence, 
-  in addition to a standard standalone language model. 
+  sentence, whereas the {\sc Textual} pathway is trained to predict the 
+  next word in the same sentence. 
   We propose a method for estimating the amount of contribution
   of individual tokens in the input to the final prediction of the networks.   
-  Using this method, we show that the image prediction pathway pays
+  Using this method, we show that the {\sc Visual} pathway pays
   selective attention to lexical categories and grammatical functions
@@
   function. Further analysis of the most informative n-gram contexts for 
-  each model shows that in comparison to the image prediction pathway,
+  each model shows that in comparison to the {\sc Visual} pathway,
   the language models react more strongly to abstract contexts 
@@
 larger linguistic units directly from the input data, without feature
-engineering. Variations of the RNN architectures have been applied in
+engineering. Variations of the RNN architecture have been applied in
 several NLP domains such as parsing \cite{vinyals2015grammar} and
@@
 object segmentation \cite{visin2015reseg}. RNNs are also important
-components of systems integrating Vision and Language, e.g.\ image
+components of systems integrating vision and language, e.g.\ image
 \cite{karpathy2015deep} and video captioning \cite{yu2015video}.
@@
 word in the descriptions. This particular
-architecture allows a comparative analysis of the hidden activations
+architecture allows a comparative analysis of the hidden activation
 patterns between networks trained on two different tasks, while
@@
 next symbol in a sequence are relatively well understood, and there
-have been some attempts to analyze their internal states e.g.
-\cite{elman1991distributed,karpathy2015visualizing}. In
+have been some attempts to analyze their internal states 
+\cite[among others]{elman1991distributed,karpathy2015visualizing}. In
 constrast, {\sc Visual} maps a complete sequence of words to
@@
 
-Our experiments show that the image prediction pathway in general 
+Our experiments show that the {\sc Visual} pathway in general 
 pays special attention to syntactic categories which carry semantic 
-content, and particularly nouns. More surprisingly, this pathway also 
+content, and particularly to nouns. More surprisingly, this pathway also 
 learns to treat word types differently depending on their 
 grammatical function and their position in the sequential structure of 
-the sentence. In contrast, the textual pathway and the standalone 
+the sentence. In contrast, the {\sc Textual} pathway and the standalone 
 language model are especially sensitive to the local syntactic 
 characteristics of the input sentences.  Further analysis of the most 
-informative n-gram contexts for each model shows that while the image 
-prediction pathway is mostly sensitive to lexical (i.e., token n-gram) 
+informative n-gram contexts for each model shows that while the 
+{\sc Visual} pathway is mostly sensitive to lexical (i.e., token n-gram) 
 contexts, the language models react more strongly to abstract contexts

### 2017-05-31 doc/related.tex
final edits on related work

@@
 layer. His analysis shows that these distributed representations  
-encode lexical categories, grammatical relationships and hierarchical constituent 
-structures. \namecite{giles1992extracting} trains RNNs similar to Elman
+encode lexical categories, grammatical relations and hierarchical constituent 
+structures. \namecite{giles1992extracting} train RNNs similar to Elman
 networks on strings generated by small 
 deterministic regular grammars with the objective to recognize grammatical and reject ungrammatical 
-strings, and develops the \emph{dynamic state partitioning} technique to extract the learned 
+strings, and develop the \emph{dynamic state partitioning} technique to extract the learned 
 grammar from the networks in the form of deterministic finite state
@@
 Short-Term Memory networks (LSTM) \cite{hochreiter1997long} 
-for phrase level sentiment analysis 
+for phrase-level sentiment analysis 
 and present novel methods 
@@
 At the decoding side this neural module assigns weights 
-$\alpha_{1},\ldots,\alpha_{i}$ to the hidden states of the decoder 
-$h_{1},\ldots,h_{i}$, which allows the decoder to selectively pay varying degrees of
+to the hidden states of the decoder, which allows the decoder to selectively pay varying degrees of
 attention to different phrases in the source sentence at different decoding time-steps.
@@
 GRUs.  They show that certain dimensions in the RNN hidden activation
-vectors have specific and interpretable functions. One of the goals of
-the present paper is also to explore the specific functions encoded
-by individual hidden units, but in a linguistically-informed way.
-
+vectors have specific and interpretable functions. Similarly,
 \namecite{li2015convergent} use a Convolutional Neural Networks (CNN)

### 2017-05-31 doc/models.tex
final edits on models

@@
 arises from the fact that they overwrite their hidden states at every
-time step with a new value computed from the current input $x_{t}$ and
+time step with a new value computed from the current input $\mathbf{x_{t}}$ and
 the previous hidden state $\mathbf{h_{t-1}}$. Similarly to LSTMs,
@@
 \end{equation}
+where $\mathbf{W}$, $\mathbf{U}$, $\mathbf{W}_z$, $\mathbf{U}_z$, $\mathbf{W}_r$ and $\mathbf{U}_r$ are learnable parameters.
 
@@
 meaning representations from textual and visual input.  It acquires 
-linguistic knowledge via comprehension and production, by receiving a
+linguistic knowledge through language comprehension, by receiving a
 description of a scene and trying to visualise it through predicting a visual
@@
 be seen from the figure, the model consists of two GRU pathways, 
-{\sc Textual} and {\sc Visual}, with a shared word-embedding matrix. 
+{\sc Textual} and {\sc Visual}, with a shared word embedding matrix. 
 The inputs to the model are pairs of image descriptions and their 

### 2017-05-31 doc/experiments.tex
last edits on experiments

@@
 %the linguistic features encoded by individual hidden dimensions. 
-In Section~\ref{sec:computeomission} we introduce \emph{omission scores},
+In Section~\ref{sec:computeomission} we introduce \emph{omission score},
 a metric to measure the contribution of each token to the prediction
-of the networks and in Section~\ref{sec:omitimaginet} we analyze how
-these scores are distributed over dependency relations and
+of the networks, and in Section~\ref{sec:omitimaginet} we analyze how
+omission scores are distributed over dependency relations and
 part-of-speech categories. 
@@
 the importance of words for the different networks depend on the words themselves,
-their sequential position and their grammatical function in the sentences.
-\todo[inline]{add a short description of the experiments in sections 5.4 and 5.5.}
+their sequential position, and their grammatical function in the sentences.
+Finally, in Section~\ref{sec:contexts} we systematically compare the types of 
+n-gram contexts that trigger individual dimensions in the hidden layers of the 
+networks, and discuss their level of abstractness.
 
 In all these experiments we report our findings based on the {\sc Imaginet}
-model, whenever appropriate compare to our two other models {\sc LM} and {\sc Sum}, 
-and discuss the generalizabilty of our methods to other architectures.
-\label{edit:experimentsgeneral}
+model, and whenever appropriate compare it to our two other models {\sc LM} and {\sc Sum}.
 For all the experiments, we trained the models on the training portion of the

### 2017-05-31 doc/categories.tex
last edits on categories.tex

@@
 from a linguistic point of view, and focus
-on the high-level understanding of what parts of input sentence the networks
+on the high-level understanding of what parts of the input sentence the networks
 pay most attention to. Furthermore, we investigate if the networks
@@
 representation. For example, the distance
-between $\mathbf{h}_\text{end}(${\it the black dog is running}$)$
-and $\mathbf{h}_\text{end}(${\it the dog is running}$)$ determines
+between $\mathbf{h}_\text{end}($``{\it the black dog is running}''$)$
+and $\mathbf{h}_\text{end}($``{\it the dog is running}''$)$ determines
 the importance of {\it black} in the first sentence. We introduce the
@@
 
-\noindent Figure~\ref{fig:omissionex} demonstrates the omission
+Figure~\ref{fig:omissionex} demonstrates the omission
 scores for the {\sc LM}, {\sc Visual} and {\sc Textual} models for an
-example caption. The omission scores are
-plotted for the
-full sentence and for the sentence with the word with the highest
-$\mathrm{omission}(i,S)$ removed. 
+example caption.  
 Figure~\ref{fig:omissionexpic} shows the images retrieved by {\sc
@@
 The omission scores for {\sc Visual} show that the model paid attention
-mostly to {\it baby} and {\it bed} and slightly to {\it laptop} and
+mostly to {\it baby} and {\it bed} and slightly to {\it laptop}, and
 retrieved an image depicting a baby sitting on a bed with a laptop.
@@
 numerals (CD) and participles (VBN), and {\sc Textual} to determiners
-(DT), pronouns (PRP), prepositions (IN) as well finite verbs
+(DT), pronouns (PRP), prepositions (IN) and finite verbs
 (VBZ, VBP).
@@
 
-It is worth noting, however, the relatively low omission scores for verbs in case \label{edit:generality}
+It is worth noting, however, the relatively low omission scores for verbs in the case 
 of {\sc Visual}. One might expect that the task of image prediction from
 descriptions requires general language understanding and so high omission
-scores for all content words in general, however, the results
+scores for all content words in general; however, the results
 suggest that this setting is not optimal for learning useful representations of verbs,
@@
   \caption{Distributions of log ratios of omission scores of {\sc LM} to {\sc Textual} per
-    POS (left) and dependency labels (right). Only labels which occur at least 1250times are included.}
+    POS (left) and dependency labels (right). Only labels which occur at least 1250 times are included.}
 \label{fig:omission-imaginet-quotient}
@@
 Models that utilize the sequential structure of language 
-have the capacity to interpret the same word-type differently depending on
+have the capacity to interpret the same word type differently depending on
 the context. The omission score distributions in Section \ref{sec:omitimaginet} 
@@
 The raw $R^2$ scores show that for the language models {\sc LM} and 
-{\sc Textual}, the word-type predicts the omission-score to much smaller 
+{\sc Textual}, the word type predicts the omission-score to much smaller 
 degree compared to {\sc Visual}. Moreover, adding information about 
@@
 
-Figure~\ref{fig:rsquared} offers a different view of the data to, showing
+Figure~\ref{fig:rsquared} offers a different view of the data, showing
 the increase or decrease in $R^2$ for the models relative to {\sc LR~+pos} 
@@
 dependency labels, the {\sc Visual} model's omission scores are the
-hardest to predict of the four models suggesting that the model may be
+hardest to predict of the four models. This suggests that {\sc Visual} may be
 encoding additional structural features not captured by these predictors. 
@@
   \includegraphics[scale=0.35]{top_words.pdf}
-  \caption{Distribution of omission scores per deprel label for the selected word types.}
+  \caption{Distribution of omission scores per dependency label for the selected word types.}
   \label{fig:top_words}
@@
 better, on average, {\sc LR~+dep} predicted their omission scores 
-compared to {\sc LM~word}.
+compared to {\sc LR~word}.
 
@@
 impact the network's representation depending on what grammatical
-function they fulfill. They all have large omission scores when they
+function they fulfil. They all have large omission scores when they
 occur as {\sc nsubj} (nominal subject) or {\sc root}, likely due to the fact that these
@@
  \includegraphics[scale=0.4]{position-coef.pdf}
+ \vspace{-2cm}
  \caption{Coefficients on the y-axis of {\sc LR full} corresponding to the
@@
 tokens in positions closer to the end of the sentence to be more important. 
-On the other hand in English the information structure of a single
+On the other hand, in English the information structure of a single
 sentence is expressed via linear ordering: the {\sc topic} of a 
@@
 contain new information in these cases. In our setting of image captions 
-however, sentences are not part of larger discourse, it is sentence
+however, sentences are not part of a larger discourse; it is sentence
 initial material that typically contains the most important 

### 2017-05-31 doc/conclusions.tex
last edits on conclusion

@@
 proposed techniques to explore what kinds of linguistic features the models
-learn to exploit on top of lexical cues.
+learn to exploit beyond lexical cues.
 
@@
 our analyses of the hidden activation patterns show that the {\sc Visual} model
-learns an abstract representation of the information structure of the
-language, and pays selective attention to lexical categories and
+learns an abstract representation of the information structure of a single
+sentence in the language, and pays selective attention to lexical categories and
 grammatical functions that carry semantic information. In contrast,
@@
 specialized units which are tuned to both lexical and structural
-patterns that are useful for the task at hand, some of which can carry
-activations to later time steps to encode long-term dependencies.
+patterns that are useful for the task at hand.
 
@@
 can be applied in general to other architectures which
-map variable length linguistic expressions to the same fixed dimensional
+map variable-length linguistic expressions to the same fixed dimensional
 space and perform predictions based on these embeddings. 
 This includes tree-structured Recursive Neural Network models such as the Tree-LSTM
-introduced in \namecite{kai2015treelstm} or the CNN architecture of \namecite{yoonneural2014} 
+introduced in \namecite{kai2015treelstm}, or the CNN architecture of \namecite{yoonneural2014} 
 for sentence classification. 
@@
 analysis is that in the case of 
-bi-directional architectures the interpretation of the features
+bi-directional architectures, the interpretation of the features
 extracted by the RNNs that process the input tokens in the reversed order
@@
 \cite{sutskever2014sequence} or the purely distributional
-sentence embedding system of \cite{kiros2015skip}. A number of
-recurrent neural models rely on a so called attention mechanism, first
+sentence embedding system of \namecite{kiros2015skip}. A number of
+recurrent neural models rely on a so-called attention mechanism, first
 introduced by \namecite{bahdanau2014neural} under the name of soft
-alignment. In these networks attention is explicitly represented and it
+alignment. In these networks attention is explicitly represented, and it
 would be interesting to see how our method of discovering implicit
 attention, the omission score, compares. For future work we also propose to
-collect data where humans assess the importance of each word in sentences
-and explore the relationsship between \emph{omission scores} for various models
+collect data where humans assess the importance of each word in a sentence
+and explore the relationship between omission scores for various models
 and human annotations.\label{edit:humanjudgement} Finally, one of the benefits

### 2017-05-31 doc/response.tex
boxplot footnote
final edits to response letter

@@
 emphasise our contributions, updated Related Work (Section~\ref{sec:related}) 
-to include relevant papers that came out while this manuscript was under review, and expanded Discussion (Section~\ref{sec:conclusion}) to discuss in more detail the relevance and generalizability of our techniques.
+to include relevant papers that came out while this manuscript was under review, and 
+expanded Discussion (Section~\ref{sec:conclusion}) to discuss in more detail the 
+relevance and generalizability of our techniques. 
 We reran all the experiments with a different, more reliable parser,
@@
 \begin{quote}
-\textsc{Author Response:}  \todo[inline]{explain the updated graph}
+\textsc{Author Response:}  We have provided an interpretation of the boxplots 
+in footnote~\ref{ft:boxplots} on page~\pageref{ft:boxplots}.
 \end{quote}
@@
 \begin{quote}
-\textsc{Author Response:}  We removed this part and added an expanded discussion of the generalizability of our techniques to Discussion (Section~\ref{sec:conclusion}).
+\textsc{Author Response:}  We removed this part and added an 
+expanded discussion of the generalizability of our techniques to 
+Discussion (Section~\ref{sec:conclusion}).
 \end{quote}
@@
 \textsc{Author Response:}  As mentioned before, we have reorganised 
-the experiments and remove the exploratory parts, including the old Section 6.3. 
+the experiments and removed the exploratory parts, including the old Section 6.3. 
 We address the issue of encoding grammatical functions in the revised manuscript in 
@@
 \begin{quote}
-\textsc{Author Response:}  There was no $\lambda$ in that equation, but in 
+\textsc{Author Response:}  There was no lambda in that equation, but in 
 Section~\ref{sec:imaginet} we report the value we use for the parameter 
@@
 unsurprising. We should expect any method which works well to recover
-such expected results. In addition, there are, however, other findings
+such expected results. In addition, however, there are other findings
 which we consider far from obvious: for example the fact that the
@@
 compare the results head-to-head. Thus {\sc Textual} serves
-mostly to show that our methods give reasonably, expected answers,
+mostly to show that our methods give reasonable, expected answers,
 while with {\sc Visual} we can generate novel, previously unfamiliar

### 2017-07-11 src/omission.py
Rename README.md, change paths in src/omission.py

@@ -7,12 +7,14 @@ from scipy.spatial.distance import cosine
 import imaginet.task
 import numpy
 
-root = '/home/gchrupal/reimaginet/'
+root = '/home/gchrupal/cl-resubmit/'
 data_path = '/home/gchrupal/cl-resubmit/data/'
-model_im_path = root + "/run-lm_visual_vanilla-1/model.r.e7.zip"
-model_lm_path = root + "/run-lm-0/model.r.e6.zip"
-model_sum_path = root + "/run-vectorsum-0/model.r.e7.zip"
-
+#model_im_path = root + "/run-lm_visual_vanilla-1/model.r.e7.zip"
+#model_lm_path = root + "/run-lm-0/model.r.e6.zip"
+#model_sum_path = root + "/run-vectorsum-0/model.r.e7.zip"
+model_im_path = root + "models/imaginet.zip"
+model_lm_path = root + "models/lm.zip"
+model_sum_path = root + "models/sum.zip"
 
 def dump_activations(sent, models):
     for name, model, task in models:

### 2017-07-12 doc/cover.tex
adding style files and cover letter

\documentclass{letter}
\signature{Akos Kadar \\ Grzegorz Chrupala \\ Afra Alishahi}

\begin{document}
\begin{letter}{}
\opening{Dear Dr.~Merlo,}

Thank you for informing us about the outcome of the review process, we are very pleased that the paper has been accepted to be published in Computational Linguistics.

Attached please find the final version of the article. We have applied the formatting requirements as specified in the author guidelines, and updated the references. Please let us know if any other information is required.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2017-07-17 doc/mi-stat/stat.R
Cleanup files.

@@ -1,58 +1,0 @@
-setwd("/home/gchrupala/repos/naacl2016/mi-stat")
-library(dplyr)
-library(ggplot2)
-
-
-data <- data.frame()
-for (n in 1:3) {
-  for (condition in c("deprel","token")) {
-    file <- paste("MI/new_results/",
-                  condition,
-                  "s",
-                  n,
-                  ".csv", sep="")
-    scores <- read.csv(file)
-    cond <- if (condition=="token") {"word"} else {"deprel"}
-    data <- rbind(data, data.frame(pathway="Visual",
-                           n=n,
-                           condition=cond,
-                           mi=as.numeric(scores[1,2:1025])))          
-    data <- rbind(data, data.frame(pathway="Textual",
-                                 n=n,
-                                 condition=cond,
-                                 mi=as.numeric(scores[2,2:1025])))          
-    }
-  }
-
-ggplot(data %>% filter(n==1), aes(x=condition, y=mi, color=pathway)) + 
-  geom_boxplot()
-
-library(boot)
-
-log_ratio <- function(data, ix) {
-  w <- log(median(filter(data[ix,], pathway=="Textual", condition=="word")$mi) / 
-        median(filter(data[ix,], pathway=="Visual", condition=="word")$mi),base=exp(1))
-  d <- log(median(filter(data[ix,], pathway=="Textual", condition=="deprel")$mi) / 
-             median(filter(data[ix,], pathway=="Visual", condition=="deprel")$mi),base=exp(1))
-  return(c(w=w,d=d))
-}
-
-
-library(reshape2)
-
-data_boot <- data.frame()
-for (i in 1:3) {  
-  res <- boot(data %>% filter(n==i), log_ratio, R=5000)
-  data_boot <- rbind(data_boot, data.frame(n=i, condition="word", log.ratio=res$t[,1]))
-  data_boot <- rbind(data_boot, data.frame(n=i, condition="deprel", log.ratio=res$t[,2]))
-}
-ggplot(data=data_boot,
-       aes(x=as.factor(n), color=condition, y=log.ratio)) +
-  geom_boxplot() +
-  coord_flip() +
-  xlab("n-gram range") +
-  ylab(expression(log(MI[C]^T / MI[C]^V ))) +
-  theme(text=element_text(size=20)) +
-  ggsave(file="bootstrappedMI.png",
-         width=6, height=4)
-

### 2017-07-17 doc/omission-stat/analysis.R

@@ -1,180 +1,0 @@
-# INSTALLATION
-
-#install.packages("Rcpp", dependencies=TRUE, repos="http://cran.r-project.org" )
-#install.packages("dplyr", dependencies=TRUE, repos="http://cran.r-project.org" )
-#install.packages("ggplot2", dependencies=TRUE, repos="http://cran.r-project.org")
-
-library("dplyr")
-library("ggplot2")
-# READING DATA
-
-read.data <- function(f) {
-  data <- read.csv(f, col.names= c('omit','id','word','pos','deprel', 'score')) %>%
-  select(-omit) %>% na.omit()
-  return(data[! data[,'pos'] %in%c('<END>','END'),])
-}
-data_v <- read.data('../data/visual_full_omission.csv')
-data_v$pathway <- "visual"
-data_v$deprel <- reorder(data_v$deprel, data_v$score, median)
-data_v$pos <- reorder(data_v$pos, data_v$score, median)
-data_t <- read.data('../data/textual_full_omission.csv')
-data_t$pathway <- "textual"
-data_t$deprel <- reorder(data_t$deprel, data_v$score, median)
-data_t$pos <- reorder(data_t$pos, data_v$score, median)
-
-data <- rbind(data_t , data_v) %>% mutate(pathway=as.factor(pathway))
-
-filter_by_count <- function(dat, col, count) {
-  dat[dat[,col] %in% names(table(dat[,col]))[table(dat[,col]) > count],]
-}
-
-MINCOUNT<-500
-subset <- filter_by_count(data %>% filter(pathway=='visual'),"pos", MINCOUNT)
-
-
-
-# OMISSION BOXPLOTS
-
-ggplot(subset, aes(x=pos, y=score)) +
-  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
-  coord_flip() +
-  theme(text=element_text(size=15))
-ggsave(file="visual-omission-pos-boxplot.png",width=5, height=5)
-
-subset <- filter_by_count(data %>% filter(pathway=='visual'),"deprel",MINCOUNT)
-ggplot(subset, aes(x=deprel, y=score)) +
-  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
-  coord_flip() +
-  theme(text=element_text(size=15))
-ggsave(file="visual-omission-deprel-boxplot.png",width=5, height=5)
-
[…]

### 2017-07-17 doc/omission-stat/me.R

@@ -1,25 +1,0 @@
-library(dplyr)
-library(ggplot2)
-library(lme4)
-
-# Visualization of predicting omission scores using various predictors
-data <- read.table(file="predictor_rsq.txt", header=TRUE) 
-for (model in levels(data$model)) {
-  data[data$model==model,"score"] = data[data$model==model,"rsq"]/data[data$model==model & data$predictors=="word","rsq"]
-}
-data <- data %>% mutate(predictors = reorder(predictors, rsq))
-ggplot(data, aes(x=model, color=predictors, y=score)) + geom_boxplot() +  theme(aspect.ratio=2/3, text=element_text(size=25)) +
-   ylab("R Squared relative to word")
-ggsave("position-new.png")
-
-data_coef<- read.table(file="position_coef.txt", header=TRUE)
-values <- c("first","second","third","middle","antepenult","penult","last")
-data_coef$predictor <- factor(data_coef$predictor, levels=values, ordered=TRUE)
-ggplot(data_coef, aes(x=as.numeric(predictor), y=coef, color=model)) + geom_point() + geom_line() +
-  xlab("Position") +
-  ylab("Coefficient") +
-  scale_x_continuous(breaks=1:length(values), 
-                     labels=values) +
-  theme(aspect.ratio=2/3, text=element_text(size=25))
-ggsave("position-coef.png")
-

### 2017-07-17 doc/makezip.sh
Script to make zip file with Latex sources.

@@ -1,0 +1,2 @@
+zip source.zip *.pdf *.png *.jpg paper.tex double_paper.tex intro.tex related.tex models.tex experiments.tex categories.tex dimensions.tex conclusions.tex refs.bib clv3.cls compling.bst
+

### 2017-07-17 doc/makezip.sh
copyright makezip

@@ -1,2 +1,2 @@
-zip source.zip *.pdf *.png *.jpg paper.tex double_paper.tex intro.tex related.tex models.tex experiments.tex categories.tex dimensions.tex conclusions.tex refs.bib clv3.cls compling.bst
+zip source.zip *.png *.jpg imaginet.pdf bootstrappedMI2.pdf raw_mutual.pdf position-new.pdf position-coef.pdf omissionex.pdf top_words.pdf bootstrappedMI.pdf paper.tex double_paper.tex intro.tex related.tex models.tex experiments.tex categories.tex dimensions.tex conclusions.tex refs.bib clv3.cls compling.bst
 

### 2017-07-17 doc/double_paper.tex
Added double-spaced version of paper.

%\documentclass[shortpaper]{clv3}  %for single-spaced version
\documentclass[manuscript]{clv3} %for double-spaced version

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{algorithm}

\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{todonotes}
\usepackage{url}
\usepackage{array,multirow,graphicx}
\usepackage{rotating}
\usepackage{times}
\usepackage{latexsym}
\usepackage{arydshln}
\usepackage{enumitem}
\setlist{nolistsep} % needs enumitem

\issue{43}{3}{2017}

\title{Representation of linguistic form and function in recurrent neural networks}
\runningtitle{Representation of linguistic form and function in recurrent neural networks}

\author{Ákos Kádár \thanks{Tilburg Center for Cognition and Communication,
Tilburg University, 5000 LE Tilburg, The Netherlands, 
E-mail: \texttt{\{a.kadar, g.chrupala, a.alishahi\}@uvt.nl}} }  
\affil{Tilburg University}

\author{Grzegorz Chrupała \footnotemark[1]}
\affil{Tilburg University}

\author{Afra Alishahi \footnotemark[1]}
\affil{Tilburg University}

\runningauthor{Kádár, Chrupała and Alishahi}

\historydates{Submission received: 21st July, 2016 \\
Revised version received: 5th June, 2017 \\
Accepted for publication: 7th July, 2017}

\begin{document}
\maketitle

\input{intro}
\input{related}
\input{models}
\input{experiments}
\input{conclusions}

\bibliographystyle{compling}
%\starttwocolumn http://cljournal.org/commonprobs.html says we need
%this, but the template doesn't do it
\bibliography{refs}

\end{document}
